### 2022-05-13 notebooks/Linear_algebra.Rmd
Initial commit. Includes first version of the article draft, along with several notebooks and ancillary output documents.

---
title: "Linear algebra"
output:
  html_notebook:
    number_sections: yes
---

Following this document:
https://www.math.hkust.edu.hk/~mabfchen/Math111/Week13-14.pdf

# Dot product of $\mathbb{R}^n$

All clear here.

# Inner product spaces

**Example 2.1 (adapted).**

For
$\mathbf{\theta}_1 = \begin{bmatrix}\theta_{11}\\ \theta_{12}\\ \end{bmatrix}$,
$\mathbf{\theta}_2 = \begin{bmatrix}\theta_{21}\\ \theta_{22}\\ \end{bmatrix}$
$\in \mathbb{R}^2$,
define

$$
\langle \mathbf{\theta}_1, \mathbf{\theta}_2 \rangle =
  \tfrac{4}{3} \theta_{11} \theta_{21} -
    \tfrac{2}{3} \theta_{11} \theta_{22} -
    \tfrac{2}{3} \theta_{12} \theta_{21} +
    \tfrac{4}{3} \theta_{12} \theta_{22}.
$$

Then $\langle,  \rangle$ is an inner product on $\mathbb{R}^2$.
The properties fulfill (see handwritten notebook "Linear algebra").

For each $\theta \in \Theta$, the **norm** or **length** of $\theta$ is
defined as

$$
\lVert \mathbf{\theta} \rVert :=
  \sqrt{ \langle \mathbf{\theta}, \mathbf{\theta} \rangle }
$$

# Examples of inner product spaces

**Example 3.1.**

**IMPORTANT:**
The vector space $\mathbb{R}^n$ with this special inner product (dot product)
is called the **Euclidean $n$-space**, and the dot product is called the
**standard inner product** on $\mathbb{R}^n$.

# Representation of inner product

**Theorem 4.1 (adapted).** Let $\Theta$ be an $n$-dimensional vector space with
an inner product $\langle, \rangle$, and let $\Sigma^{-1}$ be the matrix of
$\langle, \rangle$ relative to a basis $\mathcal{B}$.
Then for any vectors $\mathbf{\theta}_1', \mathbf{\theta}_2' \in \Theta$,

$$
\langle \mathbf{\theta}_1', \mathbf{\theta}_2' \rangle =
  \mathbf{\theta}_1^T \Sigma^{-1} \mathbf{\theta}_2,
$$

where $\mathbf{\theta}_1$ and $\mathbf{\theta}_2$ are the coordinate vectors
of $\mathbf{\theta}_1'$ and $\mathbf{\theta}_2'$, respectively, i.e.,
$\mathbf{\theta}_1 = [\mathbf{\theta}_1']_\mathcal{B}$ and
$\mathbf{\theta}_2 = [\mathbf{\theta}_2']_\mathcal{B}$.

**Example 4.1 (adapted).** For the inner product of $\mathbb{R}^2$ defined by

$$
\langle \mathbf{\theta}_1', \mathbf{\theta}_2' \rangle =
  \frac{4}{3} \theta_{11} \theta_{21} -
    \frac{2}{3} \theta_{11} \theta_{22} -
    \frac{2}{3} \theta_{12} \theta_{21} +
    \frac{4}{3} \theta_{12} \theta_{22},
$$

where
$\mathbf{\theta}_1 = \begin{bmatrix} \theta_{11}\\ \theta_{12}\\ \end{bmatrix}$,
$\mathbf{\theta}_2 = \begin{bmatrix} \theta_{21}\\ \theta_{22}\\ \end{bmatrix}$
$\in \mathbb{R}^2$,
its matrix relative to the standard basis
$\mathcal{E} = \{ \mathbf{e}_1, \mathbf{e}_2 \}$ is

$$
\Sigma^{-1} = \begin{bmatrix}
    \langle \mathbf{e}_1, \mathbf{e}_1 \rangle &
      \langle \mathbf{e}_1, \mathbf{e}_2 \rangle \\
    \langle \mathbf{e}_2, \mathbf{e}_1 \rangle &
      \langle \mathbf{e}_2, \mathbf{e}_2 \rangle \\
  \end{bmatrix} =
  \begin{bmatrix}
     \tfrac{4}{3} & -\tfrac{2}{3} \\
    -\tfrac{2}{3} &  \tfrac{4}{3} \\
  \end{bmatrix}
$$

(see handwritten notebook "Linear algebra").

With a change of variables:

$$
\Bigg\{
\begin{array} \\
  \theta_{11} =             \theta_{11}' \\
  \theta_{12} = \cos \alpha \theta_{11}' + \sin \alpha \theta_{12}'
\end{array}
$$

$$
\Bigg\{
\begin{array} \\
  \theta_{21} =             \theta_{21}' \\
  \theta_{22} = \cos \alpha \theta_{21}' + \sin \alpha \theta_{22}'
\end{array}
$$

we have (see handwritten notebook "Linear algebra")

$$
\langle \mathbf{\theta}_1', \mathbf{\theta}_2' \rangle =
  \theta_{11}'\theta_{21}' + \theta_{12}'\theta_{22}' =
  \mathbf{\theta}_1'^T \mathbf{\theta}_2'.
$$

This is equivalent to choosing a new basis

$$
\mathcal{B} = \left\{
  \mathbf{u}_1 = \begin{bmatrix} 1\\ \cos \alpha\\ \end{bmatrix},
  \mathbf{u}_2 = \begin{bmatrix} 0\\ \sin \alpha\\ \end{bmatrix}
\right\}
$$

so that the inner product relative to it is the identity matrix, i.e.
(see handwritten notebook "Linear algebra"),

$$
\begin{bmatrix}
  \langle \mathbf{u}_1, \mathbf{u}_1 \rangle &
    \langle \mathbf{u}_1, \mathbf{u}_2 \rangle \\
  \langle \mathbf{u}_2, \mathbf{u}_1 \rangle &
    \langle \mathbf{u}_2, \mathbf{u}_2 \rangle \\
\end{bmatrix} =
\begin{bmatrix}
  1 & 0 \\
  0 & 1 \\
\end{bmatrix}.
$$

Let $\mathbf{P}$ the transition matrix from the standard basis
$\{ \mathbf{e}_1, \mathbf{e}_2 \}$ to the basis
$\{ \mathbf{u}_1, \mathbf{u}_2 \}$, i.e.,

$$
[ \mathbf{u}_1, \mathbf{u}_2 ] = [ \mathbf{e}_1, \mathbf{e}_2 ] \mathbf{P} =
  [ \mathbf{e}_1, \mathbf{e}_2 ]
  \begin{bmatrix}
         1      &      0      \\
    \cos \alpha & \sin \alpha \\
  \end{bmatrix}
$$

Let $\mathbf{\theta}_1$ be the coordinate vector
of the vector $\mathbf{\theta}_1'$ relative to the basis $\mathcal{B}$.
(The coordinate vector of $\mathbf{\theta}_1'$ relative to the standard basis is
$\mathbf{\theta}_1'$ itself.)
Then

$$
\mathbf{\theta}_1' = [ \mathbf{e}_1, \mathbf{e}_2 ] \mathbf{\theta}_1' =
  [ \mathbf{u}_1, \mathbf{u}_2 ]                    \mathbf{\theta}_1  =
  [ \mathbf{e}_1, \mathbf{e}_2 ] \mathbf{P}         \mathbf{\theta}_1,
$$

and it follows that

$$
\mathbf{\theta}_1' = \mathbf{P} \mathbf{\theta}_1.
$$

Similarly,

$$
\mathbf{\theta}_2' = \mathbf{P} \mathbf{\theta}_2.
$$

Note that $\mathbf{\theta}_1'^T = \mathbf{\theta}_1^T \mathbf{P}^T$,
so by Theorem 4.1,

$$
\langle \mathbf{\theta}_1',         \mathbf{\theta}_2' \rangle =
  \mathbf{\theta}_1'^T \mathbf{I}_n \mathbf{\theta}_2'         =
  \mathbf{\theta}_1'^T              \mathbf{\theta}_2',
$$

on one hand, and

$$
\langle \mathbf{\theta}_1', \mathbf{\theta}_2' \rangle                       =
  \mathbf{\theta}_1^T               \Sigma^{-1}            \mathbf{\theta}_2 =
  \mathbf{\theta}_1'^T \mathbf{P}^T \Sigma^{-1} \mathbf{P} \mathbf{\theta}_2'
$$

on the other.

**Theorem 4.2.** Let $\Theta$ be a finite-dimensional inner product space.
Let $\Sigma^{-1}$, $\mathbf{B}$ be matrices of the inner product relative to
bases $\mathcal{B}$, $\mathcal{B}'$ of $\Theta$, respectively.
If $\mathbf{P}$ is the transition matrix
from $\mathcal{B}'$ to $\mathcal{B}$[^trans-matrix-basis],
then

[^trans-matrix-basis]: According to the document I'm following,
the transition matrix if from basis $\mathcal{B}$ to $\mathcal{B}'$,
but to my understanding, in the previous example this transition matrix
is from the standard basis to $\mathcal{B}$, and so it should be here
(as in my example, the standard basis $\mathcal{E}$
plays the role of $\mathcal{B}'$ in this theorem).

$$
\mathbf{B} = \mathbf{P}^T \Sigma^{-1} \mathbf{P}.
$$

**Application of Theorem 4.2. to my bersion of example 4.1.**

In my case, the standard basis $\mathcal{E}$ plays the role of $\mathcal{B}'$
in Theorem 4.2. The inner product matrix relative to it is $\mathbf{I}_n$,
and the transition matrix from $\mathcal{E}$ to $\mathcal{B}$ is

$$
\mathbf{P} = \begin{bmatrix}
         1      &      0      \\
    \cos \alpha & \sin \alpha \\
  \end{bmatrix}
$$

It is true that:

$$
\mathbf{I}_n = \mathbf{P}^T \Sigma^{-1} \mathbf{P}
$$

(see handwritten notebook for derivation).

# Cauchy-Schwarz inequality

**IMPORTANT HERE!**

The **angle $\alpha$ between $u$ and $v$** is defined by

$$
\cos \alpha = \frac{
    \langle \mathbf{u}, \mathbf{v} \rangle
  } {
    \lVert \mathbf{u} \rVert \lVert \mathbf{v} \rVert
  }.
$$

The angle exists and is unique.

(**IMPORTANT NOTE!!** $\cos \alpha$ apparently is independent of the inner
product, i.e., which basis it is computed in).

# Orthogonality

Let $\Theta$ be an inner product space.
Two vectors $\mathbf{u}$ and $\mathbf{v}$ are said to be **orthogonal** if

$$
\langle \mathbf{u}, \mathbf{v} \rangle = 0
$$

# Orthogonal sets and bases

Let $\Theta$ be an inner product space.
A subset $S = {\mathbf{u}_1, \mathbf{u}_1, \dots, \mathbf{u}_k}$
of nonzero vectors of $\Theta$ is called an **orthogonal set**
if every pair of vectors are orthogonal, i.e.,

$$
\langle \mathbf{u}_i, \mathbf{u}_j \rangle = 0, \, 1 \le i < j \le k.
$$

An orthogonal set $S = {\mathbf{u}_1, \mathbf{u}_1, \dots, \mathbf{u}_k}$
is called an **orthonormal set** if we further have

$$
\lVert \mathbf{u}_i \rVert = 1, \, 1 \le i \le k.
$$

An **orthonormal basis** or $\Theta$ is a basis which is also
an orthonormal set.

# Orthogonal projection

Let $\Theta$ be an inner product space.
Let $\mathbf{\theta}_1$ be a nonzero vector of $\Theta$.
We want to decompose an arbitrary vector $\mathbf{y}$ into the form

$$
\mathbf{y} = \alpha \mathbf{\theta}_1 + z, \,
\mathrm{where} \, \mathbf{z} \in \mathbf{\theta}_1^\perp.
$$

Since $z \perp \mathbf{\theta}_1$, we have

$$
\langle \mathbf{\theta}_1, \mathbf{y} \rangle =
  \langle \mathbf{\theta}_1, \alpha \mathbf{\theta}_1 \rangle =
  \alpha \langle \mathbf{\theta}_1, \mathbf{\theta}_1 \rangle.
$$

This implies that

$$
\alpha = \frac{
    \langle \mathbf{\theta}_1, \mathbf{y} \rangle
  }
  {
    \langle \mathbf{\theta}_1, \mathbf{\theta}_1 \rangle
  }.
$$

We define the vector

$$
\mathrm{Proj}_{\mathbf{\theta}_1}(\mathbf{y}) = \frac{
    \langle \mathbf{\theta}_1, \mathbf{y} \rangle
  }
  {
    \langle \mathbf{\theta}_1, \mathbf{\theta}_1 \rangle
  } \mathbf{\theta}_1,
$$

called the **orthogonal projection of $\mathbf{y}$ along $\mathbf{\theta}_1$**.
The linear transformation
$\mathrm{Proj}_{\mathbf{\theta}_1}: \Theta \rightarrow \Theta$
is called the
**orthogonal projection of $\Theta$ onto the direction $\mathbf{\theta}_1$**.

(Example 8.1 clear, and then generalization of the orthogonal projection from
a vector to a subspace. But note this is in an Euclidean space, i.e.,
a vector space with *standard inner product*.)

```{r, include=FALSE}
alpha <- pi / 3
rho   <- cos(alpha)

u_1 <- c(1, rho)
u_2 <- c(0, sin(alpha))

R     <- rbind(u_1, u_2) # R^T = [u_1, u_2]
R_inv <- R |> solve() |> unname()
R_inv_alt <- matrix(c(1, 0, -rho / sin(alpha), 1 / sin(alpha)), nrow = 2)
all.equal(R_inv, R_inv_alt) # Son iguales

Sigma <- t(R) %*% R

Sigma_inv <- solve(Sigma)

all.equal(Sigma_inv, R_inv %*% t(R_inv)) # Son iguales
```

# Adaptation to the case in hand {-}

Let's assume that the inner product of $\mathbb{R}^2$ relative to basis
$\mathcal{B} = \{ \mathbf{u}_1, \mathbf{u}_2 \}$,
where
$\mathbf{u}_1 = \begin{bmatrix} 1\\ -\frac{\cos \alpha}{\sin \alpha} \\ \end{bmatrix}$
and
$\mathbf{u}_2 = \begin{bmatrix} 0\\ \frac{1}{\sin \alpha} \\ \end{bmatrix}$,
is defined by

$$
\langle \mathbf{x}, \mathbf{y} \rangle = \frac{4}{3} x_1y_1 - \frac{2}{3} x_1y_2
  - \frac{2}{3} x_2y_1 + \frac{4}{3} x_2y_2,
$$

where
$\mathbf{x} = \begin{bmatrix} x_1\\ x_2\\ \end{bmatrix}$,
$\mathbf{y} = \begin{bmatrix} y_1\\ y_2\\ \end{bmatrix}$,
$\in \mathbb{R}^2$,

$$
\Sigma^{-1} = \begin{bmatrix}
    \langle \mathbf{e}_1, \mathbf{e}_1 \rangle &
      \langle \mathbf{e}_1, \mathbf{e}_2 \rangle \\
    \langle \mathbf{e}_2, \mathbf{e}_1 \rangle &
      \langle \mathbf{e}_2, \mathbf{e}_2 \rangle \\
  \end{bmatrix} =
  \begin{bmatrix}
     \tfrac{4}{3} & -\tfrac{2}{3} \\
    -\tfrac{2}{3} &  \tfrac{4}{3} \\
  \end{bmatrix}
$$