### 2022-05-13 output/Derivations_by_Reckase/Derivation_according_to_Reckase.Rmd

---
title: "Multidimensional difficulty and discrimination parameters according to Reckase & McKinley"
author: "Daniel Morillo"
date: "`r Sys.Date()`"
output:
  officedown::rdocx_document:
    tables:
      style: APA_Like_style
    mapstyles:
      Normal: First Paragraph
    base_format: bookdown::word_document2
    reference_docx: ../../www/Template.docx
    number_sections: no
    keep_md:         no
    fig_width:       4
    fig_height:      4
bibliography: ../../www/Multidimensional-parameters-MCLM.bib
csl:          ../../www/apa-old-doi-prefix.csl
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)

ROOT_DIR <- "../.."
DOC_DIR <- getwd()

opts_knit$set(root.dir = ROOT_DIR)
opts_chunk$set(
  echo     = FALSE,
  cache    = FALSE,
  dev      = "png",
  dpi      =   300,
  dev.args = list(type = "cairo-png"),
  fig.pos = "!H"
)

# Constants:
GRAPH_FONT   <- "serif"
LINE_WIDTH   <- .1
VECTOR_WIDTH <- 1.5
ANGLE_WIDTH  <- .5
POINT_CHAR   <- 19L

# Output configuration options:
options(digits = 4)
# theme_set(
#   theme_classic(
#     base_size      = 12,
#     base_family    = GRAPH_FONT,
#     base_line_size = LINE_WIDTH
#   ) %+replace%
#   theme(
#     axis.title.x = element_text(hjust = 1),
#     axis.title.y = element_text(vjust = 1, angle = 0)
#   )
# )
```

```{r includes, cache=FALSE}
source("R/Formulas.R")
```

# Introduction

We are going to use the method proposed by @reckase_difficulty_1985
in order to determine the difficulty of a multidimensional item
in a non-orthogonal latent space.
Then, we will determine its discrimination,
following @reckase_discriminating_1991.
In order to better understand these,
we will first review the original derivations in the orthogonal space.

# Original parameter derivations

## Multidimensional difficulty in @reckase_difficulty_1985

Equation 7 in @reckase_difficulty_1985 applies the constraint
$\sum_{k=1}^n\cos^2 \alpha_{jk} = 1$,
which implies orthogonal axes.
This constraint is necessary because the $n$ cosines are not independent;
rather, they are (non-linearly) dependent by the constraint in their Equation 7.

For a certain dimension $k$ ($k \in [1, n-1]$),
the derivative of each addend in their Equation 6 is 0
(because it is independent of $\cos \alpha_{ki}$),
except for the derivatives of the $k-\mathrm{th}$ term,
which is $\tfrac{\alpha_{ki}}{4}$,
and the $n-\mathrm{th}$ term, given by

\begin{equation}
  \begin{split}
    \frac{\delta}{\delta \cos \alpha_{jk}}
      \frac{a_{in}}{4}
      \sqrt{ 1 - \sum_{k=1}^{n-1} \cos^2 \alpha_{jk} } &=
        \frac{a_{in}}{4}
          \frac{1}{2 \sqrt{ 1 - \sum_{k=1}^{n-1} \cos^2 \alpha_{jk} } }
          \left(-2 \cos \alpha_{jk} \right) \\
      &= - \frac{a_{in}}{4}
        \frac{
            \cos \alpha_{jk}
          } {
            \sqrt{ 1 - \sum_{k=1}^{n-1} \cos^2 \alpha_{jk} }
          } \\
      &= - \frac{a_{in}}{4} \frac{\cos \alpha_{jk}}{\cos \alpha_{jn}}.
  \end{split}
\end{equation}

Therefore, the derivative of their Equation 6 is given by

\begin{equation}
  \frac{\delta}{\delta \cos \alpha_{jk}}
    \frac{1}{4}
    \sum_{k=1}^{n} a_{ik} \cos \alpha_{jk} =
  \frac{1}{4}
    \left( a_{ik} - a_{in} \frac{\cos \alpha_{jk}}{\cos \alpha_{jn}} \right).
\end{equation}

The term $\tfrac{1}{4}$ can be dropped when equating to 0,
thus resulting in the $n-1$ expressions in Equation 8 in
@reckase_difficulty_1985.
By reformulating each of these equations we can show that
$\cos \alpha_{ik}$ is proportional to $a_{ik}$ for all $k$:

\begin{equation}
  a_{ik} - a_{in} \frac{\cos \alpha_{jk}}{\cos \alpha_{jn}} = 0 \Leftrightarrow
    \frac{a_{ik}}{\cos \alpha_{jk}} = \frac{a_{in}}{\cos \alpha_{jn}} = C.
\end{equation}

Thus, the solution implies computing the proportionality constant $C$.

The system is complemented by Equation 7,
giving $n$ equations with the $n$ cosines as unknowns.
Operating on the previous equation,

\begin{equation}
  \begin{split}
    \cos \alpha_{jk} = \frac{a_{ik}}{C} &\Leftrightarrow
      \cos^2 \alpha_{jk} = \frac{a_{ik}^2}{C^2} \\
      &\Leftrightarrow \sum_{k=1}^{n} \cos^2 \alpha_{jk} =
        \sum_{k=1}^{n} \frac{a_{ik}^2}{C^2} \\
      &\Leftrightarrow \frac{1}{C^2} \sum_{k=1}^{n} a_{ik}^2 = 1 \\
      &\Leftrightarrow C^2 = \sum_{k=1}^{n} a_{ik}^2 \\
      &\Leftrightarrow C = \sqrt{ \sum_{k=1}^{n} a_{ik}^2 },
  \end{split}
\end{equation}

results in the value of $C$ which, when solving for $j = i$,
gives Equation 9 in @reckase_difficulty_1985.

The signed distance to the point of steepest slope, $D_i$,
is obtained by substituting Equation 9 for $\cos \alpha_{ik}$ in Equation 3.
The exponent of the logit function results in
(the summation index is changed from $k$ to $l$ in the denominator
to avoid collision):

\begin{equation}
  \begin{split}
    \sum_{k=1}^{n} a_{ik} `r THETA_MODULE` \cos \alpha_{jk} + d_i &=
        \sum_{k=1}^{n} a_{ik}
        `r THETA_MODULE`
        \frac{a_{ik}}{\sqrt{ \sum_{l=1}^{n} a_{il}^2} } + d_i \\
      &= \frac{ \sum_{k=1}^{n} a_{ik}^2 }{ \sqrt{ \sum_{l=1}^{n} a_{il}^2 } }
        `r THETA_MODULE`  + d_i \\
      &= \sqrt{
          \frac{ \left( \sum_{k=1}^{n} a_{ik}^2 \right)^2 }
               { \sum_{l=1}^{n} a_{il}^2 }
        }
        `r THETA_MODULE` +
        d_i \\
      &= \sqrt{ \sum_{k=1}^{n} a_{ik}^2 } `r THETA_MODULE`  + d_i.
  \end{split}
\end{equation}

Solving for $P_{ij} = .5$ implies that the exponent equals 0.
Thus solving for $D_i = `r THETA_MODULE`$,

\begin{equation}
  \begin{split}
    \sqrt{ \sum_{k=1}^{n} a_{ik}^2 } D_i + d_i = 0 &\Leftrightarrow
      \sqrt{ \sum_{k=1}^{n} a_{ik}^2 } D_i = -d_i \\
    &\Leftrightarrow D_i = - \frac{d_i}{ \sqrt{ \sum_{k=1}^{n} a_{ik}^2 } },
  \end{split}
\end{equation}

which gives Equation 10.

## Multidimensional discrimination

The multidimensional discrimination parameter ($MDISC$)
is obtained straightforwardly from the previous equations and
Equations 6 through 9 in @reckase_discriminating_1991.

# Parameters in the non-orthogonal space

## Model definition

The model is defined by Reckase's [-@reckase_difficulty_1985] Equation 1.
However, in the present case, $`r THETA_J`$ is multivariate normal:

\begin{equation}
  `r THETA_J` \sim \mathcal{N}(\mathbf{\upmu}, `r SIGMA`),
\end{equation}

where $\mathbf{\upmu}$ is an $n$-element vector of means and
$`r SIGMA` = `r S_MATRIX` `r R_MATRIX` `r S_MATRIX`'$
an $n$-dimensional positive definite covariance matrix,
with $`r R_MATRIX`$ the corresponding correlation matrix
(also positive definite) and $`r S_MATRIX`$ a scaling diagonal matrix
with $s_{ii} = \sigma_{ii}$
(i.e., the matrix of standard deviations).

When $`r THETA_J`$ is standardized, $`r S_MATRIX` = `r ID_MATRIX`$ and
$`r SIGMA` = `r R_MATRIX`$.
When $`r R_MATRIX` = `r ID_MATRIX`$, the elements of $`r THETA_J`$
are independent and $`r SIGMA` = `r S_MATRIX` `r S_MATRIX`'$,
which implies that the basis of the latent space $`r THETA`$ is orthogonal.
When both conditions occur simultaneously, $`r SIGMA` = `r ID_MATRIX`$,
and the basis of the latent space $`r THETA`$ is canonical and therefore
orthonormal.

One may assume that the latter must occur for the conditions in
@reckase_difficulty_1985 to apply.
However, Reckase's derivations apply to an orthogonal latent space,
which actually happens in the second condition.
That is, the only requirement is that the elements of $`r THETA_J`$ are
independent, being irrelevant whether they are standardized or not.
Therefore, our generalization applies to the non-orthogonal case.

## Model recasting into polar coordinates {#polar-coords}

Following the method outlined by @reckase_difficulty_1985,
the model is first transformed into polar coordinates,
as in Reckase's [-@reckase_difficulty_1985] Equation 3.
However, to apply this transformation one must consider that,
in a non-orthogonal space,
$`r THETA_J` = `r R_COS_ALPHA_THETA`$
instead of $`r THETA_MODULE` `r COS_ALPHA_J_VECTOR`$, as in Rekcase's
[-@reckase_difficulty_1985] Equation 3.

**Proof 1**.[^proof-needed]
First, let $`r THETA_ORTH` = `r B_INVERSE` `r THETA`$
be an orthogonal latent space with $`r B_MATRIX`$ a rotation matrix such that
$`r B_MATRIX` `r B_MATRIX`' = `r R_MATRIX`$.
The norm $`r THETA_J_ORTH_NORM` \equiv `r THETA_MODULE`$
of a vector $`r THETA_J_ORTH`$ in $`r THETA_ORTH`$
is defined by

\begin{equation}
  `r THETA_MODULE`^2 = `r THETA_J_ORTH`' `r THETA_J_ORTH` =
    `r THETA_J`' `r B_INVERSE`' `r B_INVERSE` `r THETA_J` =
    `r THETA_J`' (`r B_MATRIX` `r B_MATRIX`')^{-1} `r THETA_J` =
    `r THETA_J`' `r R_INVERSE` `r THETA_J`.                 (\#eq:inner-product)
\end{equation}

Therefore, we consistently define
$\langle \mathbf{x}, \mathbf{y} \rangle = \mathbf{x}' `r R_INVERSE` \mathbf{y}$ as an inner product in $`r THETA`$ with norm
$`r THETA_MODULE` = \sqrt{`r THETA_J`' `r R_INVERSE` `r THETA_J`}$.
By definition of the inner product, the cosine between
$`r THETA_J`$ and each vector $`r BASIS_VECTOR`$ of the basis of $`r THETA`$
is given by their inner product divided by the norms of the two vectors:

\begin{equation}
  `r COS_ALPHA_JK` = \frac{
      `r BASIS_VECTOR`' `r R_INVERSE` `r THETA_J`
    } {
      `r BASIS_VECTOR_NORM` `r THETA_J_NORM`
    } =
    \frac{
      `r BASIS_VECTOR`' `r R_INVERSE` `r THETA_J`
    } {
      1 `r THETA_J_NORM`
    } =
    \frac{`r BASIS_VECTOR`' `r R_INVERSE` `r THETA_J`}{ `r THETA_MODULE`}.
\end{equation}

Therefore, the vector of cosines is given by

\begin{equation}
  `r COS_ALPHA_J_VECTOR` = \frac{
      \begin{pmatrix}
        `r BASIS_VECTOR_1`' \\
        \vdots              \\
        `r BASIS_VECTOR_N`'
      \end{pmatrix}
      `r R_INVERSE` `r THETA_J`
    } {
       `r THETA_MODULE`
    } =
    \frac{`r ID_MATRIX` `r R_INVERSE` `r THETA_J`}{`r THETA_MODULE`}
\end{equation}

\begin{equation}
    `r COS_ALPHA_J_VECTOR` = \frac{`r R_INVERSE` `r THETA_J`}{`r THETA_MODULE`}.
                                                         (\#eq:director-cosines)
\end{equation}

Solving for $`r THETA_J`$,

\begin{equation}
  \begin{split}
    \frac{`r R_MATRIX` `r R_INVERSE` `r THETA_J`}{ `r THETA_MODULE`} &=
      `r R_COS_ALPHA` \\
    `r THETA_J` &= `r R_COS_ALPHA_THETA`.\blacksquare
  \end{split}
\end{equation}

[^proof-needed]: It took me quite an effort to realise that Equation 3 in
@reckase_difficulty_1985 did not apply straightforwardly,
and derivate the following proof.
However, I'm not sure if it's really necessary to proof this,
as I think it must be an already-known result of the inner product
[see e.g., @harman_modern_1970,
though probably not the most appropriate source].

Thus, the model expression in polar coordinates is given by

\begin{equation}
  `r P_MCLM_OBL` = `r LOGISTIC_OBL_POLAR`.                     (\#eq:MCLM-polar)
\end{equation}

## Multidimensional difficulty {#md-diff}

The equivalent to Reckase's [-@reckase_difficulty_1985]
Equations 4 through 6 are:

\begin{equation}
  `r IRF_DIFF2_THETA` = `r IRF_DIFF2_THETA_DEF`,
\end{equation}

\begin{equation}
  `r IRF_DIFF_THETA` = `r IRF_DIFF_THETA_DEF`,
\end{equation}

and

\begin{equation}
  \tfrac{1}{4} `r A_R_COS_ALPHA`,                               (\#eq:max-slope)
\end{equation}

To derive the direction of steepest slope we must take into account the
dependence among the components of $`r COS_ALPHA_J_VECTOR`$,
as in Reckase's [-@reckase_difficulty_1985] Equation 7.
However, this dependence is expressed here as

\begin{equation}
  `r COS_ALPHA_J_VECTOR` ' `r R_MATRIX` `r COS_ALPHA_J_VECTOR` = 1
                                                        (\#eq:cosine-constraint)
\end{equation}

**Proof 2**.[^proof-needed-2]
From Equation \@ref(eq:director-cosines), we have that

\begin{equation}
  \begin{split}
    `r R_MATRIX` `r COS_ALPHA_J_VECTOR` &=
      \frac{`r THETA_J`}{`r THETA_MODULE`},\\
    `r B_MATRIX` `r B_MATRIX`' `r COS_ALPHA_J_VECTOR` &=
      \frac{`r THETA_J`}{`r THETA_MODULE`}.
  \end{split}
\end{equation}

Also, in the orthogonal space,

\begin{equation}
  \begin{split}
    `r COS_ALPHA_J_ORTH_VECTOR` &= \frac{`r THETA_J_ORTH`}{`r THETA_MODULE`}, \\
    `r COS_ALPHA_J_ORTH_VECTOR` &=
      \frac{`r B_INVERSE` `r THETA_J`}{`r THETA_MODULE`}, \\
    `r B_MATRIX` `r COS_ALPHA_J_ORTH_VECTOR` &=
      \frac{`r THETA_J`}{`r THETA_MODULE`}.
  \end{split}
\end{equation}

Equating the left terms in both equations,

\begin{equation}
  `r B_MATRIX` `r B_MATRIX`' `r COS_ALPHA_J_VECTOR` =
    `r B_MATRIX` `r COS_ALPHA_J_ORTH_VECTOR`,
\end{equation}

\begin{equation}
  `r B_MATRIX`' `r COS_ALPHA_J_VECTOR` = `r COS_ALPHA_J_ORTH_VECTOR`
                                                      (\#eq:cosines-equivalence)
\end{equation}

and squaring,

\begin{equation}
  \begin{split}
    `r COS_ALPHA_J_VECTOR`' `r B_MATRIX` `r B_MATRIX`' `r COS_ALPHA_J_VECTOR` &=
      `r COS_ALPHA_J_ORTH_VECTOR`' `r COS_ALPHA_J_ORTH_VECTOR`; \\
    `r COS_ALPHA_J_VECTOR`' `r R_MATRIX` `r COS_ALPHA_J_VECTOR` &= 1.
      \blacksquare
  \end{split}
\end{equation}

[^proof-needed-2]: As before, I doubt this proof is strictly necessary.

(Now, I've come to a dead end here, as I don't know how to solve
\@ref(eq:cosine-constraint) for $`r COS_2_ALPHA_JN`$.
I arrive to the covariance matrix of $`r COS_ALPHA_J_VECTOR`$,
$\mathrm{cov}(`r COS_ALPHA_J_VECTOR`) = `r R_INVERSE` `r S_MATRIX` `r R_MATRIX` `r S_MATRIX`' `r R_INVERSE`'$,
but I don't see how this can be solved for a single component of
$`r COS_ALPHA_J_VECTOR`$.
A possible solution would be to define
$`r THETA_ORTH` = `r T_INVERSE` `r THETA`$ as an orthonormal latent space
with $`r T_MATRIX` = `r S_MATRIX` `r B_MATRIX`$,
such that $`r T_MATRIX` `r T_MATRIX`' = `r SIGMA`$.
This would lead to the substitution of $`r R_MATRIX`$ by $`r SIGMA`$ in all
the equations in sections \@ref(polar-coords) and \@ref(md-diff),
and therefore $\mathrm{cov}(`r COS_ALPHA_J_VECTOR`) = `r SIGMA`^{-1}$.
However, I don't know how to continue from here either.)

### Derivating the multidimensional difficulty by orthogonalization

The most straightforward way of solving it seems to be by orthogonalizing
the latent space, computing the solution in this orthogonalized space,
and then transforming the space back into oblique coordinates
(as I do in my dissertation).
Using this approach, we may find the solution in the orthogonal space
applying the method in @reckase_difficulty_1985.

First, we define $`r A_VECTOR_ORTH` = `r B_MATRIX`' `r A_VECTOR`$ such that
$`r A_VECTOR_ORTH`' `r THETA_J_ORTH` = `r A_VECTOR`' `r B_MATRIX` `r B_INVERSE` `r THETA_J` = `r A_VECTOR`' `r THETA_J`$.
Then, we can compute $`r COS_ALPHA_J_ORTH_VECTOR`$ and $`r MDIFF`$ applyting
Reckase's [-@reckase_difficulty_1985] equations 3 through 10.
This results in

\begin{equation}
  `r COS_ALPHA_IK_ORTH` =
    \frac{`r A_COMPONENT_ORTH`}{\sqrt{`r A_VECTOR_ORTH`' `r A_VECTOR_ORTH`}}
\end{equation}

and

\begin{equation}
  `r MDIFF` =
    \frac{- `r D_PARAM`}{\sqrt{`r A_VECTOR_ORTH`' `r A_VECTOR_ORTH`}}.
\end{equation}

The first one can be expressed in vector terms:

\begin{equation}
  `r COS_ALPHA_I_ORTH_VECTOR` =
    \frac{`r A_VECTOR_ORTH`}{\sqrt{`r A_VECTOR_ORTH`' `r A_VECTOR_ORTH`}},
\end{equation}

and thus both can be expressed in terms of $`r A_VECTOR`$ instead of
$`r A_VECTOR_ORTH`$, with
$`r A_VECTOR_ORTH`' `r A_VECTOR_ORTH` = `r A_VECTOR`' `r B_MATRIX` `r B_MATRIX`' `r A_VECTOR` = `r A_VECTOR`' `r R_MATRIX` `r A_VECTOR`$:

\begin{equation}
  `r COS_ALPHA_I_ORTH_VECTOR` = \frac{
      `r B_MATRIX`' `r A_VECTOR`
    } {
      \sqrt{`r A_VECTOR`' `r R_MATRIX` `r A_VECTOR`}
    },
\end{equation}

and

\begin{equation}
  `r MDIFF` =
    \frac{- `r D_PARAM`}{\sqrt{`r A_VECTOR`' `r R_MATRIX` `r A_VECTOR`}}.
\end{equation}

From Equation \@ref(eq:cosines-equivalence) we have that 
$`r COS_ALPHA_I_ORTH_VECTOR` = `r B_MATRIX`' `r COS_ALPHA_I_VECTOR`$, so
[^warning-direction]

\begin{equation}
  `r B_MATRIX`' `r COS_ALPHA_I_VECTOR` = \frac{
      `r B_MATRIX`' `r A_VECTOR`
    } {
      \sqrt{`r A_VECTOR`' `r R_MATRIX` `r A_VECTOR`}
    },
\end{equation}

\begin{equation}
  `r COS_ALPHA_I_VECTOR` =
    \frac{`r A_VECTOR`}{\sqrt{`r A_VECTOR`' `r R_MATRIX` `r A_VECTOR`}}.
                                                         (\#eq:director-cosines)
\end{equation}

[^warning-direction]: Warning!
This result is not the same as what I found in the dissertation (Equation 3.7).
This one seems to be the correct one though.
I think the explanation is that, in my dissertation,
the projections should be on the orthogonal axes instead, that is,
$`r COS_ALPHA_I_ORTH_VECTOR`'$ in the left side.

### Derivating the multidimensional difficulty form the equation in polar coordinates

Using the method outlined above, one can derivate the director cosines which
describe the measurement direction.
However, the absolute value of the multidimensional difficulty can be derived
by applying this solution to the model expression in polar coordinates,
i.e. Equation \@ref(eq:MCLM-polar).

Substituting Equation \@ref(eq:director-cosines) for $`r COS_ALPHA_I_VECTOR`$ in
Equation \@ref(eq:MCLM-polar) we may find the maximum slope, solving for
$`r MDIFF` = `r THETA_MODULE`$ when $P_ij = .5$,
that is, $`r P_MCLM_OBL_MDIFF` = .5$:

\begin{equation}
  \begin{split}
    `r LOGISTIC_OBL_POLAR_MDIFF` = .5 &\Leftrightarrow
      `r EXPONENT_OBL_POLAR_MDIFF` = 0 \\
      &\Leftrightarrow \frac{
        `r A_VECTOR`' `r R_MATRIX` `r A_VECTOR`
      } {
        \sqrt{`r A_VECTOR`' `r R_MATRIX` `r A_VECTOR`}
      } `r MDIFF` = - `r D_PARAM`
  \end{split}
\end{equation}

\begin{equation}
  \begin{split}
      `r MDIFF` = - \frac{
        `r D_PARAM`
      } {
        \sqrt{`r A_VECTOR`' `r R_MATRIX` `r A_VECTOR`}
      }
  \end{split}                                              (\#eq:MDIFF-solution)
\end{equation}

## Multidimensional discrimination

Analogously, we apply Equations 4 through 9 in @reckase_discriminating_1991,
arriving at the equivalent of their Equation 9 in the orthogonal space:

\begin{equation}
  `r MDISC` = \sqrt{`r A_VECTOR_ORTH`' `r A_VECTOR_ORTH`} =
    \sqrt{`r A_VECTOR`' `r R_MATRIX` `r A_VECTOR`}.
\end{equation}

## Properties of the multidimensional parameters

### Multidimensional discrimination

Substituting $`r COS_ALPHA_J_VECTOR`$ in expression \@ref(eq:max-slope) by
$`r COS_ALPHA_I_VECTOR`$,
and the substituting Equation \@ref(eq:director-cosines) into
\@ref(eq:max-slope) results in

\begin{equation}
  \tfrac{1}{4} \sqrt{`r A_VECTOR`' `r R_MATRIX` `r A_VECTOR`} =
    \tfrac{1}{4} `r MDISC`.
\end{equation}

Therefore, $`r MDISC`$ is analogous
to the $a_i$ parameter in unidimensional IRT:
Its scale amounts to four times the slope and can be interpreted accordingly,
as in the orthogonal case.

When $a_{il} \neq 0$ and $a_{ik} = 0, \forall k \neq l$,
then $`r MDISC` = a_{il}$.
That is, $`r MDISC`$ reduces to the unidimensional discrimination parameter
as expected, the same way it does in the orthogonal case.

Finally, @reckase_difficulty_1985 [p. 404] states that "when $`r ALPHA_JK` = 0$
and $\alpha_{jl} = 90$ for $l \neq k$,
indicating that the slope of the surface at $P_{ij} = .5$
is determined parallel to axis $k$, $`r COS_ALPHA_JK` = 1$ and
$\cos \alpha_{jl} = 0$.
The slope parallel to axis $k$ is then given by $\tfrac{1}{4} a_{ik}$."

In the non-orthogonal space, the equality $\alpha_{jl} = 90$
does not necessarily hold,
and thus it is not generally true that $\cos \alpha_{jl} = 0$.
In fact, $`r COS_ALPHA_JK`$ equals the inner product of $`r A_VECTOR`$
with the basis vector
$`r BASIS_VECTOR`' = (e_1, \ldots, e_k, \ldots, e_n), e_k = 1, e_l = 0 \, \forall l \neq k$:

\begin{equation}
  `r COS_ALPHA_JK` = \frac{
      `r BASIS_VECTOR`' `r R_MATRIX` `r A_VECTOR`
    } {
      \left\| e \right\| \left\| `r A_VECTOR` \right\|
    } =
    \frac{
      `r BASIS_VECTOR`' `r R_MATRIX` `r A_VECTOR`
    } {
      \left\| e \right\| \left\| `r A_VECTOR` \right\|
    } =
    \frac{
      (\rho_{k1}, \ldots, 1, \ldots, \rho_{kn}) `r A_VECTOR`
    } {
      \sqrt{`r A_VECTOR`' `r R_MATRIX` `r A_VECTOR`}
    } =
    \frac{
      a_{ik} + \sum_{l = 1, l \neq k}^n \rho_{kl} a_{il}
    } {
      \sqrt{`r A_VECTOR`' `r R_MATRIX` `r A_VECTOR`}
    }.
\end{equation}

Therefore, this condition does not generally hold in the non-orthogonal space.

### Multidimensional difficulty

$`r MDIFF`$ has the same relationship with $`r D_PARAM`$
as $b_i$ for the 2PL model, i.e.

\begin{equation}
  `r MDIFF` = - \frac{`r D_PARAM`}{`r MDISC`}.       (\#eq:relationship-MDIFF-d)
\end{equation}

# Alternative definition of the inner product

## Orthogonal interpretation of the coordinates for correlated traits

The definition of the inner product in $`r THETA`$ as
$\langle \mathbf{x}, \mathbf{y} \rangle = \mathbf{x}' `r R_INVERSE` \mathbf{y}$
is arbitrary, and it has been done in order to make it consistent with the
inner product
$\langle \mathbf{x}, \mathbf{y} \rangle = \mathbf{x}' \mathbf{y}$
in the orthogonal space $`r THETA_ORTH`$.
This is convenient in order to transform between the oblique and orthogonal
space back and forth,
and thus leverage on Reckase's [-@reckase_difficulty_1985] derivations to
solve the non-orthogonal case.
However, we could want to define an inner product in $`r THETA`$ such that
the inner product in the test space
$\langle \mathbf{x}, \mathbf{y} \rangle = \mathbf{x}' \mathbf{y}$.
This would imply an inner product
$\langle \mathbf{x}, \mathbf{y} \rangle = \mathbf{x}' \mathbf{y}$
in the latent trait space, and all the formulae in @reckase_difficulty_1985
and @reckase_discriminating_1991 would apply.
However, in applications where one would need to compute the equivalent
parameters in an orthogonal space (i.e. equating, linking, or multigroup IRT),
additional complications would arise,
as the inner product in the orthogonal space would be defined by
the inner product
$\langle \mathbf{x}, \mathbf{y} \rangle = \mathbf{x}' (`r B_MATRIX`' `r B_MATRIX`)^{-1} \mathbf{y}$.

## Orthonormalization

The orthogonal latent space $`r THETA_ORTH`$ could instead be defined as
$`r THETA_ORTH` = `r T_INVERSE` `r THETA`$,
with $`r T_MATRIX` = `r S_MATRIX` `r B_MATRIX`$,
such that $\mathrm{cov}(`r THETA_ORTH`) = `r ID_MATRIX`$.

**Proof 3.** The covariance of $`r THETA_ORTH`$ is given by
$\mathrm{cov}(`r THETA_ORTH`) = `r T_INVERSE` `r SIGMA` `r T_INVERSE`'$, and
$`r T_MATRIX` `r T_MATRIX`' = `r S_MATRIX` `r B_MATRIX` `r B_MATRIX`' `r S_MATRIX`' = `r SIGMA`$.
Thus substituting:

\begin{equation}
  \mathrm{cov}(`r THETA_ORTH`) =
    `r T_INVERSE` `r T_MATRIX` `r T_MATRIX`' `r T_INVERSE`' =
    `r ID_MATRIX`.\blacksquare
\end{equation}

Using this definition results in an inner product as defined in
@zhang_theoretical_1999; all the derivations from \@ref(eq:inner-product) to
\@ref(eq:relationship-MDIFF-d) would be equivalent,
substituting $`r B_MATRIX`$ by $`r T_MATRIX`$ and $`r R_MATRIX`$ by $`r SIGMA`$.
This has the advantage that different latent spaces can be referred to a common
metric.
However, it also has the disadvantage of changing the metric (i.e. covariances)
of the latent space, resulting in a change in variances.
This is detrimental for graphical representation purposes, on one side,
and for the property of analogy with unidimensional IRT on the other,
as $`r MDISC`$ is no longer analogous to its unidimensional counterpart $a_i$.

The original derivation, on the other side,
has a consistent metric with the original, oblique latent space
(but $`r THETA_ORTH`$ apparently does not satisfy
$\mathrm{cov}(`r THETA_ORTH`) = `r S_MATRIX` `r S_MATRIX`'$, in general).

# References