### 2022-07-27 notebooks/Space_transformation_theorem.qmd
Cosine director lemmas, initial version [WIP]

---
title: "Space transformation in Multidimensional IRT (Theorem)"
author: "Mario Luzardo, Daniel Morillo"
format:
  docx:
    number-sections: true
    highlight-style: github  
  html: default
callout-collapse: true
editor_options: 
  chunk_output_type: console
---

# Introduction

Transcription of Mario's notes

```{r setup, include=FALSE}
# Includes:

library(knitr)

ROOT_DIR <- ".."
DOC_DIR <- getwd()

opts_knit$set(root.dir = ROOT_DIR)
# opts_chunk$set(
#   echo     = FALSE,
#   cache    = FALSE,
#   dev      = "png",
#   dpi      =   300,
#   dev.args = list(type = "cairo-png"),
#   fig.pos = "!H"
# )
```

```{r sources, cache=FALSE, include=FALSE}
source("R/Formulae.R", encoding = 'UTF-8')
```

# Main

Let's consider a $`r N_DIMS`$ latent space $`r LATENT_SPACE`$ with Euclidean
vector structure, and let $`r LS_BASIS_EQ`$ be a basis of $`r LATENT_SPACE`$.

::: callout-note
Stating that $`r LS_BASIS`$ is a basis implies that its vectors are linearly
independent and are a generator of $`r LATENT_SPACE`$; i.e., any vector
$`r TRAIT_VEC_IN_LS`$ can be expressed as a linear combination of
$`r LS_BASIS`$.
:::

Let's consider a dichotomous item $`r ITEM_INDEX`$ such that the probability
that respondent with trait $`r LS_BASIS_EQ`$ gives a correct[^1] response is
given by

$$
  `r M2PL_FORMULATION`
$$ {#eq-M2PL-formulation}

Let's consider in $`r LATENT_SPACE`$ the standard basis $`r LS_STD_BASIS_EQ`$ as
well.

For all $`r TRAIT_VEC_IN_LS`$ we have both representations[^2]:

$`r TRAIT_VECTOR_DEF_BASIS_EQ`$, and $`r TRAIT_VECTOR_DEF_STD_BASIS_EQ`$

Let $`r BASIS_CHANGE_DEF`$ be the change of basis matrix.

[^1]: I prefer to refer to the responses as "positive/negative", instead of
    "correct/incorrect", as this terminology contemplates the possibility of
    non-cognitive as well as cognitive domains.

[^2]: I'm preferring here the notation I use in the draft, that is,
    $`r DIM_INDEX`$ for indexing the dimensions (instead of $`r ITEM_INDEX`$,
    which is also used to index the items), $`r TRAIT_MODULE`$ for the
    coordinates of $`r TRAIT_VECTOR`$ in $`r LS_BASIS`$,
    and $`r TRAIT_VEC_STD_EL`$ for the coordinates in $`r LS_STD_BASIS`$.
    Nevertheless, as Mario reminded me, it is worth noting that this notation
    for the vector coordinates is ambiguous, as we are using the same symbol
    ($`r TRAIT_VECTOR`$) for the "geometric vector" and
    the "algebraic vector" (i.e. column matrix) of its coordinates in
    $`r LATENT_SPACE`$ (this notation is usually accepted though).