### 2022-08-05 notebooks/Space_transformation_theorem.qmd
Direction cosines complete
Rebase merge conflict addressed. Adds footnote nº 9

---
title: "Space transformation in Multidimensional IRT (Lemmas)"
author: "Mario Luzardo, Daniel Morillo"
format:
  docx:
    number-sections: true
    highlight-style: github  
callout-collapse: true
editor_options: 
  chunk_output_type: console
---

# Introduction

Transcription of Mario's notes

```{r setup, include=FALSE}
# Includes:

library(knitr)

ROOT_DIR <- ".."
DOC_DIR <- getwd()

opts_knit$set(root.dir = ROOT_DIR)
# opts_chunk$set(
#   echo     = FALSE,
#   cache    = FALSE,
#   dev      = "png",
#   dpi      =   300,
#   dev.args = list(type = "cairo-png"),
#   fig.pos = "!H"
# )
```

```{r sources, cache=FALSE, include=FALSE}
source("R/Formulae.R", encoding = 'UTF-8')
```

# Main

Let's consider a $`r N_DIMS`$ latent space $`r LATENT_SPACE`$ with Euclidean
vector structure, and let $`r LS_BASIS_EQ`$ be a basis of $`r LATENT_SPACE`$.

::: callout-note
Stating that $`r LS_BASIS`$ is a basis implies that its vectors are linearly
independent and are a generator of $`r LATENT_SPACE`$; i.e., any vector
$`r TRAIT_VEC_IN_LS`$ can be expressed as a linear combination of
$`r LS_BASIS`$.
:::

Let's consider a dichotomous item $`r ITEM_INDEX`$ such that the probability
that respondent with trait $`r LS_BASIS_EQ`$ gives a positive[^1] response is
given by

$$
  `r M2PL_FORMULATION`
$$ {#eq-M2PL-formulation}

Let's consider in $`r LATENT_SPACE`$ the standard basis $`r LS_STD_BASIS_EQ`$ as
well.

For all $`r TRAIT_VEC_IN_LS`$ we have both representations[^2]:

$`r TRAIT_VECTOR_DEF_BASIS_EQ`$, and $`r TRAIT_VECTOR_DEF_STD_BASIS_EQ`$

Let $`r BASIS_CHANGE_DEF`$ be the change of basis matrix; then, if we denote
$`r TRAIT_COORDS_DEF`$ and $`r TRAIT_STD_COORDS_DEF`$ we have that

$$
  `r BASIS_CHANGE_EQ`
$$ {#eq-change-basis}

The $`r DIM_INDEX`$-th column of $`r TRANSFORM_MATRIX`$ is given by the
coordinates of $`r BASIS_VECTOR_ANY`$ in basis $`r LS_STD_BASIS`$. If
$`r BASIS_VEC_STD_COORDS_DEF`$ then[^3]

$$
  `r TRANSFORM_MATRIX_EQ`
$$ {#eq-transform-matrix-def}

In $`r LATENT_SPACE`$ with the standard basis we have the usual scalar product.
In $`r LATENT_SPACE`$ with basis $`r LS_BASIS`$ we define an inner product
$`r INNER_PROD_BASIS`$ such that the norm of $`r TRAIT_VECTOR`$ is invariant,
i.e., $`r TRAIT_NORMS_EQ`$. Thus,[^4]

$$
  `r TRAIT_NORM_STD_EQ`.
$$

Therefore we must define the inner product $`r INNER_PROD_BASIS`$ with Gram
matrix $`r INNER_PROD_MAT_EQ`$. That is, $`r INNER_PROD_EQ`$.[^5]

The Gram matrix has as element $`r INNER_PROD_MAT_EL_EQ`$. As
$`r TRANSFORM_MATRIX`$ (and consequently, $`r TRANSFORM_MATRIX_TRANSP`$) is
invertible so $`r INNER_PROD_MATRIX`$ is, and
$`r INNER_PROD_MATRIX_INV_EQ`$.[^6]

Let's consider now the vectors $`r DIR_COS_VEC`$ and $`r DIR_COS_VEC_STD`$, made
up by $`r N_DIMS`$ direction cosines of $`r TRAIT_VECTOR`$ with the basis
vectors of $`r LS_BASIS`$ and $`r LS_STD_BASIS`$, respectively.[^7]

::: {#lem-dircos}
Under the previously defined conditions, for all $`r TRAIT_VEC_IN_LS`$,[^8]

$$
  `r DIR_COS_VEC_EQ`
$$

with

$$
  `r DIAG_MATRIX_INNER_PROD_INV_SQ_EQ`.
$$

Therefore,

$$
  `r TRAIT_VECTOR_POLAR_EQ2`.
$$

::: proof
For every $`r DIM_INDEX`$; $`r DIM_ENUM_EQ`$,

$$
  `r DIR_COSINE_PROOF`,
$$

with $`r INNER_PROD_MATRIX_ROW`$ the $`r DIM_INDEX`$-th row of
$`r INNER_PROD_MATRIX`$.

Therefore, $`r DIR_COS_VEC_EQ2`$ and $`r TRAIT_VECTOR_POLAR_EQ2`$.[^9]
$\blacksquare$
:::
:::

::: {#lem-convert-spaces}
Let's consider another basis $`r NEW_BASIS_EQ`$ in $`r LATENT_SPACE`$ and the
inner product $`r INNER_PROD_NEWBS`$ that keeps the norm invariant.

Let $`r DIR_COS_VEC_NEWBS`$ be the vector of direction cosines of
$`r TRAIT_VECTOR`$ (i.e., the cosines with the elements of $`r NEW_BASIS`$).

Then,

$$
  `r LEMMA_2_EQ`,
$$ {#eq-lemma-2}

where

$$
  `r DIAG_MATRIX_INNER_PROD_NEWBS_INV_SQ_EQ`
$$

and $`r BASIS_CHANGE_NEWBS_DEF`$.

::: proof
Let's call $`r TRAIT_NEWBS_COORDS_DEF`$ and let $`r BASIS_CHANGE_NEWBS_DEF`$ be
the change of basis matrix such that $`r BASIS_CHANGE_NEWBS`$.

We have that $`r BASES_CHANGE_EQ`$.

The inner product $`r INNER_PROD_NEWBS`$ that keeps the norm invariant is
defined by[^10] $`r INNER_PROD_NEWBS_EQ`$, and we define
$`r INNER_PROD_MAT_NEWBS_EQ`$.

By @lem-dircos, for each case we have:

1.  $`r TRAIT_VECTOR_POLAR_STD_EQ`$

2.  $`r TRAIT_VECTOR_POLAR_EQ2`$

3.  $`r TRAIT_VEC_NEWBS_POLAR_EQ`$

Given (1),

$$
  `r TRAIT_COORDS_EQUIV`
$$

$$
  `r TRAIT_COORDS_EQUIV_NEWBS`.
$$

Using (2) and (3) we get[^11]

$$
  `r TRAIT_COORDS_EQUIV_EXPANDED`
$$

$$
  `r TRAIT_COORDS_EQUIV_NEWBS_EXPANDED`.
$$

Therefore,

$$
  `r TRAIT_COORDS_EQUIV_EXPANDED_EQ`,
$$

and thus $`r LEMMA_2_EQ`$.$\blacksquare$
:::
:::

[^1]: I prefer to refer to the responses as "positive/negative", instead of
    "correct/incorrect", as this terminology contemplates the possibility of
    non-cognitive (as well as cognitive) domains.

[^2]: I'm preferring here the notation I use in the draft, that is,
    $`r DIM_INDEX`$ for indexing the dimensions (instead of $`r ITEM_INDEX`$,
    which is also used to index the items), $`r TRAIT_MODULE`$ for the
    coordinates of $`r TRAIT_VECTOR`$ in $`r LS_BASIS`$, and
    $`r TRAIT_VEC_STD_COMP`$ for the coordinates in $`r LS_STD_BASIS`$.
    Nevertheless, as Mario reminded me, it is worth noting that this notation
    for the vector coordinates is ambiguous, as we are using the same symbol
    ($`r TRAIT_VECTOR`$) for the "geometric vector" and the "algebraic vector"
    (i.e. column matrix) of its coordinates in $`r LATENT_SPACE`$ (this notation
    is usually accepted though).

[^3]: Is it ok to use square brackets here? I'm using it instead of parentheses
    because I had already written this (with a function), but if it's not
    correct I change it to parentheses.

[^4]: Mario wrote in the leftmost therm of the equality "the norm of
    $`r TRAIT_VECTOR`$", but it should be "the *squared* norm of
    $`r TRAIT_VECTOR`$".

[^5]: Having already defined $`r TRAIT_COORDS_DEF`$, I have found this notation
    more convenient (consistent with what I had already written, for which I had
    it already coded).

[^6]: I haven't considered it necessary to add the observation, as I think is a
    rather known theorem in matrix algebra (see e.g. [this response in Math
    Stackexchange](https://math.stackexchange.com/a/340234/1082326 "Response to 'Transpose of inverse vs inverse of transpose'")).

[^7]: I'm omitting here (for the moment) the mathematical expressions defining
    the direction cosine vectors.

[^8]: Mario is using $V$ here but I understand he refers to the latent (vector)
    space, which I have denoted it by $`r LATENT_SPACE`$. Is this correct?

[^9]: I don't quite understand the whole proof, particularly, the equivalence
    $`r latex_eq(latex(latex_transp(STD_BASIS_VECTOR_ANY), INNER_PROD_MATRIX), INNER_PROD_MATRIX_ROW)`$,
    and the step from $`r DIR_COS_ANY`$ to $`r DIR_COS_VEC`$. This step is
    similar to my step from Equation 6 to Equation 7 in the draft, which I had
    the feeling was missing some intermediate steps.

[^10]: Again I'm using here the notation I have used in the draft and before
    (see also footnote 5).

[^11]: I have skipped some basic matrix algebra steps here.