### 2022-08-11 notebooks/Space_transformation_theorem.qmd
Direction cosines lemmas revised for errata

---
title: "Space transformation in Multidimensional IRT (Lemmas)"
author: "Mario Luzardo, Daniel Morillo"
format:
  docx:
    number-sections: true
    highlight-style: github  
callout-collapse: true
editor_options: 
  chunk_output_type: console
---

# Introduction

Transcription of Mario's notes

```{r setup, include=FALSE}
# Includes:

library(knitr)

ROOT_DIR <- ".."
DOC_DIR <- getwd()

opts_knit$set(root.dir = ROOT_DIR)
# opts_chunk$set(
#   echo     = FALSE,
#   cache    = FALSE,
#   dev      = "png",
#   dpi      =   300,
#   dev.args = list(type = "cairo-png"),
#   fig.pos = "!H"
# )
```

```{r sources, cache=FALSE, include=FALSE}
source("R/Formulae.R", encoding = 'UTF-8')
```

# Direction consines transformations

Let's consider a $`r N_DIMS`$-dimensional latent space $`r LATENT_SPACE`$ with
Euclidean vector structure, and let $`r LS_BASIS_EQ`$ be a basis of
$`r LATENT_SPACE`$.

::: callout-note
Stating that $`r LS_BASIS`$ is a basis implies that its vectors are linearly
independent and are a generator of $`r LATENT_SPACE`$; i.e., any vector
$`r TRAIT_VEC_IN_LS`$ can be expressed as a linear combination of
$`r LS_BASIS`$.
:::

Let's consider a dichotomous item $`r ITEM_INDEX`$ such that the probability
that a respondent with trait $`r TRAIT_VEC_IN_LS`$ gives a positive[^1] response
is given by

$$
  `r M2PL_FORMULATION`
$$ {#eq-M2PL-formulation}

Let's consider as well the standard basis $`r LS_STD_BASIS_EQ`$ in
$`r LATENT_SPACE`$.

For all $`r TRAIT_VEC_IN_LS`$ we have both representations[^2]:

$`r TRAIT_VECTOR_DEF_BASIS_EQ`$, and $`r TRAIT_VECTOR_DEF_STD_BASIS_EQ`$

Let $`r BASIS_CHANGE_DEF`$ be the change of basis matrix; then, if we denote
$`r TRAIT_COORDS_DEF`$ and $`r TRAIT_STD_COORDS_DEF`$ we have that

$$
  `r BASIS_CHANGE_EQ`.
$$ {#eq-change-basis}

The $`r DIM_INDEX`$-th column of $`r TRANSFORM_MATRIX`$ is given by the
coordinates of $`r BASIS_VECTOR_ANY`$ in basis $`r LS_STD_BASIS`$. If
$`r BASIS_VEC_STD_COORDS_DEF`$ then[^3]

$$
  `r TRANSFORM_MATRIX_EQ`
$$ {#eq-transform-matrix-def}

With the standard basis in $`r LATENT_SPACE`$ we have the usual scalar product.
With basis $`r LS_BASIS`$ in $`r LATENT_SPACE`$ we define an inner product
$`r INNER_PROD_BASIS`$ such that the norm of $`r TRAIT_VECTOR`$ is invariant,
i.e., $`r TRAIT_NORMS_EQ`$. Thus,[^4]

$$
  `r TRAIT_NORM_STD_EQ`.
$$

Therefore we must define the inner product $`r INNER_PROD_BASIS`$ with Gram
matrix $`r INNER_PROD_MAT_EQ`$. That is, $`r INNER_PROD_EQ`$,[^5] which keeps
the norm invariant.

The Gram matrix has as element $`r INNER_PROD_MAT_EL_EQ`$.[^6] As
$`r TRANSFORM_MATRIX`$ (and consequently, $`r TRANSFORM_MATRIX_TRANSP`$) is
invertible so $`r INNER_PROD_MATRIX`$ is, and
$`r INNER_PROD_MATRIX_INV_EQ`$.[^7]

Let's consider now the vectors $`r DIR_COS_VEC`$ and $`r DIR_COS_VEC_STD`$, made
up by the $`r N_DIMS`$ direction cosines of $`r TRAIT_VECTOR`$ with the basis
vectors of $`r LS_BASIS`$ and $`r LS_STD_BASIS`$, respectively.[^8]

::: {#lem-dircos}
Under the previously defined conditions, for all $`r TRAIT_VEC_IN_LS`$,[^9]

$$
  `r DIR_COS_VEC_EQ`
$$

with

$$
  `r DIAG_MATRIX_INNER_PROD_INV_SQ_EQ`.
$$

Therefore,

$$
  `r TRAIT_VECTOR_POLAR_EQ2`.
$$

::: proof
For every $`r DIM_INDEX`$; $`r DIM_ENUM_EQ`$,

$$
  `r DIR_COSINE_PROOF`,
$$

with $`r INNER_PROD_MATRIX_ROW`$ the $`r DIM_INDEX`$-th row of
$`r INNER_PROD_MATRIX`$.

Therefore, $`r DIR_COS_VEC_EQ2`$ and $`r TRAIT_VECTOR_POLAR_EQ2`$.
$\blacksquare$
:::
:::

::: {#lem-convert-spaces}
Let's consider another basis $`r NEW_BASIS_EQ`$ in $`r LATENT_SPACE`$ and the
inner product $`r INNER_PROD_NEWBS`$ that keeps the norm of the Euclidean space
invariant.

Let $`r DIR_COS_VEC_NEWBS`$ be the vector of direction cosines of
$`r TRAIT_VECTOR`$ in $`r NEW_BASIS`$ (i.e., the cosines with the elements of
$`r NEW_BASIS`$).

Then,

$$
  `r LEMMA_2_EQ`,
$$ {#eq-lemma-2}

where

$$
  `r DIAG_MATRIX_INNER_PROD_NEWBS_INV_SQ_EQ`
$$

and $`r BASIS_CHANGE_NEWBS_DEF`$.

::: proof
Let's call $`r TRAIT_NEWBS_COORDS_DEF`$ and let $`r BASIS_CHANGE_NEWBS_DEF`$ be
the change of basis matrix such that $`r BASIS_CHANGE_NEWBS`$.

We have that $`r BASES_CHANGE_EQ`$.

The inner product $`r INNER_PROD_NEWBS`$ that keeps the norm invariant is
defined by[^10] $`r INNER_PROD_NEWBS_EQ`$, and we define
$`r INNER_PROD_MAT_NEWBS_EQ`$.

By @lem-dircos, for each case we have:

1.  $`r TRAIT_VECTOR_POLAR_STD_EQ`$

2.  $`r TRAIT_VECTOR_POLAR_EQ2`$

3.  $`r TRAIT_VEC_NEWBS_POLAR_EQ`$

Given (1),

$$
  `r TRAIT_COORDS_EQUIV`
$$

$$
  `r TRAIT_COORDS_EQUIV_NEWBS`.
$$

Using (2) and (3) we get[^11]

$$
  `r TRAIT_COORDS_EQUIV_EXPANDED`
$$

$$
  `r TRAIT_COORDS_EQUIV_NEWBS_EXPANDED`.
$$

Therefore,

$$
  `r TRAIT_COORDS_EQUIV_EXPANDED_EQ`,
$$

and thus $`r LEMMA_2_EQ`$.$\blacksquare$
:::
:::

# Point of maximum slope

[^1]: I prefer to refer to the responses as "positive/negative", instead of
    "correct/incorrect", as this terminology allows for non-cognitive (as well
    as cognitive) domains.

[^2]: I'm preferring here the notation I use in the draft, that is,
    $`r DIM_INDEX`$ for indexing the dimensions (instead of $`r ITEM_INDEX`$,
    which is also used to index the items), $`r TRAIT_COMPONENT`$ for the
    coordinates of $`r TRAIT_VECTOR`$ in $`r LS_BASIS`$, and
    $`r TRAIT_VEC_STD_COMP`$ for the coordinates in $`r LS_STD_BASIS`$.
    Nevertheless, as Mario reminded me, it is worth noting that this notation
    for the vector coordinates is ambiguous, as we are using the same symbol
    ($`r TRAIT_VECTOR`$) for the "geometric vector" (i.e. "arrow" in vector
    space) and the "algebraic vector" (i.e. column matrix) of its coordinates in
    $`r LATENT_SPACE`$ (although this notation is usually accepted).

[^3]: Is it ok to use square brackets here? I'm using it instead of parentheses
    because I had already written this (with a function), but if it's not
    correct I change it to parentheses. **Also:** Is this a common algebra
    result? It feels like the this step is left implicit without justification;
    in my case, in the draft I needed to go all the way by Equations 2 through 4
    (section 2.2.1)

[^4]: Mario wrote in the leftmost therm of the equality "the norm of
    $`r TRAIT_VECTOR`$", but it should be "the *squared* norm of
    $`r TRAIT_VECTOR`$".

[^5]: Having already defined $`r TRAIT_COORDS_DEF`$, I have found this notation
    more convenient (consistent with what I had already written, for which I had
    it already coded).

[^6]: There is an erratum here in Mario's notes, I think. It should be the inner
    product of $`r BASIS_VECTOR_AUX_TRANSF`$ and $`r BASIS_VECTOR_ANY_TRANSF`$
    as I have written, instead of $`r BASIS_VECTOR_AUX`$ and
    $`r BASIS_VECTOR_ANY`$, as Mario wrote.

[^7]: I haven't considered it necessary to add the observation, as I think is a
    rather known theorem in matrix algebra (see e.g. [this response in Math
    Stackexchange](https://math.stackexchange.com/a/340234/1082326 "Response to 'Transpose of inverse vs inverse of transpose'")).

[^8]: I'm omitting here (for the moment) the mathematical expressions defining
    the direction cosine vectors.

[^9]: Mario is using $V$ here but I understand he refers to the latent (vector)
    space, which I have denoted it by $`r LATENT_SPACE`$. Is this correct?

[^10]: Again I'm using here the notation I have used in the draft and before
    (see also footnote 5).

[^11]: I have skipped some basic matrix algebra steps here.