### 2022-09-16 notebooks/Multidimensional_information.qmd
Multidimensional information notebook versioned

---
title: "Multidimensional information"
author: "Daniel Morillo"
format: html
bibliography: ../www/Multidimensional-parameters-MCLM.bib
editor: visual
---

# Introduction

Hereby I am going to derive the multidimensional information, according to
@ackerman_creating_1994.

# Derivations

@ackerman_creating_1994's Equations 1 through 7 and 13 through 16 for
bidimensional latent spaces apply directly to the 3-or-more dimensional case,
regardless of the latent covariance structure, as this is not involved at all in
those expressions. The only caveat is that he uses the normal parameterization,
and thus for the M2PL as we have been formulating it the 1.7 factor in Equation
1 does not apply (and subsequently, the 1.7\^2 = 2.89 factor in Equations 7 and
14).

The first deviation from his work comes in the paragraph right before Equation
13, where he states that

> the unidimensional trait level $\theta$ is equal to the linear composite,
> $\theta_c$, defined by $\theta_c = \beta_1 \theta_1 + \beta_2 \theta_2$, in
> which $\theta_c$ represents a linear composite in the angular direction of
> $\alpha$ and $\beta_1 = \cos \alpha$ and $\beta_2 = \sin \alpha$.

In our case, $\mathbf{\beta}$ must be a $D$-dimensional vector, being its $k$-th
component $\beta_k = \cos \alpha_k$, $k = [1, \dots, n]$.

## Components of $\mathbf{\beta}$

The main problem with determining the composite direction given by
$\mathbf{\beta}$ is that its components must be the cosines of the angles with
the $\theta_k$ axes. At this point, my main doubt is which space to use for
computing the angles: the latent trait space or the test space?

I will first try with a correlated, bidimensional latent space as an example to
explore the solution. In this space, the information of simple-structure items
in the direction parallel to the axes of its tapped trait should be equal to the
information of the unidimensional case. In the direction of the other trait
axis, the information should be more positive the more positive the correlation
is (constrained by the information of the former axis as maximum value).

I first define the latent space, by considering a bivariate standardize normal
distribution with correlated dimensions:

```{r}
rho     <- .5
sigma   <- c(1, rho, rho, 1) |> matrix(2)
sig_inv <- sigma |> solve()
```

I define the item discrimination parameters:

```{r}
a_item <- c(1, 0) |> matrix()
d_item <- 0
```

Then, the composite direction must be given by any arbitrary vector that is then
multiplied (by the inner product) with the basis vectors to be transformed into
a vector of cosines:

```{r}
comp_vec <- c(0, 1) |> matrix()
```

Now, computing the module of the basis vectors and the composite vector requires
the inner product matrix for the first time. I will try with the inverse of the
covariance matrix first.

```{r}
library(magrittr)

ip_matrix <- sig_inv

comp_vec_mod <- t(comp_vec) %*% ip_matrix %*% comp_vec |>
  drop() |>
  sqrt()

u_1 <- c(1, 0) |> matrix()
u_2 <- c(0, 1) |> matrix()

basis_matrix <- cbind(u_1, u_2)
basis_vec_mods <- t(basis_matrix) %*% ip_matrix %*% basis_matrix |>
  diag() |>
  sqrt()

comp_cos_vec <- t(comp_vec) %*% ip_matrix %*% basis_matrix |>
  divide_by(comp_vec_mod * basis_vec_mods) |>
  matrix()

comp_cos_vec
```

These are the cosines, the components of the vector $\beta$. The covariance
error matrix, Equation 13 in @ackerman_creating_1994, is given by the inverse of
the information matrix. The latter can be expressed as
$\mathbf{A}^T\mathbf{A}PQ$, being $\mathbf{A}$ the matrix of all item
parameters; the covariance matrix can be subsequently expressed as
$(\mathbf{A}^T\mathbf{A})^{-1}PQ^{-1}$. In our case, for a single item,
$\mathbf{A} = a^T$.

```{r}
A_matrix <- a_item |> t()

A_sq     <- t(A_matrix) %*% A_matrix
# A_sq_inv <- A_sq |> solve() ## Not run
```

As @ackerman_creating_1994 states (and I already knew) "for a single item or for
$n$ items that measure in exactly the same direction \[...\], Equation 7 (i.e.
the information matrix) is singular" (p. 259). This implies that I need at least
two items, and I want the information to be mostly aligned with one of the axes,
but they must not measure in the same direction. Therefore, I create a second
item, with near simple structure on the trait of the first axis.

```{r}
a_item_1 <- a_item
a_item_2 <- c(.95, .1) |> matrix()

A_matrix <- cbind(a_item_1, a_item_2) |> t()

A_sq     <- t(A_matrix) %*% A_matrix
A_sq_inv <- A_sq |> solve()
```

Once we have the error covariance matrix (except for the $PQ^{-1}$ factor), we
can pre- and post- multiply by (transposed) $\mathbf{\beta}$ to compute first
the error variance in that direction, and then its inverse to obtain the
information.

```{r}
error_var <- t(comp_cos_vec) %*% A_sq_inv %*% comp_cos_vec
comp_info <- 1 / error_var
```

The resulting multidimensional information coefficient for the direction
parallel to the second axis is `r comp_info`. Let's try to compute it in the
direction of the first axis:

```{r}
comp_vec     <- c(1, 0) |> matrix()
comp_vec_mod <- t(comp_vec) %*% ip_matrix %*% comp_vec |>
  drop() |>
  sqrt()
comp_cos_vec <- t(comp_vec) %*% ip_matrix %*% basis_matrix |>
  divide_by(comp_vec_mod * basis_vec_mods) |>
  matrix()

error_var <- t(comp_cos_vec) %*% A_sq_inv %*% comp_cos_vec
comp_info <- 1 / error_var
```

In the first axis the coefficient is `r comp_info`. This seems rather small when
one should expect a much higher information than in the second axis.