### 2022-09-28 notebooks/Multidimensional_information.qmd
Update to the multidimensional information notebook

---
title: "Multidimensional information"
author: "Daniel Morillo"
format:
  html:
    toc: true
    toc-location: left
    toc-depth: 4
bibliography: ../www/Multidimensional-parameters-MCLM.bib
editor: visual
---

```{r setup}
#| include: false
options(digits = 5)
```

# Introduction

Hereby I am going to try to derive the multidimensional information, according
to @ackerman_creating_1994.

# Derivations

## Multidimensional information

@ackerman_creating_1994's Equations 1 through 7 and 13 through 16 for
bidimensional latent spaces apply directly to the 3-or-more dimensional case,
regardless of the latent covariance structure, as this is not involved at all in
those expressions. The only caveat is that he uses the normal parameterization,
and thus for the M2PL as we have been formulating it the 1.7 factor in Equation
1 does not apply (and subsequently, the 1.7^2^ = 2.89 factor in Equations 7 and
14).

## Linear composite $\mathbf{\beta}$

The first deviation from his work comes in the paragraph right before Equation
13, where he states that

> the unidimensional trait level $\theta$ is equal to the linear composite,
> $\theta_c$, defined by $\theta_c = \beta_1 \theta_1 + \beta_2 \theta_2$, in
> which $\theta_c$ represents a linear composite in the angular direction of
> $\alpha$ and $\beta_1 = \cos \alpha$ and $\beta_2 = \sin \alpha$.

In our case, $\mathbf{\beta}$ must be a $D$-dimensional vector, being its $k$-th
component $\beta_k = \cos \alpha_k$, $k = [1, \dots, n]$.

### Components of $\mathbf{\beta}$

The main problem with determining the composite direction given by
$\mathbf{\beta}$ is that its components must be the cosines of the angles with
the $\theta_k$ axes. At this point, my main doubt is which space to use for
computing the angles: the latent trait space or the test space?

## Example with a correlated bidimensional space

I will first try with a correlated, bidimensional latent space as an example to
explore the solution. In this space, the information of simple-structure items
in the direction parallel to the axes of its tapped trait should be equal to the
information of the unidimensional case. In the direction of the other trait
axis, the information should be higher the more positive the correlation is
(constrained by the information of the former axis as maximum value).

I first define the latent space, by considering a bivariate standardized normal
distribution with correlated dimensions:

```{r def-corr}
rho <- .5
```

```{r def-covariance}
sigma   <- c(1, rho, rho, 1) |> matrix(2)
sig_inv <- sigma |> solve()
```

I define the item discrimination parameters:

```{r def-item}
a_item <- c(1, 0) |> matrix()
d_item <- 0
```

Then, the composite direction must be given by any arbitrary vector that is then
multiplied (by the inner product) with the basis vectors to be transformed into
a vector of cosines:

```{r def-composite}
comp_vec <- c(0, 1) |> matrix()
```

### Computation in the latent space {#latent-space}

Now, computing the module of the basis vectors and the composite vector requires
the inner product matrix for the first time. I will try with the inverse of the
covariance matrix first.

```{r def-inner-product}
ip_matrix <- sig_inv
```

```{r compute-composite-components}
library(magrittr)

comp_vec_mod <- t(comp_vec) %*% ip_matrix %*% comp_vec |>
  drop() |>
  sqrt()

u_1 <- c(1, 0) |> matrix()
u_2 <- c(0, 1) |> matrix()

basis_matrix   <- cbind(u_1, u_2)
basis_vec_mods <- t(basis_matrix) %*% ip_matrix %*% basis_matrix |>
  diag() |>
  sqrt()

comp_cos_vec <- t(comp_vec) %*% ip_matrix %*% basis_matrix |>
  divide_by(comp_vec_mod * basis_vec_mods) |>
  matrix()

comp_cos_vec
```

These are the cosines, the components of the vector $\beta$. The error
covariance matrix, Equation 13 in @ackerman_creating_1994, is given by the
inverse of the information matrix. The latter can be expressed as
$\mathbf{A}^T\mathbf{A}PQ$, being $\mathbf{A}$ the matrix of all item
parameters; the error covariance matrix can be subsequently expressed as
$(\mathbf{A}^T\mathbf{A})^{-1}PQ^{-1}$. In our case, for a single item,
$\mathbf{A} = a^T$.

```{r compute-discr-matrix-attemt}
A_matrix <- a_item |> t()

A_sq     <- t(A_matrix) %*% A_matrix
# A_sq_inv <- A_sq |> solve() ## Not run
```

As @ackerman_creating_1994 states "for a single item or for $n$ items that
measure in exactly the same direction \[...\], Equation 7 (i.e. the information
matrix) is singular" (p. 259). This I already knew, but hadn't realized it was
the case; therefore, my previous assertion that "in the direction of the other
trait axis, the information should be higher the more positive the correlation
is" is actually false (for one item). Therefore, I need at least two items, and
I want the information to be mostly aligned with one of the axes, but they must
not measure in the same direction. Thus I create a second item, with near simple
structure on the trait of the first axis.

```{r def-items}
a_item_1 <- a_item
a_item_2 <- c(1.5, .1) |> matrix()

A_matrix <- cbind(a_item_1, a_item_2) |> t()

A_sq     <- t(A_matrix) %*% A_matrix
A_sq_inv <- A_sq |> solve()
```

Once we have the error covariance matrix (except for the $PQ^{-1}$ factor), we
can pre- and post- multiply by (transposed) $\mathbf{\beta}$ to compute first
the error variance in that direction, and then its inverse to obtain the
information.

```{r compute-info-axis2}
error_var  <- t(comp_cos_vec) %*% A_sq_inv %*% comp_cos_vec
comp_info2 <- 1 / error_var
```

The resulting multidimensional information coefficient for the direction
parallel to the second axis is `r comp_info2`. Let's try to compute it in the
direction of the first axis:

```{r compute-info-axis1}
comp_vec     <- c(1, 0) |> matrix()
comp_vec_mod <- t(comp_vec) %*% ip_matrix %*% comp_vec |>
  drop() |>
  sqrt()
comp_cos_vec <- t(comp_vec) %*% ip_matrix %*% basis_matrix |>
  divide_by(comp_vec_mod * basis_vec_mods) |>
  matrix()

error_var  <- t(comp_cos_vec) %*% A_sq_inv %*% comp_cos_vec
comp_info1 <- 1 / error_var

ratio <- comp_info1 / comp_info2
```

In the first axis the coefficient is `r comp_info1`. This seems rather small
when one should expect a much higher information than in the second axis; the
ratio in the two directions is `r ratio`.

#### Test with higher correlation

Common sense tells us that the more aligned the axes are, the more similar the
information will be on the two dimensions parallel with the two axes. To test
this commonsense assumption, we try the previous computations again with a
higher correlation.

```{r test-higher-corr}
# Correlation and covariance matrix
rho       <- .93
sigma     <- c(1, rho, rho, 1) |> matrix(2)
sig_inv   <- sigma |> solve()
ip_matrix <- sig_inv

# Basis vectors
basis_vec_mods <- t(basis_matrix) %*% ip_matrix %*% basis_matrix |>
  diag() |>
  sqrt()

# Composite cosine vectors
comp_vecs     <- diag(2)
comp_vecs_mod <- t(comp_vecs) %*% ip_matrix %*% comp_vecs |>
  diag() |>
  sqrt()
comp_cos_vecs <- t(comp_vecs) %*% ip_matrix %*% basis_matrix |>
  divide_by(comp_vecs_mod * basis_vec_mods)

# Error variance and information in the composite directions
error_var <- t(comp_cos_vecs) %*% A_sq_inv %*% comp_cos_vecs |> diag()
comp_info <- 1 / error_var

ratio <- comp_info[1] / comp_info[2]
```

The information in the two axes is now `r comp_info[1]` and `r comp_info[2]`,
being their ratio `r ratio`. This fulfills the assumption we wanted to test.
However, it doesn't make much sense that the information in the direction of the
first axis has decreased.

### Computation in the test space

We are going to test the same assumption, but this time in the test space. This
implies using the covariance matrix as inner product matrix, instead of its
inverse.

```{r}
rho <- .5
```

We try first with a correlation of `r rho`.

```{r compute-test-space}
# Covariance matrix
sigma     <- c(1, rho, rho, 1) |> matrix(2)
ip_matrix <- sigma

# Basis vectors
basis_vec_mods <- t(basis_matrix) %*% ip_matrix %*% basis_matrix |>
  diag() |>
  sqrt()

# Composite cosine vectors
comp_vecs     <- diag(2)
comp_vecs_mod <- t(comp_vecs) %*% ip_matrix %*% comp_vecs |>
  diag() |>
  sqrt()
comp_cos_vecs <- t(comp_vecs) %*% ip_matrix %*% basis_matrix |>
  divide_by(comp_vecs_mod * basis_vec_mods)

# Error variance and information in the composite directions
error_var <- t(comp_cos_vecs) %*% A_sq_inv %*% comp_cos_vecs |> diag()
comp_info <- 1 / error_var

ratio <- comp_info[1] / comp_info[2]
```

The information in both axes directions would be `r comp_info[1]` and
`r comp_info[2]`, and their ratio `r ratio`.

```{r}
rho <- .93
```

Now we compute the same result, but with a correlation of `r rho` as we did
before.

```{r compute-test-space-high-corr}
# Covariance matrix
sigma     <- c(1, rho, rho, 1) |> matrix(2)
ip_matrix <- sigma

# Basis vectors
basis_vec_mods <- t(basis_matrix) %*% ip_matrix %*% basis_matrix |>
  diag() |>
  sqrt()

# Composite cosine vectors
comp_vecs     <- diag(2)
comp_vecs_mod <- t(comp_vecs) %*% ip_matrix %*% comp_vecs |>
  diag() |>
  sqrt()
comp_cos_vecs <- t(comp_vecs) %*% ip_matrix %*% basis_matrix |>
  divide_by(comp_vecs_mod * basis_vec_mods)

# Error variance and information in the composite directions
error_var <- t(comp_cos_vecs) %*% A_sq_inv %*% comp_cos_vecs |> diag()
comp_info <- 1 / error_var

ratio <- comp_info[1] / comp_info[2]
```

The information would be now `r comp_info[1]` and `r comp_info[2]` for the first
and second axis direction, respectively, and their ratio `r ratio`.

### Interpretation of the differences in the results in both spaces

The information in the direction of the first axis decreases also, as happened
when computing the directions in the latent space. However, the information in
the direction of the second axis increases, while when computing it in the
latent space it also decreased. It makes more sense to assume that the direction
would increase in the second axis direction with an increase in the correlation,
as it gets more aligned with the first axis.

On the other hand, exploring the cosines that determine the composite direction
one can see that, when using the latent space, they get more aligned **in the
opposite direction** when the correlation increases. This makes me think that,
in this case, the information is more evenly distributed across all direction in
the latent space. However, what I would expect is that the information get more
"concentrated" in a single direction.

### Correction about the computation of the information matrix

There is also one last thing when computing the information with two (or more)
items: I have assumed that the $PQ^{-1}$ factor can be factored out of the
matrix (see section [Computation in the latent space](#latent-space) above).
However, this is wrong. Its value is different for each item, and thus
it must be taken into account when computing the sum of the information for all
the items. This implies that the information matrix cannot be computed as
$\mathbf{A}^T\mathbf{A}PQ$ as I assumed before, for more than one item.