### 2022-10-21 notebooks/Distributions_Carlson_2017.qmd
Versions notebook about the issuess in Carlson's (2017) report

---
title: "Distributions in Carlson (2017)"
format: html
editor: visual
bibliography: ../www/Multidimensional-parameters-MCLM.bib
---

# Introduction

These distributions are taken from @carlson_unidimensional_2017's Table 2.

```{r table-2}
covs <- c(
  0.496, 0.138, 0.678,
  0.496, 0.223, 0.870,
  0.507, 0.328, 0.902,
  0.295, 0.428, 0.928,
  0.085, 0.383, 0.808,
  0.069, 0.546, 0.380
) |>
  matrix(ncol = 3, byrow = TRUE)

get_corr_mat <- function(vals) {

  cov <- vals[3] |> matrix(2, 2)
  diag(cov) <- vals[1:2]

  cov2cor(cov)
}

get_corr_mat(covs[1, ])
get_corr_mat(covs[2, ])
get_corr_mat(covs[3, ])
get_corr_mat(covs[4, ])
get_corr_mat(covs[5, ])
get_corr_mat(covs[6, ])
```