### 2022-12-22 notebooks/Distributions_Carlson_2017.qmd
Updates the loading of the items and starts the simulations

---
title: "Distributions in Carlson (2017)"
format: html
editor: visual
bibliography: ../www/Multidimensional-parameters-MCLM.bib
---

```{r setup}
#| message: false
library(tidyverse)
library(waldo)
library(mnormt)
```

```{r functions}
get_cov_mat <- function(vals) {

  vals <- vals                 |>
    pivot_longer(everything()) |>
    deframe()

  corr_matrix <- diag(2)
  corr_matrix[1, 2] <- corr_matrix[2, 1] <- vals["r"]

  vars <- vals['sd_' |> paste0(1:2)]

  diag(vars) %*% corr_matrix %*% diag(vars)
}

get_mean_vec <- function(vals) {

  vals                               |>
    pivot_longer(everything())       |>
    filter(name |> str_detect("^m_")) |>
    deframe()
}

simulate <- function(params, N = 2000) {

  covs  <- params |> get_cov_mat()
  means <- params |> get_mean_vec()

  rmnorm(N, mean = means, varcov = covs)
}
```

# Item parameters

The item parameters are taken from the Appendix, and are as follows:

```{r table-appendix}
items_table <- read_delim(
  "../dat/Table_Appendix_Carlson_2017.csv", 
  delim         = ";",
  escape_double = FALSE,
  trim_ws       = TRUE,
  col_types     = cols(
    Item           = col_character(),
    `Level admin.` = col_character(),
    .default       = col_double()
  )
)

# Set apart items and their descriptives:
suppressWarnings(
  items_table <- items_table |> mutate(
    is_descriptive = Item |>
      as.numeric()        |>
      is.na()
  )
)

item_parameters <- items_table |>
  filter(!is_descriptive)      |>
  select(-is_descriptive)      |>
  mutate(
    `Level admin.` = `Level admin.` |>
      str_split(", ")               |>
      map(parse_integer)
  )
```

We check that the items follow the specifications stated:

```{r check-items}
item_parameters          |>
  count(`Level admin.`)  |>
  unnest(`Level admin.`) |>
  count(`Level admin.`, wt = n)

item_parameters |>
  ggplot(aes(b_1, b_2, color = `Level admin.`)) +
  geom_point()
```

The number of items per level are as specified (but they are not coincident
with Table 1; there seems to be a mistake in that table). Also, the graphical
representation seems to be equivalent to @carlson_unidimensional_2017's Figure
5.

Now, we check whether the rectangular parameters are properly computed from the
polar parameters provided. In order to that, we apply Equations 23, 24, and 25
in @carlson_unidimensional_2017 to compute the rectangular parameters
(`d`, `b_m` `a_m`, respectively, for $m = {1, 2}$).

```{r check-item-parameters-rectangular}
computed_item_params <- item_parameters |> transmute(
  Item,
  b_1 =  `b*` * cospi(omega_1 / 180),
  b_2 =  `b*` * sinpi(omega_1 / 180),
  d   = -`b*` / `a*`,
  a_1 =  `a*` * cospi(omega_1 / 180),
  a_2 =  `a*` * sinpi(omega_1 / 180)
)

item_parameters |> select(Item, b_1:a_2) |>
  compare(computed_item_params, tolerance = 5e-3)
```

The values of `d` seem to be miscalculated.

We extract the descriptive statistics of the parameters and compare them with
the ones provided in the appendix table:

```{r check-item-descriptives}
items_descriptives <- items_table                                     |>
  filter(is_descriptive)                                              |>
  select(stat = Item, everything(), -is_descriptive, -`Level admin.`) |>
  pivot_longer(-stat, names_to = "parameter")                         |>
  mutate(across(-value, parse_factor))

# Check item descriptives:
computed_item_desc <- item_parameters                                      |>
  summarize(
    across(omega_1:a_2, .fns = lst(mean, sd, min, max))
  )                                                                        |>
  pivot_longer(everything())                                               |>
  separate(col = name, into = c("parameter", "stat"), sep = "_(?=[^_]+$)") |>
  mutate(
    across(-value, parse_factor),
    stat = stat                    |>
      fct_relabel(str_to_sentence) |>
      fct_recode(SD = "Sd")
  )                                                                        |>
  select(stat, parameter, value)                                           |>
  arrange(stat)

compare(items_descriptives, computed_item_desc, tolerance = 5e-3)
compare(items_descriptives, computed_item_desc, tolerance = 1e-3)
```

Coincidence is exact (up to the second decimal).

We will compute now the multidimensional parameters using these distribution
parameters:

```{r simulate}
params |> slice(1) |> get_cov_mat()
```

# Latent trait distributions

These distributions are taken from @carlson_unidimensional_2017's Table 2.
They are all bivariate normal with the following parameters:

```{r table-2}
params <- tribble(
  ~m_1,   ~m_2,   ~sd_1,  ~sd_2,  ~r,
  -3.228, -1.093,  0.496,  0.138,  0.678,
  -1.528, -0.779,  0.496,  0.223,  0.870,
  -0.265, -0.101,  0.507,  0.328,  0.902,
   0.598,  0.801,  0.295,  0.428,  0.928,
   1.100,  1.893,  0.085,  0.383,  0.808,
   1.330,  3.233,  0.069,  0.546,  0.380
)

params
```

# Simulations

2,000 latent trait values are drawn from bivariate normal distributions with the
parameters given above.

```{r simulate-traits}
set.seed(1325498132)

sim_traits <- params |>
  nrow()             |>
  seq_len()          |>
  map_df(
    ~params      |>
      slice(.x)  |>
      simulate() |>
      as_tibble(),
    .id = "level"
  )                  |>
  mutate(level = level |> parse_integer())

sim_traits                                |>
  mutate(level = level |> as.character()) |>
  ggplot(aes(V1, V2, color = level)) +
  geom_point(alpha = .1)
```

The resulting distributions quite resemble @carlson_unidimensional_2017's Figure
6.

With these values, item responses are first simulated to the items that
correspond to each level.

```{r simulate-responses}
sim_responses <- sim_traits |>
  full_join(
    item_parameters          |>
      unnest(`Level admin.`) |>
      select(Item, `Level admin.`, d:a_2),
    by = c(level = "Level admin.")
  )        |>
  mutate(
    prob     = V1 * a_1 + V2 * a_2 + d,
    response = prob |> rbinom(n = n(), size = 1)
  )
```