### 2024-01-12 notebooks/Mirt_solution_exploration.Rmd
Creates initial version of the notebook

---
title: 'Exploration of the MIRT solutions'
output:
  html_document: 
    df_print:        paged
    code_folding:    hide
    toc:             yes
    toc_float:       yes
editor_options: 
  chunk_output_type: console
---

```{r setup}
library(knitr)
library(mirt)

ROOT_DIR <- rstudioapi::getActiveProject()
DOC_DIR  <- getwd()

opts_knit$set(root.dir = ROOT_DIR)
opts_chunk$set(
  echo     = TRUE,
  message  = FALSE,
  cache    = FALSE
)

# Output configuration options:
options(digits = 3)
```

# Introduction

We want to explore the solutions of package `{mirt}` when computing latent trait
estimates, given that, apparently, it computes a rotated solution. Rotated
solutions typically try to find the simplest structure possible, thus performing
an oblique rotation, if necessary, to find the highest number of null
discrimination parameters.

However, the MDISC and MID parameters are possibly estimated from the original,
unrotated solution.