### 2024-02-02 notebooks/Mirt_solution_exploration.Rmd
Improves notebook output to make it clearer

---
title: 'Exploration of the MIRT solutions'
output:
  html_document: 
    df_print:        paged
    code_folding:    show
    toc:             yes
    toc_float:       yes
editor_options: 
  chunk_output_type: inline
---

```{r setup}
#| message: false
library(knitr)
library(mirt)
library(tidyverse)

ROOT_DIR <- rstudioapi::getActiveProject()
DOC_DIR  <- getwd()

opts_knit$set(root.dir = ROOT_DIR)
opts_chunk$set(
  echo     = TRUE,
  message  = FALSE,
  cache    = FALSE,
  results  = 'hold'
)

# Output configuration options:
options(digits = 3)
```

```{r sources}
source('R/Mirt_toolbox.R', encoding = 'UTF-8')
```

# Introduction

We want to explore the solutions of package `{mirt}` when computing latent trait
estimates, given that, apparently, it computes a rotated solution. Rotated
solutions typically try to find the simplest structure possible, thus performing
an oblique rotation, if necessary, to find the highest number of null
discrimination parameters.

However, the MDISC and MID parameters are possibly estimated from the original,
unrotated solution. Note that the values of these multidimensional parameters
is invariant to the rotation.

# Model estimation

We start by fitting a model with sample data bundled with the `{mirt}` package
itself. Let's start by using the first multidimensional example:

```{r fit-model}
#| message: false
data <- expand.table(LSAT7)
itemstats(data)

mod2 <- mirt(data, 2, verbose = FALSE)
coef(mod2)
summary(mod2, rotate = 'oblimin') #oblimin rotation
```

## Parameter computation

The default model is a 2-PL, so the items only have two discrimination
parameters (one for each dimension), and an intercept parameter. Parameters `g`
and `u` are restricted to `0` and `1`, respectively.

After applying the `oblimin` rotation, the item parameters and covariances are:

```{r rotated-irt-params}
coefs <- coef(mod2, rotate = 'oblimin') |> map(as_tibble)

item_pars <- coefs |>
  bind_rows()      |>
  slice(1:5)       |>
  select(a1:d)     |>
  rownames_to_column(var = "item")

cov_pars <- coefs[['GroupPars']] |>
  pivot_longer(everything())     |>
  deframe()

cov_matrix <- cov_pars[c('COV_11', 'COV_21', 'COV_21', 'COV_22')] |>
  matrix(nrow = 2)

item_pars
```

In order to check the result expressed in factor-analytic metric, we turn the
factor-analytic parameters into IRT metric to see whether they are coincident:

```{r param-equivalence}
sum <- summary(mod2, rotate = 'oblimin') #oblimin rotation
sum$rotF      |>
  as_tibble() |>
  mutate(
    across(everything(), ~1.702 * . / sqrt(1-as.numeric(sum$h2)))
  )           |>
  # Update the column names so they refer correctly to the discrimination
  #   parameters instead of the factor loadings.
  rename_with(str_replace, everything(), pattern = 'F', replacement = 'a')
```

The equivalence holds.

## Factor scores

Factor scores are computed by means of the `fscores()` function. By default,
`{mirt}` applies an `oblimin` rotation.

We use the original LSAT7 response patterns to compute the factor scores:

```{r compute-fscores}
fscores(mod2, response.pattern = LSAT7[, 1:5])
```

## Multidimensional IRT parameters

We compute the multidimensional IRT parameters using the `{mirt}` functions, and
also the IRT parameters of the rotated solution, and compare them.

```{r multidim-params}
mod2 |> MDISC()
mod2 |> MDIFF()

item_pars |>
  compute_mirt_params(d, a1:a2, cov_matrix = cov_matrix, dir_out = 'cos')
```

We see that they have the same values. Also, we can use the original, unrotated
solution:

```{r multidim-params-unrotated}
coefs_unrot <- coef(mod2) |> map(as_tibble)

item_pars_unrot <- coefs_unrot |>
  bind_rows()                  |>
  slice(1:5)                   |>
  select(a1:d)                 |>
  rownames_to_column(var = "item")

item_pars_unrot |>
  compute_mirt_params(d, a1:a2, dir_out = 'cos')
```

As expected, the MDISC and the location component of the MDIFF are coincident,
but the direction cosines change.

A visual inspection of the `MDISC()` and `MDIFF()` functions shows that the
parameters are computed applying Reckase's formulae to the unrotated solution.