### 2024-07-05 output/paper_draft/paper_M2PLM_parameters.Rmd
Moves properties to covariance-based section
The properties of the multidimensional parameters need to be in the covariance-based version section, because the noncompliance of this version with the "scale invariance" property (Property 1 in Reckase & McKinley, 1991) is used to justify and introduce the correlation-based version.

---
title: "A generalized definition of multidimensional item response theory parameters"
author: "Daniel Morillo"
date: "`r format(Sys.Date(), '%d/%m/%Y')`"
output:
  officedown::rdocx_document:
    tables:
      style: APA_Like_style
    mapstyles:
      Normal: First Paragraph
    base_format: bookdown::word_document2
    reference_docx: ../../www/Template.docx
    number_sections: no
    keep_md:         no
    fig_width:       4
    fig_height:      4
bibliography: [../../www/Multidimensional-parameters-MCLM.bib, ../../www/packages.bib]
csl:          ../../www/apa-old-doi-prefix.csl
editor_options:
  chunk_output_type: console
---

```{r setup}
#| include: false

# Includes:

library(knitr)
library(patchwork)
library(rstudioapi)

ROOT_DIR <- getActiveProject()
DOC_DIR  <- getwd()

opts_knit$set(root.dir = ROOT_DIR)
opts_chunk$set(
  echo     = FALSE,
  message  = FALSE,
  cache    = FALSE,
  dev      = "png",
  dpi      =   300,
  dev.args = list(type = "cairo-png")
)

# Output configuration options:
options(digits = 3)
```

```{r write-bib}
#| cache: true

# automatically create a bib database for R packages
write_bib(
  c(.packages(), 'bookdown', 'knitr', 'rmarkdown'),
  'www/packages.bib'
)
```

```{r sources}
#| cache:   false
#| include: false
source("R/Formulae.R",  encoding = 'UTF-8')
source("R/Constants.R", encoding = 'UTF-8')
```

# Introduction

Multidimensional item response theory (MIRT) models have been widely used since
their inception.<!--# Ref. needed? -->
The so-called *Multidimensional 2-Parameter Logistic*
[`r MODEL_ACRONYM`, @mckinley_extension_1983] model is among the most important
and widely used ones.
This is partly due to its close relationship to the common factor model
[@mcdonald_test_1999], but also to the formal tractability of its formulation.<!--# Ref. needed? -->
It consists of a generalization of the 2-parameter logistic model
[@birnbaum_latent_1968] to more than one latent trait
[@mckinley_extension_1983].
Its usefulness lies in describing the variability of the population when
the correct response does not depend on a single trait;
more specifically, when a deficit in a certain trait
and a surplus in another one cancel out their effects. Because of this property,
it is sometimes referred to as a *compensatory* model.

@reckase_difficulty_1985 argues for a need to take dimensionality into account
when characterizing the statistical properties of multidimensional items.
As such, he outlines a general procedure for determining their difficulty.
Following this procedure, he also derives
a *multidimensional difficulty* parameter (MID) for `r MODEL_ACRONYM` items.
This parameter consists of two parts:
The first one is a scalar value representing the distance from
the origin of the multidimensional latent space to
the maximally discriminating locus of the item.
The second one, the direction of the vector from the origin to that same locus.
Following a similar rationale, @reckase_discriminating_1991 derive a
*multidimensional discrimination* (`r MDISC_SYM`) parameter definition
for that model.

A crucial assumption leading to those formulations is that the axes of the
latent space are orthogonal
[@ackerman_using_1994; @ackerman_multidimensional_2005; see also
@reckase_difficulty_1985, p. 404 and Equation 7].
Of course, orthogonality is a property of a geometrical space instead of a
latent trait space.
Thus, nothing prevents a researcher from interpreting an abstraction such as
a latent trait space, no matter its structure, as an orthogonal space
[@ackerman_creating_1994]
---exploratory factor analysis software, such as SPSS <!--# Reference needed? -->or
the R package *psych*, <!--# Reference needed? -->graphically represents
factor loadings on orthogonal axes even after an oblique rotation.
Nevertheless, the assimilation of correlated latent traits to non-orthogonal
axes is prevalent in the common factor literature
[see, e.g., @harman_modern_1970].<!--# Thurstone's Multiple Factor Analysis, 1947? -->
This assumption is not so easily found in the MIRT literature though,
which mostly deals with uncorrelated dimensions
[see @reckase_multidimensional_2009 for a comprehensive treatment of the topic].
This fact is probably due to the different spotlights
these two theoretical programs put the focus on;
factor analysis aims at a clear interpretation of
the latent covariance structure, and thus focuses on finding
a structure matrix as simple as possible. MIRT literature, on the other hand,
aims at modeling responses to certain stimuli and estimating person scores,
being the model interpretation secondary in relevance.<!--# Reference needed? -->
Nevertheless, the confluence of the two theoretical approaches into a common
modern measurement theory [@mcdonald_test_1999] requires
the relaxation of traditional constraints on both approaches.
It is not uncommon to find recent examples of MIRT models where the
correlation matrix of the latent dimensions is freely estimated
[see, e.g., @reckase_multidimensional_2009, pp. 224-228]. Despite this, the
correspondence between correlated latent traits and orthogonal coordinate axes
is usually implicit in the MIRT literature without further consideration<!--# Examples? -->
[see @zhang_theoretical_1999 for a notable exception].
Moreover, disregarding this geometrical interpretation may lead to
misconceptions that put the measurement and its conclusions into question.<!-- completar?? -->

A latent space where the traits are correlated is easier to understand as a
geometrical space with non-orthogonal axes,
often referred to as *general Cartesian coordinates* [@harman_modern_1970].
Such a space has the advantage that the cosines of the angles between two axes
can be interpreted as the correlation of the corresponding latent dimensions.
<!--# Reference needed? -->
<!-- Problem here: The latent space is not defined like that, according to
the results of section "Linear algebra of non-orthogonal coordinates". -->
The general case of MIRT model estimation implies estimating possibly
non-null covariances among latent dimensions, as well as non-standard scales.
Assuming orthonormality among traits is usually unrealistic, be it in the
cognitive [e.g., @reckase_multidimensional_2009] or the non-cognitive domain<!--# Reference to example? -->.
Moreover, certain applications impose the estimation of
non-identity latent covariance matrices (e.g. multigroup equating) even when<!-- # Reference needed? -->
interpreting the latent dimensions as substantive constructs is not a goal.
As we will argue, the usual definitions of the MID and MDISC parameters
do not generalize to the non-orthonormal geometry
necessary to accurately represent correlated latent spaces.

In the non-cognitive domain, the use of so-called negatively-keyed items
--items that tap the negative pole of a certain trait,
usually reverse scored-- is also widespread.<!--# Reference and example? -->
It is common to find instruments that combine
positively- and negatively-keyed items.
Moreover, a complex item may have discrimination parameters with opposite signs
for different latent dimensions.<!--# Examples? -->
Take the multidimensional version of the nominal categories model
[@thissen_nominal_2010] as an example,
and consider two response categories tapping a different construct each:
A positive change in the dimension tapped by one category will inevitably
lead to a decrease in the probability of response to the other category.
One can easily note that a monotonically decreasing probability for a certain
dimension is given by a negative discrimination parameter
(and an unchanging probability by a null discrimination parameter),<!--# Actually, this also applies to the cognitive domain, which may not be very well explained. -->
contrary to the constraints imposed by @reckase_difficulty_1985 [p. 411].
However, the multidimensional item parameters can also be defined for
items with monotonically non-increasing probabilities, as we will see.

The main objective of this paper is thus to provide a generalization of
the `r MODEL_ACRONYM` multidimensional parameters in two aspects:
On one hand, in an oblique coordinate system
that can be soundly applied to the correlated trait case; on the other, to
items that yield monotonous response probabilities, but in the broader sense of
unchanging monotony. Additionally, we aim at providing a method for
graphically representing these items in the fashion of @ackerman_creating_1994
but taking the non-orthogonality of the coordinate axes into account.
In order to pursue these objectives, we first define the generalization of
the `r MODEL_ACRONYM` model and describe its scope of application.
Then, we follow the method outlined by @reckase_difficulty_1985 and
@reckase_discriminating_1991 to obtain
the formulation of the multidimensional parameters.
We will introduce two versions of these indices,
showing that either the original or one of the new ones is obtained
as a function of how we define and interpret the latent space.
Next, we discuss the properties of the indices,
propose a graphical representation method,
and present two practical examples applying them.
We round up with a discussion about
the applicability and generalization of our results.

# Method

## Model formulation

Consider a $`r N_DIMS`$-dimensional latent space $`r LATENT_SPACE`$
in basis $`r LS_BASIS_EQ`$, being $`r LS_BASIS`$ not necessarily orthonormal,
where a respondent is represented by vector $`r TRAIT_VEC_IN_LS`$.
According to the `r MODEL_ACRONYM` model, the probability of
a positive response to item $`r ITEM_INDEX`$ is

\begin{equation}
  `r M2PL_FORMULATION`,                                  (\#eq:M2PL-formulation)
\end{equation}

where $`r DISCR_VECTOR`$ is an $`r N_DIMS`$-dimensional vector of discrimination
parameters, and $`r INTERCEPT_PARAM`$ an intercept parameter related to the
location of the item in the latent space.

We must stop here and make a few remarks. First, as we are not constraining
ourselves to cognitive traits,
we speak of _positive_ instead of _correct response_. This may imply giving
a correct response indeed, if we are talking about a cognitive test.
But it may mean endorsing an item (or a certain response option)
in a non-cognitive domain.
Second, although not explicitly stated, $`r DISCR_VECTOR`$ is a vector of
non-negative real numbers in Reckase
[-@reckase_difficulty_1985, a decrease in the probability of a correct response
with an increase in ability makes little sense].
In our case, $`r DISCR_VECTOR`$ is a vector of real numbers; the value
of the $`r DIM_INDEX`$-th component $`r DISCR_PARAM`$ can be negative if
the item is *inverse* for (i.e., taps the negative pole of)
latent dimension $`r DIM_INDEX`$ (e.g. a negatively-keyed item).
Finally, as responses are not necessarily interpreted as correct or incorrect,
parameter $`r INTERCEPT_PARAM`$ is said to be related to the item _location_
instead of the item difficulty.

Following @reckase_difficulty_1985’s procedure,
we must find in the first place
the point of maximum slope of $`r IRF_ABBR`$ in $`r LATENT_SPACE`$.
Because the slope is dependent on the direction considered,
we need to reparameterize Equation \@ref(eq:M2PL-formulation)
in polar coordinates, as in @reckase_difficulty_1985.
However, Reckase's Equation 3 implicitly assumes orthogonality.
In order to contemplate the general case,
we need to introduce a few linear algebra results before proceeding further.

## Linear algebra of non-orthogonal coordinates

Let's consider an orthonormal basis $`r LS_ORTH_BASIS_EQ`$ in $`r LATENT_SPACE`$
with the dot product as inner product.
Let $`r TRAIT_COORDS_EQ`$ and $`r TRAIT_COORDS_ORTH_EQ`$ be the coordinates of
any $`r TRAIT_VEC_IN_LS`$ in $`r LS_BASIS`$ and $`r LS_ORTH_BASIS`$,
respectively, such that $`r TRAIT_VECTOR_DEF_EQ`$.

Let $`r BASIS_CHANGE_DEF`$ be the change of basis matrix between the two bases;
then, we have that

\begin{equation}
  `r BASIS_CHANGE`.                                          (\#eq:change-basis)
\end{equation}

The $`r DIM_INDEX`$-th column of $`r TRANSFORM_MATRIX`$ is given by the
coordinates of $`r BASIS_VECTOR_ANY`$ in basis $`r LS_ORTH_BASIS`$;
formally, if $`r BASIS_VECTOR_ANY_TRANSF_EQ`$, then

\begin{equation}
  `r TRANSFORM_MATRIX_EQ`.                           (\#eq:transform-matrix-def)
\end{equation}

As the norm of $`r TRAIT_VECTOR`$ must be invariant (i.e., $`r TRAIT_NORM_EQ`$),
we need to define $`r LS_BASIS`$ in $`r LATENT_SPACE`$ with inner product
$`r LS_BASIS_INNER_PROD`$ as

\begin{equation}
  `r TRAIT_NORM_SQ_EQ`.                                 (\#eq:inner-product-def)
\end{equation}

Thus, inner product $`r LS_BASIS_INNER_PROD`$ has $`r INNER_PROD_MAT_EQ`$ as
Gram matrix, i.e., $`r INNER_PROD_EQ`$, with $`r INNER_PROD_MAT_ELEMENT_EQ`$ as
the $`r INNER_PROD_MATRIX_INDEX`$-th element of $`r INNER_PROD_MATRIX`$.

### Test space definition

Let $`r TEST_SPACE`$ be the vector space made up by
the set of $`r DISCR_VECTOR`$ vectors,
with the operations sum and product by a scalar as in $`r REAL_N_SPACE`$.
We want @reckase_difficulty_1985's definition of
the multidimensional parameters in rectangular coordinates to be
a particular case of our more general definition;
we can thus consider an orthonormal basis $`r TEST_SPACE_STD_BASIS`$
in $`r TEST_SPACE`$ such that $`r DISCR_STD_COORDS_ALT_EQ`$
corresponds in the model to $`r TRAIT_ORTH_COORDS`$.
Since the response probability to the `r MODEL_ACRONYM`,
given by Equation \@ref(eq:M2PL-formulation), must be invariant with respect to
the change of basis expressed in Equation \@ref(eq:change-basis),
$`r INVARIANCE_EQ`$.
Therefore, we find a basis $`r TEST_SPACE_BASIS_EQ`$ in $`r TEST_SPACE`$
defined by

\begin{equation}
  `r DISCR_STD_COORDS_EQ`,                               (\#eq:discr-vector-std)
\end{equation}

with $`r DISCR_COORDS_ALT_EQ`$, and
$`r TRANSFORM_MATRIX_TRANSP_INV`$ a change of basis matrix.
Note that $`r TRANSFORM_MATRIX_TRANSP_INV`$ determines $`r TEST_SPACE_BASIS`$.

To make @reckase_difficulty_1985's definition a particular case of ours
we can conveniently assume that the inner product in
$`r TEST_SPACE_STD_BASIS`$ is the dot product,
as we did with $`r LS_ORTH_BASIS`$.
As $`r TRANSFORM_MATRIX`$ (and consequently $`r TRANSFORM_MATRIX_TRANSP`$) is
invertible, so $`r INNER_PROD_MATRIX`$ is, and $`r INNER_PROD_MATRIX_INV_EQ`$.
The norm of $`r DISCR_VECTOR`$ thus fulfills

\begin{equation}
  `r DISCR_VECTOR_MODULE_EQ`.
\end{equation}

Therefore, an inner product $`r TEST_SPACE_INNER_PROD`$ with Gram matrix
$`r INNER_PROD_MATRIX_INV`$ can be defined, such that $`r INNER_PROD_DISCR_EQ`$
keeps the norm of $`r DISCR_VECTOR`$ invariant.

### Direction cosines in the latent space

The cosine between two vectors

\begin{equation}
  `r COS_VECTORS_EQ`,                                         (\#eq:cos-vectors)
\end{equation}

is independent of whether the inner product is standard or not.
Therefore, to compute the direction cosine $`r DIR_COS_ANY_ORG`$
of vector $`r TRAIT_VECTOR`$ along dimension $`r DIM_INDEX`$,
we plug it into Equation \@ref(eq:cos-vectors) along with
the corresponding basis vector $`r BASIS_VECTOR_ANY`$:

\begin{equation}
  `r DIR_COSINE_EQ`.                                     (\#eq:direction-cosine)
\end{equation}

Let's consider now the vector $`r DIR_COS_VEC_ORG`$, made up by the $`r N_DIMS`$
direction cosines of $`r TRAIT_VECTOR`$ with
the basis vectors of $`r LS_BASIS`$.

::: {#prp-dir-cosines-generalized-basis}

**Proposition 1** Under the previously defined conditions, for all
$`r TRAIT_VEC_IN_LS`$,

\begin{equation}
  `r DIR_COS_VEC_EQ`                              (\#eq:direction-cosine-vector)
\end{equation}

with

\begin{equation}
  `r DIAG_MATRIX_INNER_PROD_EQ`.
\end{equation}

Therefore,

\begin{equation}
  `r TRAIT_VECTOR_POLAR_EQ`.                           (\#eq:trait-vector-polar)
\end{equation}

::: {.proof}
*Proof.*

For every $`r DIM_INDEX`$; $`r DIM_ENUM_EQ`$,

\begin{equation}
  `r DIR_COSINE_PROOF_EQ`,
\end{equation}

with $`r STD_BASIS_VECTOR_ANY`$ the $`r DIM_INDEX`$-th standard unitary vector,
and $`r INNER_PROD_MATRIX_ROW`$ the $`r DIM_INDEX`$-th row of
$`r INNER_PROD_MATRIX`$.

Therefore, $`r DIR_COS_VEC_EQ`$ and $`r TRAIT_VECTOR_POLAR_EQ`$.
$`r END_OF_PROOF`$
:::

:::

We can also obtain a result that relates the direction cosines of any two bases.

::: {#prp-dir-cosines-bases-relationship}

**Proposition 2** Let's consider another basis $`r ALT_BASIS_EQ`$ in
$`r LATENT_SPACE`$ and the inner product $`r ALT_BASIS_INNER_PROD`$ that keeps
the norm of the space invariant.

Let $`r DIR_COS_VEC_STD`$ and $`r DIR_COS_VEC_ALT`$ be the vectors of
direction cosines of $`r TRAIT_VECTOR`$ in $`r LS_ORTH_BASIS`$ and
$`r ALT_BASIS`$, respectively.

Then,

\begin{equation}
  `r DIR_COS_VEC_ALT_EQ`,                      (\#eq:dir-cosines-transformation)
\end{equation}

where

\begin{equation}
  `r DIAG_MATRIX_INNER_PROD_ALT_EQ`
\end{equation}

and $`r TRANSFORM_MATRIX_ALT_EQ`$.

::: {.proof}
*Proof.*

Let's call $`r TRAIT_COORDS_ALT_EQ`$ and let $`r TRANSFORM_MATRIX_ALT_EQ`$
be the change of basis matrix, such that $`r TRAIT_TRANSFORM_ALT_EQ`$.

Then, we have that $`r BASIS_CHANGE_ORG_ALT_EQ`$.

The inner product $`r ALT_BASIS_INNER_PROD`$ that keeps the norm invariant is
defined by $`r INNER_PROD_TRAIT_ALT_EQ`$, and we define
$`r INNER_PROD_MATRIX_ALT_EQ`$.

By proposition 1, for each case we have:

i. $`r TRAIT_VEC_ORTH_POLAR_EQ`$

i. $`r TRAIT_VECTOR_POLAR_EQ`$

i. $`r TRAIT_VEC_ALT_POLAR_EQ`$

Given 1,

\begin{equation}
  `r TRAIT_VEC_ORTH_POLAR_EQ_XPAND`
\end{equation}

and

\begin{equation}
  `r TRAIT_VEC_ALT_POLAR_ALT_EQ`.
\end{equation}

Using 2 and 3 we get that

\begin{equation}
  `r TRAIT_VEC_ORTH_POLAR_ORG_EQ`
\end{equation}

and

\begin{equation}
  `r TRAIT_VEC_ORTH_POLAR_ALT_EQ`.
\end{equation}

Therefore,

\begin{equation}
  `r TRAIT_VEC_POLAR_ORG_ALT_COEFF_EQ`,
\end{equation}

and thus $`r DIR_COS_VEC_ALT_EQ`$.$`r END_OF_PROOF`$

:::

:::

## Model recasting into polar coordinates

Applying Proposition 1 to
the model formulation in Equation \@ref(eq:M2PL-formulation),
the `r MODEL_ACRONYM` model is expressed in polar coordinates as

\begin{equation}
  `r IRF_POLAR_EQ`.                                    (\#eq:M2PL-formula-polar)
\end{equation}

As Equation \@ref(eq:M2PL-formula-polar) shows, the inner product matrix is
involved in the expression of the model in polar coordinates.
Importantly, this shows that @reckase_difficulty_1985's Equation 3 (p. 403)
already assumes the orthonormality of the latent space, as we stated before.

## Point of maximum slope

Following the procedure outlined by @reckase_difficulty_1985
[@reckase_discriminating_1991],
we first compute the point of maximum slope by finding the root(s) of
the second derivative with respect to $`r TRAIT_NORM`$.

The slope in direction $`r DIR_ANGLE_VEC_ORG`$ is given by

\begin{equation}
  `r IRF_1ST_DIFF_EQ`,                                      (\#eq:M2PL-1st-diff)
\end{equation}

and

\begin{equation}
  `r IRF_2ND_DIFF_EQ`,                                      (\#eq:M2PL-2nd-diff)
\end{equation}

The root we are interested in is found for $`r IRF_MAX_SLOPE_EQ`$
(the other ones occur when $`r IRF_ABBR`$ equals 0 and 1,
which result in improper values of $`r TRAIT_NORM`$).

The slope in direction $`r DIR_ANGLE_ITEM_VEC_EQ`$ when $`r IRF_MAX_SLOPE_EQ`$
is given by

\begin{equation}
  `r SLOPE_MAX_EQ`.                                             (\#eq:slope-max)
\end{equation}

Notation aside, these equations differ from
Reckase's [-@reckase_difficulty_1985] Equations 3 through 6
only in the term $`r INNER_PROD_INV_DIAG_SQ_PROD`$.

To compute the direction of the maximum slope,
Reckase leverages on the property that the squared direction cosines sum up to 1
(see Equation 7 therein).
However, this assertion implicitly assumes orthogonal coordinates,
so it does not apply in general to $`r DIR_COS_VEC_ORG`$.
Instead, we may consider the direction cosine vector with an orthonormal basis
$`r LS_ORTH_BASIS`$.
Now, applying Proposition 2 (with $`r TRANSFORM_MATRIX_ALT`$
an orthonormal matrix and $`r ALT_MATRIX_AS_ID`$) we obtain

\begin{equation}
  \begin{split}
    `r DIR_COS_AS_DIR_COS_ORTH_EQ`, \\
    `r TRAIT_POLAR_COEFF_COS_ORTH_EQ`.
  \end{split}                                 (\#eq:direction-cosine-vectors-eq)
\end{equation}

Therefore, substituting in \@ref(eq:slope-max),

\begin{equation}
  `r SLOPE_MAX_STD_COSINES_EQ`.                     (\#eq:slope-max-std-cosines)
\end{equation}

As $`r DISCR_VECTOR`$ is defined in space $`r TEST_SPACE`$ with basis
$`r TEST_SPACE_BASIS`$, its algebraic vector values correspond to its
coordinates, i.e., $`r DISCR_VECTOR_COORDS_EQ`$. Therefore, by Equation
\@ref(eq:discr-vector-std),

\begin{equation}
  `r SLOPE_MAX_STD_EQ`.                                     (\#eq:slope-max-std)
\end{equation}

We are thus considering the direction cosines with respect to the orthonormal
axes, so we can just apply @reckase_difficulty_1985's results, obtaining

\begin{equation}
  `r DIR_COS_VEC_STD_EQ`.                                 (\#eq:dir-cos-vec-std)
\end{equation}

Substituting $`r DIR_COS_ITEM_VEC_STD`$
(Equation \@ref(eq:direction-cosine-vectors-eq)) and
$`r DISCR_STD_COORDS`$ (Equation \@ref(eq:discr-vector-std))
by their respective expressions results in the
expression for the direction cosine vector of item $i$ in $`r LATENT_SPACE`$,

\begin{equation}
  `r DIR_COS_ITEM_VEC_EQ`,                                    (\#eq:dir-cos-vec)
\end{equation}

which gives the direction from the origin to the point where
the slope is maximum.
Finally, to determine the signed distance $`r DISTANCE_PARAM`$
from the origin to that point,
we solve Equation \@ref(eq:M2PL-formula-polar) for $`r IRF_MAX_SLOPE_EQ`$ to get
$`r TRAIT_NORM_SOLVED_EQ`$. Using equation \@ref(eq:dir-cos-vec),

\begin{equation}
  `r DISTANCE_PARAM_EQ`.                                   (\#eq:distance-param)
\end{equation}

The slope $`r SLOPE_MAX_PARAM`$ at the point defined by Equations
\@ref(eq:dir-cos-vec) and \@ref(eq:distance-param)
is given by Equation \@ref(eq:slope-max).
Substituting Equation \@ref(eq:dir-cos-vec) in Equation \@ref(eq:slope-max),
[@reckase_discriminating_1991],

\begin{equation}
  `r SLOPE_MAX_PARAM_EQ`,                                     (\#eq:slope-param)
\end{equation}

and thus, we have that

\begin{equation}
  \begin{split}
    `r DIR_COS_ITEM_VEC_NORM_EQ` \\
    `r DISTANCE_PARAM_NORM_EQ`   \\
    `r SLOPE_MAX_PARAM_NORM_EQ`.
  \end{split}                             (\#eq:item-params-discrimination-norm)
\end{equation}

# Results

## Relationship between test space and latent space

Before presenting the results about the parameters,
we must make a few remarks regarding the vector space $`r TEST_SPACE`$.
As shown before, this space is an entity by itself,
distinct from the latent space $`r LATENT_SPACE`$.
Despite this, the dot product in the exponent of Equation
\@ref(eq:M2PL-formulation) induces a bijective relationship between
their two sets of respective bases.
Explicitly, if $`r TRAIT_VECTOR`$ is represented in basis $`r LS_BASIS`$
with inner product Gram matrix $`r INNER_PROD_MATRIX`$,
$`r DISCR_VECTOR`$ is represented in basis $`r TEST_SPACE_BASIS`$
with inner product Gram matrix $`r INNER_PROD_MATRIX_INV`$,
for any pair of bases $`r LS_BASIS`$ and $`r TEST_SPACE_BASIS`$.
Only when the two bases are orthonormal,
the two spaces share a common inner product (i.e., the dot product),
and thus can be mistaken.
However, it is worth noting that, although their representation
can be superimposed in this case,
they are still two different spaces. <!-- # TODO: To discussion? -->
Following @zhang_theoretical_1999, we shall refer to space $`r TEST_SPACE`$ as
the *test space*, as it is the space where
the items (i.e., the test elements) are represented.

We must make a few observations here.
First, the transformation to the orthonormal bases of the two spaces allows
mapping the $`r MODEL_ACRONYM`$ parameters indistinctly in the two spaces,
by projecting one space onto the other.
This allows representing the coordinates of the discrimination vectors
in the latent space and, conversely,
the coordinates of the latent trait vectors in the test space.
As we can see, the invariance of the model holds,
no matter in which space we represent the coordinates.
<!-- # TODO: Incluir la condición de invarianza de las coordenadas en los dos espacios (Foto pizarra)? -->
Second, and most importantly, if the test space is
where the item parameters naturally lie, they must be represented in this space,
and not in the latent space.
Therefore, we may find the direction angles of $`r DISCR_VECTOR`$
in the test space:
As both $`r LS_ORTH_BASIS`$ and $`r TEST_SPACE_STD_BASIS`$ are orthonormal,
we can map one directly onto the other. Without loss of generality, we may thus
apply Proposition 2, so if $`r ANGLE_VEC_TS_ITEM`$ is
the vector of direction angles of $`r DISCR_VECTOR`$,
in the test space $`r TEST_SPACE`$ we get

\begin{equation}
  `r DIR_COS_ITEM_VEC_TS_EQ`,                      (\#eq:dir-cos-vec-test-space)
\end{equation}

with $`r DIAG_MATRIX_INNER_PROD_INV`$ a diagonal matrix where
$`r MATRIX_INNER_PROD_INV_EL_DIAG_EQ`$.
<!-- # TODO: Revisar esto bien. Esto sólo aplica correctamente si efectivamente
"equiparamos" las bases U y U*. Además, en la Proposición 2 estamos asumiendo
que la "base de enlace" es la canónica, no una base ortonormal cualquiera. -->
As we will see later, this representation is very convenient, as
the direction of the item vectors with respect to the coordinate axes
has a direct, meaningful interpretation.

## Generalized multidimensional parameters

The item location $`r MIL_PARAM`$ is defined as the distance and direction
from the origin to the point of maximum slope [@reckase_difficulty_1985].
By analogy with the unidimensional case, $`r MDISC_SYM`$ is defined as
$4 `r SLOPE_MAX_PARAM`$ [@reckase_discriminating_1991].
Therefore, we may define the $`r MIL_PARAM`$ and $`r MDISC_SYM`$ using
Equations \@ref(eq:distance-param) through \@ref(eq:dir-cos-vec-test-space).
However, as they depend on the inner product matrix $`r INNER_PROD_MATRIX`$,
the choice we make of this matrix will lead us to different definitions of
the parameters.

### Agnostic version

Up to this point,
we have derived the maximum slope and its location in the latent space
$`r LATENT_SPACE`$ independently of its basis $`r LS_BASIS`$.
If we assume that $`r LS_BASIS`$ is orthonormal, i.e., $`r BASIS_EQ`$,
$`r TRANSFORM_MATRIX`$ will also be orthonormal
and thus $`r INNER_PROD_MAT_STD_EQ`$.
In such a case the multidimensional item location and discrimination simplify
to the expressions derived by @reckase_difficulty_1985
and @reckase_discriminating_1991.
As these expressions implicitly assume the orthonormality of $`r LS_BASIS`$,
which bears no meaning besides its pure algebraic purpose,
we shall refer to them as the *agnostic* version of the parameters:

\begin{equation}
  \begin{split}
    `r MDISC_AG_PARAM_EQ` \\
    `r MIL_AG_PARAM_EQ`.
  \end{split}
\end{equation}

### Covariance-based version

On the other hand, we can let the geometrical representation of
the latent space $`r LATENT_SPACE`$ account for its covariance structure,
thus giving $`r LS_BASIS`$ a statistical meaning
in the context of our MIRT modeling approach.
To do that, we may assume that $`r TRAIT_COORDS`$ is
a random vector distributed with covariance $`r COV_MATRIX_EQ`$,
being $`r COV_MATRIX`$ an $`r N_DIMS`$-dimensional positive definite matrix,
$`r CORR_MATRIX`$ its corresponding correlation matrix
(also positive definite) and $`r SD_MATRIX`$ a scaling diagonal matrix
with $`r VAR_COV_EQ`$ (i.e., the variances).

To make the representation of $`r LATENT_SPACE`$ meaningful,
the axes in $`r LS_BASIS`$ must have
an interpretation in terms of the latent space structure.
Although we do not know, in principle, what
the interpretation of a latent space basis should be, we may assume that
it must be somewhat related to the covariance structure.
Therefore, an orthonormal basis should represent independent,
standard coordinates;
formally, if $`r COV_MATRIX_ORTH`$ is the covariance matrix of
$`r TRAIT_ORTH_COORDS`$, then $`r COV_MATRIX_ORTH_EQ`$.
Applying Equation \@ref(eq:change-basis), $`r COV_MATRIX_TRANSF_EQ`$, and thus,

\begin{equation}
  \begin{split}
    `r COV_MATRIX_TRANSF_ID_EQ` \\
    `r COV_MATRIX_RESULT_EQ`.
  \end{split}
\end{equation}

Therefore, we may define $`r TRANSFORM_MATRIX`$ such that
$`r INNER_PROD_COV_COND`$;
in this case, $`r DIAG_MATRIX_INNER_PROD_INV_VAR_EQ`$, resulting in a
*covariance-based* version of the parameters:

\begin{equation}
  \begin{split}
    `r MDISC_COV_PARAM_EQ` \\
    `r MIL_COV_PARAM_EQ`.
  \end{split}
\end{equation}

According to @reckase_discriminating_1991, the multidimensional parameters must
meet three conditions for being regarded as a valid generalization
of the (unidimensional) IRT parameters:

1. If an item measures only dimension $`r DIM_INDEX`$,
   then $`r MDISC_UNIDIM_EQ`$.

2. The distance $`r DISTANCE_PARAM`$ has the same relationship with the
   intercept as $b_i$ in the unidimensional case,
   i.e., $`r DIST_INTERCEPT_REL`$.

3. $`r MDISC_SYM`$ is four times the maximum slope $`r SLOPE_MAX_PARAM`$.

The second property derives straightforwardly from
Equation \@ref(eq:distance-param),
and the third one is implicit in the definition of the $`r MDISC_SYM`$.
However, when an item measures only dimension $`r DIM_INDEX`$,
simple arithmetic can show that $`r MDISC_UNIDIM_GENERALIZED_EQ`$.
In the covariance-based case, this means that $`r MDISC_COV_BASED_UNIDIM_EQ`$
and, as $`r COV_SYM_DIAG`$ is generally different from 1,
$`r MDISC_COV_PARAM`$ does not fulfill the first property
(we shall refer to this property as *scale invariance*, from now on).
Nevertheless, it can be interpreted as a *scaled version* of the parameter,
with the standard deviation of the corresponding dimension as scaling factor.
This transformation may be made in the unidimensional case as well,
thus finding a *standard-metric* discrimination parameter
(i.e., referred to a unitary variance space).

## Properties of the parameters

The (signed) distance to the origin,
given by the first component of $`r MIL_PARAM`$, has an interesting property:
By the definition of the inner product,
$`r MDISC_SYM`$ is strictly positive for non-null vectors.
Therefore, the sign of $`r DISTANCE_PARAM`$ is equal to
the sign of $`r INTERCEPT_PARAM`$.
The implication is that the item is displaced along
its measurement direction forwards or backwards with respect to the origin.

The measurement direction is given by the signs of $`r DIR_COS_ITEM_VEC`$ which,
also due to the strict positivity of $`r MDISC_SYM`$,
are equal to the signs of $`r DISCR_VECTOR`$.
This property leads directly to generalizing the parameters to the monotonically
non-increasing case: The measurement direction relative to dimension
$`r DIM_INDEX`$ will be negative when $`r IRF_ABBR`$ is
monotonically decreasing with respect to variations along $`r TRAIT_COMPONENT`$.
When $`r IRF_ABBR`$ is constant with respect to
variations along $`r TRAIT_COMPONENT`$, $`r SIGN_COS_VEC_ITEM_EQ`$,
which means that $`r MIL_PARAM`$ is orthogonal to the $`r DIM_INDEX`$-th axis.
However, note that this does not necessarily imply that the vector is
parallel to any other axis or strictly contained in the hyperplane formed
by other axes; this will always be true for the $`r MIL_AG_PARAM`$ due to the
orthogonality assumption, but it will depend on the correlation matrix
$`r CORR_MATRIX`$ for the other two versions of the parameter.

## Graphical representation

The aforementioned properties are easier to apprehend with
a visual representation of the items.
A vector plot [@ackerman_creating_1994] is the most appropriate way of
representing the items graphically;
it allows one to plot several items altogether and analyze them
in terms of their multidimensional parameters [@ackerman_graphical_1996]:
An item vector will be applied at a direction and (signed) distance
from the origin of the coordinate system given by its $`r MIL_PARAM`$ parameter
and will have a length equal to its $`r MDISC_SYM`$ parameter.

As plotting more than two dimensions is difficult on a bidimensional display
(and hardly possible at all for more than three dimensions with the
currently available technology),
we will limit ourselves to the bidimensional case here.
Plotting on a two-dimensional display usually requires providing
coordinates in a rectangular Cartesian system.
Therefore, we will need to compute the equivalent rectangular coordinates
of our general, possibly non-rectangular coordinates.
However, this implies pre-multiplying by $`r TRANSFORM_MATRIX`$,
instead of $`r TRANSFORM_MATRIX_TRANSP_INV`$ as would be expected from
Equation \@ref(eq:discr-vector-std).
This is because computing the equivalent rectangular coordinates
of the oblique ones is different from computing the item parameters
transformed to an orthogonal space.
While the latter implies rotating the axes to an orthogonal configuration and
computing the new coordinates in these axes,
the former implies transforming the space altogether along with the
axes transformation.

Therefore, we need to define a convenient value for $`r TRANSFORM_MATRIX`$
that allows us to represent the test space structure
and the items by transforming their coordinates to
a rectangular Cartesian system.
However, to facilitate the interpretation,
we want to avoid changing the scaling of the parameters.
We will thus prefer the correlation-based over the covariance-based parameters.
Thus, we define $`r TRANSFORM_MATRIX`$ such that
$`r TRANSFORM_MATRIX_TRANSP_INV`$ is a square root matrix of $`r CORR_MATRIX`$.

The general procedure for plotting an item $`r ITEM_INDEX`$ consists of
the following steps:

1. Define $`r TRANSFORM_MATRIX`$ such that $`r TRANSF_MATRIX_SQ_CORR_INV_EQ`$.

1. Compute $`r MDISC_CORR_PARAM_ITEM`$ and $`r MIL_CORR_PARAM_ITEM`$,
   the item $`r MDISC_CORR_PARAM`$ and $`r MIL_CORR_PARAM`$ parameters,
   respectively.

1. Compute the origin coordinates $`r ORIGIN_ITEM_EQ`$, with
   $`r DISTANCE_CORR_PARAM`$ and $`r DIR_CORR_PARAM`$ the distance and direction
   component, respectively, of $`r MIL_CORR_PARAM_ITEM`$.

1. Compute the end coordinates $`r END_ITEM_EQ`$.

1. Compute the rectangular coordinates
   $`r ORIGIN_ITEM_TRANSF`$ and $`r END_ITEM_TRANSF`$ by pre-multiplying
   $`r ORIGIN_ITEM`$ and $`r END_ITEM`$, respectively, by
   the transformation matrix $`r TRANSFORM_MATRIX`$
   ($`r ORIGIN_ITEM_TRANSF_EQ`$; $`r END_ITEM_TRANSF_EQ`$).

### Graphical representation example

```{r read-chunks-graphical-example}
read_chunk("src/Graphical_example_paper.R")
```

```{r libraries}
#| warning: false
```

```{r sources}
```

```{r constants}
```

```{r graphical-output-conf}
```

```{r compute-example-items}
```

The example in Figure \@ref(fig:item-plot-out) showcases a representation of
a set of items in two different bidimensional spaces.
In the first case, the two latent dimensions are independent,
with $`r TRANSFORM_MATRIX`$ being an identity matrix.
In the second one, the correlation between the dimensions is $`r CORR_OBL_OUT`$.
In order to keep the horizontal axis invariant,
we define $`r TRANSFORM_MATRIX`$ such that it only rotates the vertical axis
[@harman_modern_1970]:

\begin{equation}
  `r TRANSFORM_MATRIX_EXAMPLE_EQ`.
\end{equation}

The sample items have the same `r MODEL_ACRONYM` parameters
in both latent spaces.
However, their multidimensional parameters differ due to the correlation among
the latent dimensions.
The parameters in both spaces,
along with their corresponding `r MODEL_ACRONYM` parameters,
are shown in Table \@ref(tab:example-items-table-out).

```{r compose-example-items-table}
```

```{r set-flextable-wd}
#| cache: false

# Necessary for officedown to find the template file:
opts_knit$set(root.dir = DOC_DIR)
```

```{r example-items-table-out}
#| cache:   false
#| tab_id:  example-items-table-out
#| tab.cap: Item parameters for the graphical example
item_params_output
```

<br>

```{r compose-oblique-plot}
```

```{r compose-orthogonal-plot}
```

```{r item-plot-out}
#| cache:      false
#| fig.cap:   Item plots in an orthogonal (left) and oblique (right) test space.
#| fig.height: 3.1
#| fig.width:  6.65
plot_orth + plot_oblique
```

Figure \@ref(fig:item-plot-out) illustrates the effect of
a positive correlation on the $`r MDISC_CORR_PARAM`$
(and consequently the $`r MIL_CORR_PARAM`$) value:
The components of the discrimination parameters tend to
*sum up* in the same direction as they get *more aligned*.
Thus, if the discrimination parameters have the same sign, as in items 1 and 5,
the $`r MDISC_CORR_PARAM_ITEM`$ value tends to increase;
on the contrary, when the discrimination parameters have opposite signs
they tend to cancel each other out and $`r MDISC_CORR_PARAM_ITEM`$ decreases,
as happens with items 2 and 3.
This effect can be especially noticed by comparing
the $`r MDISC_CORR_PARAM_ITEM`$ of items 3 and 5:
Their discrimination parameters are equal in absolute value,
so their $`r MDISC_CORR_PARAM_ITEM`$ values are equal in the orthogonal space;
however, the effect of the correlation shrinks the former and
stretches the latter.
Finally, note that item 4 has one discrimination parameter equal to 0,
so its $`r MDISC_CORR_PARAM_ITEM`$ is unaffected by the correlation.

We can also see how the sign of $`r INTERCEPT_PARAM`$ affects
the distance to the origin:
The item vector is applied at the origin when its value is null (item 1),
and tends to be shifted
*against* the item direction when its value is positive (items 2 and 4) and
*towards* the item direction when it is negative (items 3 and 5).
Finally, we can see how the direction is determined by the discrimination
parameters *and* the correlation; the sign of the discrimination parameters
determines where the item will point at;
thus, an item with two positive (negative) discrimination parameters will point
at the first (third) quadrant,
while an item with opposite-sign discrimination parameters will be in the
second or fourth quadrant.
The direction relative to the axes will be given by the relative absolute value
of the two parameters, but also by the correlation:
Items 2 and 3 especially illustrate this effect,
as each of them is orthogonal to one of the axes in the oblique space,
even when they are not parallel to the other axis.

## Examples with empirical data

```{r reset-root-wd}
#| cache: false

# Necessary for knitr to find the source file:
opts_knit$set(root.dir = ROOT_DIR)
```

```{r read-chunks-empirical-example}
read_chunk("src/Empirical_example.R")
```

```{r libraries}
```

```{r sources}
```

```{r general-constants}
```

```{r set-cov-matrix-Reckase2009}
# Version in Reckase, 2009, p. 153; this represents a usual 3D case
COV_MATRIX <- matrix(
  c(
    1.210, .297, 1.232,
     .297, .810,  .252,
    1.232, .252, 1.960
  ),
  nrow = 3
)
# (see also Reckase, 2009, p. 183 for a "close-to-1D" case).
```

```{r read-items}
```

```{r compute-orthogonal-params}
```

```{r compute-oblique-params}
```

```{r create-params-table}
```

```{r compose-output-table}
```

```{r set-flextable-wd}
#| cache: false
```

```{r empirical-table-Reckase2009-out}
#| cache:   false
#| tab.id:  empirical-table-Reckase2009-out
#| tab.cap: Item parameters for the empirical example
item_params_output
```

```{r set-cov-matrix-Tezza2018}
# See Tezza et al., 2018, p. 927:
# > No presente estudo, a correlação entre a dimensão 1 e a dimensão 4
# > foi de aproximadamente 0,4.
COV_MATRIX <- diag(4)
COV_MATRIX[1, 4] <- COV_MATRIX[4, 1] <- .4
```

```{r reset-root-wd}
#| cache: false
```

```{r read-items-Tezza2018}
items <- read_csv2(ITEM_PARAMS_TEZZA_2018_FILEPATH)
```

```{r compute-orthogonal-params}
```

```{r compare-orthogonal-params}
#| results: hide
```

<!-- Computed item parameters in the orthonormal space match the ones in given
in Tezza 2018, table 5 -->

```{r compute-oblique-params}
```

```{r create-params-table}
```

```{r compose-output-table}
```

```{r set-flextable-wd}
#| cache: false
```

<!---BLOCK_LANDSCAPE_START--->

```{r empirical-table-Tezza2018-out}
#| cache:   false
#| tab.id:  empirical-table-Tezza2018-out
#| tab.cap: Item parameters for the empirical example
item_params_output
```

<!---BLOCK_LANDSCAPE_STOP--->

# Discussion

@reckase_difficulty_1985 proposed a definition of the $`r MIL_PARAM`$ parameter
general enough to "be used with any model that yields probabilities that
increase monotonically with an increase in ability on any dimension" (p. 411).
However, we have discussed that this definition implicitly assumes
the orthonormality of the latent space.
Applying the general procedure outlined by @reckase_difficulty_1985 and later
extended by @reckase_discriminating_1991
we have obtained a set of multidimensional parameters
that generalize the original results in two aspects:
(1) to a non-orthonormal space that accounts for the covariance structure of
the multivariate latent variable, and (2) to any case of unchanging monotony.
It is worth noting that @zhang_theoretical_1999 already proposed a formulation
of the $`r MDISC_SYM`$ equivalent to our covariance-based version,
but for a more general, semi-parametric formulation
(although constrained to non-decreasing monotony).
However, they provided no formal proof for this result.
Our results provide this formal derivation, although only for the specific
case of the `r MODEL_ACRONYM` model.

As we have seen, taking the latent space structure into account may have a
substantial effect in the resulting values and interpretation of the parameters,
something not contemplated by the original
(hereby referred to as agnostic) version.
Moreover, in applications where the latent space structure is critical
(e.g. multi-group IRT, equating, and linking),
the agnostic version may overlook relevant differences in
the values of the multidimensional parameters.
Thus, the covariance structure of the latent space should be considered
when computing the multidimensional parameters of a MIRT model.

With that purpose in mind, we have defined two versions of these parameters:
One that takes into account the whole covariance structure,
and another one that only considers the correlations among latent dimensions.
The covariance-based parameter has the drawback of not fulfilling
the first of @reckase_discriminating_1991's properties, namely,
that $`r MDISC_UNIDIM_EQ`$ if an item only measures dimension $`r DIM_INDEX`$.
However, we have provided a rationale for the violation of this property,
arguing that it is also paralleled in unidimensional IRT.
Given its equivalence to the inner product definition in a space with
an identity covariance matrix,
it also refers the item parameters to an orthonormal metric,
comparable among different latent spaces.
Once again, this is important in applications where
potentially different latent spaces are involved.
Given these considerations, we propose to adopt the covariance-based version of
the multidimensional parameters and refer to them simply as
the $`r MDISC_SYM`$ and $`r MIL_PARAM`$ parameters.
The correlation-based version, on the other hand,
has the advantage of preserving the metric of the parameters,
which is convenient for fidelity when representing the items.
Therefore, we recommend using the correlation-based version of the parameters
for graphical representation purposes only.

One may consider that the formal derivations presented here are overly
complicated,
when one might simply transform the space to rectangular Cartesian coordinates
with Equations \@ref(eq:change-basis) and \@ref(eq:discr-vector-std) and
then compute the parameters in the transformed space using
@reckase_difficulty_1985's and @reckase_discriminating_1991's procedure.
After all, interpreting the axes of this space as substantive latent traits
depends on "the distinction between coordinate axes and the constructs that are
the target of the instrument"
(Reckase, M.D., personal communication, 28 April 2015).
However, the computations in the oblique space allow us to obtain
the direction cosines relative to the original latent dimensions,
which may give relevant information if these are actual, substantive traits.
Moreover, strictly following @reckase_difficulty_1985's method in the original,
oblique space has also provided us with an important insight:
The distinction between the latent space and the test space.
Our derivations show that, in order to properly represent the items,
we need to define a vector space that has
the latent covariance (or correlation) matrix as its inner product matrix.
On the other hand, the latent space,
where the person parameters are represented,
has the inverse of the covariance (correlation) matrix as
its inner product matrix.
It is worth noting that a similar relationship exists in
the common factor literature
between the *primary factors* and Thurstone's *reference axes*
[@harman_modern_1970, see p. 291].
Nevertheless, one must not make the mistake of assimilating the test space and
the *common-factor space*.
Their parameterization, albeit related [@mcdonald_test_1999], is different,
and we do not mean to imply that the derivations made here are directly
generalizable to the common factor model.
Those relationships deserve further exploration on their own.

Finally, we must also warn against generalizing these results
to other IRT models without further consideration.
First, we have only considered the case of dichotomous items,
whereas non-cognitive applications often (and sometimes cognitive ones as well)
require modeling several response categories.
Although models such as the graded-scale [@samejima_estimation_1968] and
the nominal response [@bock_estimating_1972] model are closely related to
the 2-parameter logistic model, it would be hasty to assume that these
parameters generalize in any way to
the multidimensional versions of those models without a formal proof.
The same can be said from the multidimensional parameters of other models
for dichotomous items different from the `r MODEL_ACRONYM` model.
We expect this work to help set the foundations for investigating those
generalizations.
Hopefully, this will contribute to a better understanding
of the underlying measurement theory, its implications and its interpretation.

\newpage

# References