### 2024-07-18 output/paper_draft/paper_M2PLM_parameters.Rmd

---
title: "A generalized definition of multidimensional item response theory parameters"
author: "Daniel Morillo"
date: "`r format(Sys.Date(), '%d/%m/%Y')`"
output:
  officedown::rdocx_document:
    tables:
      style: APA_Like_style
    mapstyles:
      Normal: First Paragraph
    base_format: bookdown::word_document2
    reference_docx: ../../www/Template.docx
    number_sections: no
    keep_md:         no
    fig_width:       4
    fig_height:      4
bibliography: [../../www/Multidimensional-parameters-MCLM.bib, ../../www/packages.bib]
csl:          ../../www/apa-old-doi-prefix.csl
editor_options: 
  chunk_output_type: console
---

```{r setup}
#| include: false

# Includes:

library(knitr)
library(patchwork)
library(rstudioapi)

ROOT_DIR <- getActiveProject()
DOC_DIR  <- getwd()

opts_knit$set(root.dir = ROOT_DIR)
opts_chunk$set(
  echo     = FALSE,
  message  = FALSE,
  cache    = FALSE,
  dev      = "png",
  dpi      =   300,
  dev.args = list(type = "cairo-png")
)

# Output configuration options:
options(digits = 3)
```

```{r write-bib}
#| cache: true

# automatically create a bib database for R packages
write_bib(
  c(.packages(), 'bookdown', 'knitr', 'rmarkdown'),
  'www/packages.bib'
)
```

```{r sources}
#| cache:   false
#| include: false
source("R/Formulae.R",  encoding = 'UTF-8')
source("R/Constants.R", encoding = 'UTF-8')
```

# Introduction

Multidimensional item response theory (MIRT) models have been widely used since
their inception.
The so-called *Multidimensional 2-Parameter Logistic*
[`r MODEL_ACRONYM`, @mckinley_extension_1983] model is among the most important
and widely used ones.
This is partly due to its close relationship to the common factor model
[@mcdonald_test_1999], but also to the formal tractability of its formulation.
It consists of a generalization of the 2-parameter logistic model
[@birnbaum_latent_1968] to more than one latent trait
[@mckinley_extension_1983].
Its usefulness lies in describing the variability of the population when
the correct response does not depend on a single trait;
more specifically, when a deficit in a certain trait
and a surplus in another one cancel out their effects. Because of this property,
it is sometimes referred to as a *compensatory* model.

@reckase_difficulty_1985 argues for a need to take dimensionality into account
when characterizing the statistical properties of multidimensional items.
As such, he outlines a general procedure for determining their difficulty.
Following this procedure, he also derives
a *multidimensional difficulty* parameter (MID) for `r MODEL_ACRONYM` items.
This parameter consists of two parts:
The first one is a scalar value representing the distance from
the origin of the multidimensional latent space to
the maximally discriminating locus of the item;
the second one, the direction of the vector from the origin to that same locus.
Following a similar rationale, @reckase_discriminating_1991 derive a
*multidimensional discrimination* (`r MDISC_SYM`) parameter definition
for that model.

A crucial assumption leading to those formulations is that the axes of the
latent space are orthogonal
[@ackerman_using_1994; @ackerman_multidimensional_2005; see also
@reckase_difficulty_1985, p. 404 and Equation 7].
Of course, orthogonality is a property of a geometrical space instead of a
latent trait space.
Thus, nothing prevents a researcher from interpreting an abstraction such as
a latent trait space, no matter its structure, as an orthogonal space
---exploratory factor analysis software, such as SPSS [@ibm_corp_ibm_2022] or
the R package *psych* [@revelle_psych_2024], graphically represents
factor loadings on orthogonal axes even after an oblique rotation.
Nevertheless, the assimilation of correlated latent traits to non-orthogonal
axes is prevalent in the common factor literature
[see, e.g., @harman_modern_1970].
This assumption is not so easily found in the MIRT literature though,
which mostly deals with uncorrelated dimensions
[see @reckase_multidimensional_2009 for a comprehensive treatment of the topic].
This fact is probably due to the different spotlights
these two theoretical programs put the focus on;
factor analysis aims at a clear interpretation of
the latent covariance structure, and thus focuses on finding
a structure matrix as simple as possible. MIRT literature, on the other hand,
aims at modeling responses to certain stimuli and estimating person scores,
being the model interpretation secondary in relevance.<!--# Reference needed? -->
Nevertheless, the confluence of the two theoretical approaches into a common
modern measurement theory [@mcdonald_test_1999] requires
the relaxation of traditional constraints on both approaches.
It is not uncommon to find recent examples of MIRT models where the
correlation matrix of the latent dimensions is freely estimated
[see, e.g., @reckase_multidimensional_2009, pp. 224-228]. Despite this, the
correspondence between correlated latent traits and orthogonal coordinate axes
is usually implicit in the MIRT literature without further consideration
[see @zhang_theoretical_1999 for a notable exception].
Moreover, disregarding this geometrical interpretation may lead to
critical misconceptions about the measurement model.

The representation of a latent space where the traits are correlated is more
accurate on a geometrical coordinate system with non-orthogonal axes
[@ackerman_multidimensional_2005-1] and (possibly) non-standard units,
often referred to as *general Cartesian coordinates* [@harman_modern_1970].
Such a space has the advantage that the basis vectors and the axes they yield
have a meaningful interpretation (albeit counterintuitive, as we will see)
in a multivariate statistical sense.
The general case of MIRT model estimation implies estimating possibly
non-standard latent dimensions and non-null covariances among them
(a more realistic case than an identity matrix).
Assuming orthonormality among traits is usually unrealistic, be it in the
cognitive [e.g., @reckase_multidimensional_2009] or the non-cognitive domain
[e.g., @thielmann_comparability_2022].
Moreover, certain applications impose the estimation of
non-identity latent covariance matrices (e.g., multigroup equating) even when
interpreting the latent dimensions as substantive constructs is not a goal
[for for multidimensional linking and equating procedures,
see @reckase_linking_2009].
As we will argue, the usual definitions of the multidimensional difficulty and
discrimination parameters do not generalize to the non-orthonormal geometry
necessary to accurately represent correlated latent spaces.

On an unrelated basis, negative loadings/discrimination parameters
are often irrelevant in MIRT applications in the cognitive testing domain.
It has been argued that discrimination parameters
"are constrained to be positive" [@ackerman_graphical_1996, p. 315],
or that “in an [independent-clusters] solution, the loadings can always
be selected to be positive” [@mcdonald_basis_2000, p. 110].
However, when such instances of negative parameters appear,
they are often dismissed as “particularly puzzling” (p. 109),
or implicitly assuming item malfunctioning without further discussion
[e.g., @ackerman_multidimensional_2005-1].
In the non-cognitive domain though, the use of so-called negatively-keyed items
---items that tap the negative pole of a certain trait, and thus
usually reverse-scored--- is widespread
[consider, e.g., the statement "Neglect my duties" in the Conscientiousness item
of the Big-Five Factor Markers, @noauthor_international_nodate].
Instruments that combine positively- and negatively-keyed items
are relatively common, and
complex indicators may even have discrimination parameters
with opposite signs for different latent dimensions
[see, e.g., @mclarnon_be_2016,
for item parcels with both positive and negative loadings,
on method factors in the case of the negative ones].
Following @ackerman_multidimensional_2005-1, one can easily note that
a monotonically decreasing probability for a certain dimension
is given by a non-positive discrimination parameter
(and an unchanging probability by a null discrimination parameter,
which also applies to cognitive instruments),
contrary to the constraints imposed by @reckase_difficulty_1985 [p. 411]
for determining the multidimensional difficulty.
However, the multidimensional item parameters can also be defined for
items with monotonically non-increasing probabilities, as we will see.

The main objective of this paper is thus to provide a generalization of
the MIRT parameters in two aspects: on one hand, in an oblique coordinate system
that can be soundly applied to the correlated trait case; on the other, to
items that yield monotonous response probabilities, but in the broader sense of
unchanging monotony. Additionally, we aim to provide a method for
graphically representing these items in the fashion of @ackerman_creating_1994
but taking the non-orthogonality of the coordinate axes into account.
To pursue these objectives, we first define the generalization of
the `r MODEL_ACRONYM` model and describe its scope of application,
which will require introducing some linear algebra concepts and results.
Then, we follow the method outlined by @reckase_difficulty_1985 and
@reckase_discriminating_1991 to obtain
the formulation of the multidimensional parameters.
We will introduce three different versions,
showing that either the original or one of the new ones is obtained
as a function of how we define and interpret the latent space.
Next, we discuss their properties,
propose a graphical representation method based on the new formulation,
and present two examples of application of our results.
We round up with a discussion about
the applicability and generalization of our results.

# Method

## Model formulation

Consider a $`r N_DIMS`$-dimensional latent space $`r LATENT_SPACE`$
in basis $`r LS_BASIS_EQ`$, being $`r LS_BASIS`$ not necessarily orthonormal,
where a respondent is represented by vector $`r TRAIT_VEC_IN_LS`$.
According to the `r MODEL_ACRONYM` model, the probability of
a positive response to item $`r ITEM_INDEX`$ is

\begin{equation}
  `r M2PL_FORMULATION`,                                  (\#eq:M2PL-formulation)
\end{equation}

where $`r DISCR_VECTOR`$ is an $`r N_DIMS`$-dimensional vector of discrimination
parameters, and $`r INTERCEPT_PARAM`$ an intercept parameter related to the
location of the item in the latent space.

We must stop here and make a few remarks. First, as we are not constraining
ourselves to cognitive traits,
we speak of _positive_ instead of _correct response_. This may imply giving
a correct response indeed, if we are talking about a cognitive test.
But it may mean endorsing an item (or a certain response option)
in a non-cognitive domain.
Second, although not explicitly stated, $`r DISCR_VECTOR`$ is a vector of
non-negative real numbers in Reckase
[-@reckase_difficulty_1985, a decrease in the probability of a correct response
with an increase in ability makes little sense].
In our case, $`r DISCR_VECTOR`$ is a vector of real numbers; the value
of the $`r DIM_INDEX`$-th component $`r DISCR_PARAM`$ can be negative if
the item is *inverse* for (i.e., taps the negative pole of)
latent dimension $`r DIM_INDEX`$ (e.g., a negatively-keyed item).
Finally, as responses are not necessarily interpreted as correct or incorrect,
parameter $`r INTERCEPT_PARAM`$ is said to be related to the item _location_
instead of the item difficulty.

Following @reckase_difficulty_1985’s procedure,
we must find in the first place
the point of maximum slope of $`r IRF_ABBR`$ in $`r LATENT_SPACE`$.
Because the slope is dependent on the direction considered,
we need to reparameterize Equation \@ref(eq:M2PL-formulation)
in polar coordinates, as in @reckase_difficulty_1985.
However, Reckase's Equation 3 implicitly assumes orthogonality.
To contemplate the general case,
we need to introduce a few linear algebra results before proceeding further.

## Linear algebra of non-orthogonal coordinates

Let's consider an orthonormal basis $`r LS_ORTH_BASIS_EQ`$ in $`r LATENT_SPACE`$
with the dot product as inner product.
Let $`r TRAIT_COORDS_EQ`$ and $`r TRAIT_COORDS_ORTH_EQ`$ be the coordinates of
any $`r TRAIT_VEC_IN_LS`$ in $`r LS_BASIS`$ and $`r LS_ORTH_BASIS`$,
respectively, such that $`r TRAIT_VECTOR_DEF_EQ`$.

Let $`r BASIS_CHANGE_DEF`$ be the change of basis matrix between the two bases;
then, we have that

\begin{equation}
  `r BASIS_CHANGE`.                                          (\#eq:change-basis)
\end{equation}

The $`r DIM_INDEX`$-th column of $`r TRANSFORM_MATRIX`$ is given by the
coordinates of $`r BASIS_VECTOR_ANY`$ in basis $`r LS_ORTH_BASIS`$;
formally, if $`r BASIS_VECTOR_ANY_TRANSF_EQ`$, then

\begin{equation}
  `r TRANSFORM_MATRIX_EQ`.                           (\#eq:transform-matrix-def)
\end{equation}

As the norm of $`r TRAIT_VECTOR`$ must be invariant (i.e., $`r TRAIT_NORM_EQ`$),
we need to define $`r LS_BASIS`$ in $`r LATENT_SPACE`$ with inner product
$`r LS_BASIS_INNER_PROD`$ as

\begin{equation}
  `r TRAIT_NORM_SQ_EQ`.                                 (\#eq:inner-product-def)
\end{equation}

Thus, inner product $`r LS_BASIS_INNER_PROD`$ has $`r INNER_PROD_MAT_EQ`$ as
Gram matrix, i.e., $`r INNER_PROD_EQ`$, with $`r INNER_PROD_MAT_ELEMENT_EQ`$ as
the $`r INNER_PROD_MATRIX_INDEX`$-th element of $`r INNER_PROD_MATRIX`$.

### Test space definition

Let $`r TEST_SPACE`$ be the vector space made up by
the set of $`r DISCR_VECTOR`$ vectors,
with the operations sum and product by a scalar as in $`r REAL_N_SPACE`$.
We want @reckase_difficulty_1985's definition of
the multidimensional parameters in rectangular coordinates to be
a particular case of our more general definition;
we can thus consider an orthonormal basis $`r TEST_SPACE_STD_BASIS`$
in $`r TEST_SPACE`$ such that $`r DISCR_STD_COORDS_ALT_EQ`$
corresponds in the model to $`r TRAIT_ORTH_COORDS`$.
Since the response probability to the `r MODEL_ACRONYM`,
given by Equation \@ref(eq:M2PL-formulation), must be invariant with respect to
the change of basis expressed in Equation \@ref(eq:change-basis),
$`r INVARIANCE_EQ`$.
Therefore, we find a basis $`r TEST_SPACE_BASIS_EQ`$ in $`r TEST_SPACE`$
defined by

\begin{equation}
  `r DISCR_STD_COORDS_EQ`,                               (\#eq:discr-vector-std)
\end{equation}

with $`r DISCR_COORDS_ALT_EQ`$, and
$`r TRANSFORM_MATRIX_TRANSP_INV`$ a change of basis matrix.
Note that $`r TRANSFORM_MATRIX_TRANSP_INV`$ determines $`r TEST_SPACE_BASIS`$.

To make @reckase_difficulty_1985's definition a particular case of ours
we can conveniently assume that the inner product in
$`r TEST_SPACE_STD_BASIS`$ is the dot product,
as we did with $`r LS_ORTH_BASIS`$.
As $`r TRANSFORM_MATRIX`$ (and consequently $`r TRANSFORM_MATRIX_TRANSP`$) is
invertible, so $`r INNER_PROD_MATRIX`$ is, and $`r INNER_PROD_MATRIX_INV_EQ`$.
The norm of $`r DISCR_VECTOR`$ thus fulfills

\begin{equation}
  `r DISCR_VECTOR_MODULE_EQ`.
\end{equation}

Therefore, an inner product $`r TEST_SPACE_INNER_PROD`$ with Gram matrix
$`r INNER_PROD_MATRIX_INV`$ can be defined, such that $`r INNER_PROD_DISCR_EQ`$
keeps the norm of $`r DISCR_VECTOR`$ invariant.
Following @zhang_theoretical_1999, we shall refer to space $`r TEST_SPACE`$ as
the *test space*, as it is the space where
the items (i.e., the test elements) are represented.

### Direction cosines in the latent space

The cosine between two vectors

\begin{equation}
  `r COS_VECTORS_EQ`,                                         (\#eq:cos-vectors)
\end{equation}

is independent of whether the inner product is standard or not.
Therefore, to compute the direction cosine $`r DIR_COS_ANY_ORG`$
of vector $`r TRAIT_VECTOR`$ along dimension $`r DIM_INDEX`$,
we plug it into Equation \@ref(eq:cos-vectors) along with
the corresponding basis vector $`r BASIS_VECTOR_ANY`$:

\begin{equation}
  `r DIR_COSINE_EQ`.                                     (\#eq:direction-cosine)
\end{equation}

Let's consider now the vector $`r DIR_COS_VEC_ORG`$, made up by the $`r N_DIMS`$
direction cosines of $`r TRAIT_VECTOR`$ with
the basis vectors of $`r LS_BASIS`$.

::: {#prp-dir-cosines-generalized-basis}

**Proposition 1** Under the previously defined conditions, for all
$`r TRAIT_VEC_IN_LS`$,

\begin{equation}
  `r DIR_COS_VEC_EQ`                              (\#eq:direction-cosine-vector)
\end{equation}

with

\begin{equation}
  `r DIAG_MATRIX_INNER_PROD_EQ`.
\end{equation}

Therefore,

\begin{equation}
  `r TRAIT_VECTOR_POLAR_EQ`.                           (\#eq:trait-vector-polar)
\end{equation}

::: {.proof}
*Proof.*

For every $`r DIM_INDEX`$; $`r DIM_ENUM_EQ`$,

\begin{equation}
  `r DIR_COSINE_PROOF_EQ`,
\end{equation}

with $`r STD_BASIS_VECTOR_ANY`$ the $`r DIM_INDEX`$-th standard unitary vector,
and $`r INNER_PROD_MATRIX_ROW`$ the $`r DIM_INDEX`$-th row of
$`r INNER_PROD_MATRIX`$.

Therefore, $`r DIR_COS_VEC_EQ`$ and $`r TRAIT_VECTOR_POLAR_EQ`$.
$`r END_OF_PROOF`$
:::

:::

We can also obtain a result that relates the direction cosines of any two bases.

::: {#prp-dir-cosines-bases-relationship}

**Proposition 2** Let's consider another basis $`r ALT_BASIS_EQ`$ in
$`r LATENT_SPACE`$ and the inner product $`r ALT_BASIS_INNER_PROD`$ that keeps
the norm of the space invariant.

Let $`r DIR_COS_VEC_STD`$ and $`r DIR_COS_VEC_ALT`$ be the vectors of
direction cosines of $`r TRAIT_VECTOR`$ in $`r LS_ORTH_BASIS`$ and
$`r ALT_BASIS`$, respectively.

Then,

\begin{equation}
  `r DIR_COS_VEC_ALT_EQ`,                      (\#eq:dir-cosines-transformation)
\end{equation}

where

\begin{equation}
  `r DIAG_MATRIX_INNER_PROD_ALT_EQ`
\end{equation}

and $`r TRANSFORM_MATRIX_ALT_EQ`$.

::: {.proof}
*Proof.*

Let's call $`r TRAIT_COORDS_ALT_EQ`$ and let $`r TRANSFORM_MATRIX_ALT_EQ`$
be the change of basis matrix, such that $`r TRAIT_TRANSFORM_ALT_EQ`$.

Then, we have that $`r BASIS_CHANGE_ORG_ALT_EQ`$.

The inner product $`r ALT_BASIS_INNER_PROD`$ that keeps the norm invariant is
defined by $`r INNER_PROD_TRAIT_ALT_EQ`$, and we define
$`r INNER_PROD_MATRIX_ALT_EQ`$.

By proposition 1, for each case we have:

i. $`r TRAIT_VEC_ORTH_POLAR_EQ`$

i. $`r TRAIT_VECTOR_POLAR_EQ`$

i. $`r TRAIT_VEC_ALT_POLAR_EQ`$

Given i,

\begin{equation}
  `r TRAIT_VEC_ORTH_POLAR_EQ_XPAND`
\end{equation}

and

\begin{equation}
  `r TRAIT_VEC_ALT_POLAR_ALT_EQ`.
\end{equation}

Using ii and iii we get that

\begin{equation}
  `r TRAIT_VEC_ORTH_POLAR_ORG_EQ`
\end{equation}

and

\begin{equation}
  `r TRAIT_VEC_ORTH_POLAR_ALT_EQ`.
\end{equation}

Therefore,

\begin{equation}
  `r TRAIT_VEC_POLAR_ORG_ALT_COEFF_EQ`,
\end{equation}

and thus $`r DIR_COS_VEC_ALT_EQ`$.$`r END_OF_PROOF`$

:::

:::

## Model recasting into polar coordinates

Applying Proposition 1 to
the model formulation in Equation \@ref(eq:M2PL-formulation),
the `r MODEL_ACRONYM` model is expressed in polar coordinates as

\begin{equation}
  `r IRF_POLAR_EQ`.                                    (\#eq:M2PL-formula-polar)
\end{equation}

As Equation \@ref(eq:M2PL-formula-polar) shows, the inner product matrix is
involved in the expression of the model in polar coordinates.
Importantly, this shows that @reckase_difficulty_1985's Equation 3 (p. 403)
already assumes the orthonormality of the latent space, as we stated before.

## Point of maximum slope

Following the procedure outlined by @reckase_difficulty_1985
[@reckase_discriminating_1991],
we first compute the point of maximum slope by finding the root(s) of
the second derivative with respect to $`r TRAIT_NORM`$.

The slope in direction $`r DIR_ANGLE_VEC_ORG`$ is given by

\begin{equation}
  `r IRF_1ST_DIFF_EQ`,                                      (\#eq:M2PL-1st-diff)
\end{equation}

and

\begin{equation}
  `r IRF_2ND_DIFF_EQ`,                                      (\#eq:M2PL-2nd-diff)
\end{equation}

The root we are interested in is found for $`r IRF_MAX_SLOPE_EQ`$
(the other ones occur when $`r IRF_ABBR`$ equals 0 and 1,
which result in improper values of $`r TRAIT_NORM`$).

The slope in direction $`r DIR_ANGLE_ITEM_VEC_EQ`$ when $`r IRF_MAX_SLOPE_EQ`$
is given by

\begin{equation}
  `r SLOPE_MAX_EQ`.                                             (\#eq:slope-max)
\end{equation}

Notation aside, these equations differ from
Reckase's [-@reckase_difficulty_1985] Equations 3 through 6
only in the term $`r INNER_PROD_INV_DIAG_SQ_PROD`$.

To compute the direction of the maximum slope,
Reckase leverages on the property that the squared direction cosines sum up to 1
(see Equation 7 therein).
However, this assertion implicitly assumes orthogonal coordinates,
so it does not apply in general to $`r DIR_COS_VEC_ORG`$.
Instead, we may consider the direction cosine vector with an orthonormal basis
$`r LS_ORTH_BASIS`$.
Now, applying Proposition 2 (with $`r TRANSFORM_MATRIX_ALT`$
an orthonormal matrix and $`r ALT_MATRIX_AS_ID`$) we obtain

\begin{equation}
  \begin{split}
    `r DIR_COS_AS_DIR_COS_ORTH_EQ`, \\
    `r TRAIT_POLAR_COEFF_COS_ORTH_EQ`.
  \end{split}                                 (\#eq:direction-cosine-vectors-eq)
\end{equation}

Therefore, substituting in \@ref(eq:slope-max),

\begin{equation}
  `r SLOPE_MAX_STD_COSINES_EQ`.                     (\#eq:slope-max-std-cosines)
\end{equation}

As $`r DISCR_VECTOR`$ is defined in space $`r TEST_SPACE`$ with basis
$`r TEST_SPACE_BASIS`$, its algebraic vector values correspond to its
coordinates, i.e., $`r DISCR_VECTOR_COORDS_EQ`$. Therefore, by Equation
\@ref(eq:discr-vector-std),

\begin{equation}
  `r SLOPE_MAX_STD_EQ`.                                     (\#eq:slope-max-std)
\end{equation}

We are thus considering the direction cosines with respect to the orthonormal
axes, so we can just apply @reckase_difficulty_1985's results, obtaining

\begin{equation}
  `r DIR_COS_VEC_STD_EQ`.                                 (\#eq:dir-cos-vec-std)
\end{equation}

Substituting $`r DIR_COS_ITEM_VEC_STD`$
(Equation \@ref(eq:direction-cosine-vectors-eq)) and
$`r DISCR_STD_COORDS`$ (Equation \@ref(eq:discr-vector-std))
by their respective expressions results in the
expression for the direction cosine vector of item $i$ in $`r LATENT_SPACE`$,

\begin{equation}
  `r DIR_COS_ITEM_VEC_EQ`,                                    (\#eq:dir-cos-vec)
\end{equation}

which gives the direction from the origin to the point where
the slope is maximum.
Finally, to determine the signed distance $`r DISTANCE_PARAM`$
from the origin to that point,
we solve Equation \@ref(eq:M2PL-formula-polar) for $`r IRF_MAX_SLOPE_EQ`$ to get
$`r TRAIT_NORM_SOLVED_EQ`$. Using equation \@ref(eq:dir-cos-vec),

\begin{equation}
  `r DISTANCE_PARAM_EQ`.                                   (\#eq:distance-param)
\end{equation}

The slope $`r SLOPE_MAX_PARAM`$ at the point defined by Equations
\@ref(eq:dir-cos-vec) and \@ref(eq:distance-param)
is given by Equation \@ref(eq:slope-max).
Substituting Equation \@ref(eq:dir-cos-vec) in Equation \@ref(eq:slope-max),
[@reckase_discriminating_1991],

\begin{equation}
  `r SLOPE_MAX_PARAM_EQ`,                                     (\#eq:slope-param)
\end{equation}

and thus, we have that

\begin{equation}
  \begin{split}
    `r DIR_COS_ITEM_VEC_NORM_EQ` \\
    `r DISTANCE_PARAM_NORM_EQ`   \\
    `r SLOPE_MAX_PARAM_NORM_EQ`.
  \end{split}                             (\#eq:item-params-discrimination-norm)
\end{equation}

# Results

## Relationship between test space and latent space

Before presenting the results about the parameters,
we must make a few remarks regarding the vector space $`r TEST_SPACE`$.
As shown before, this space is an entity by itself,
distinct from the latent space $`r LATENT_SPACE`$.
Despite this, the dot product in the exponent of Equation
\@ref(eq:M2PL-formulation) induces a bijective relationship between
their two sets of respective bases.
Explicitly, if $`r TRAIT_VECTOR`$ is represented in basis $`r LS_BASIS`$
with inner product Gram matrix $`r INNER_PROD_MATRIX`$,
$`r DISCR_VECTOR`$ is represented in basis $`r TEST_SPACE_BASIS`$
with inner product Gram matrix $`r INNER_PROD_MATRIX_INV`$,
for any pair of bases $`r LS_BASIS`$ and $`r TEST_SPACE_BASIS`$.
Only when the two bases are orthonormal,
the two spaces share a common inner product (i.e., the dot product),
and thus can be mistaken.
However, it is worth noting that, although their representation
can be superimposed in this case,
they are still two different spaces. <!-- # TODO: To discussion? -->

We must make a few observations here.
First, the transformation to the orthonormal bases of the two spaces allows
mapping the $`r MODEL_ACRONYM`$ parameters indistinctly in the two spaces,
by projecting one space onto the other.
This allows representing the coordinates of the discrimination vectors
in the latent space and, conversely,
the coordinates of the latent trait vectors in the test space.
As we can see, the invariance of the model holds,
no matter in which space we represent the coordinates.
<!-- # TODO: Incluir la condición de invarianza de las coordenadas en los dos espacios (Foto pizarra)? -->
Second, and most importantly, if the test space is
where the item parameters naturally lie, they must be represented in this space,
and not in the latent space.
Therefore, we may find the direction angles of $`r DISCR_VECTOR`$
in the test space:
As $`r LS_ORTH_BASIS`$ and $`r TEST_SPACE_STD_BASIS`$ are both
orthonormal in $`r REAL_N_SPACE`$, without loss of generality
we can consider that $`r BASES_EQ`$ (save for a rotation).
We can then map the latent space directly onto the test space and
apply Proposition 2, so if $`r ANGLE_VEC_TS_ITEM`$ is
the vector of direction angles of $`r DISCR_VECTOR`$,
in the test space $`r TEST_SPACE`$ we get

\begin{equation}
  `r DIR_COS_ITEM_VEC_TS_EQ`,                      (\#eq:dir-cos-vec-test-space)
\end{equation}

with $`r DIAG_MATRIX_INNER_PROD_INV`$ a diagonal matrix where
$`r MATRIX_INNER_PROD_INV_EL_DIAG_EQ`$.
As we will see later, this representation is very convenient, as
the direction of the item vectors with respect to the coordinate axes
has a direct, meaningful interpretation.

## Generalized multidimensional parameters

The item location $`r MIL_PARAM`$ is defined as the distance and direction
from the origin to the point of maximum slope [@reckase_difficulty_1985].
By analogy with the unidimensional case, $`r MDISC_SYM`$ is defined as
$4 `r SLOPE_MAX_PARAM`$ [@reckase_discriminating_1991].
Therefore, we may define the $`r MIL_PARAM`$ and $`r MDISC_SYM`$ using
Equations \@ref(eq:distance-param) through \@ref(eq:dir-cos-vec-test-space).
However, as they depend on the inner product matrix $`r INNER_PROD_MATRIX`$,
the choice we make of this matrix will lead us to different definitions of
the parameters.

### Agnostic version

Up to this point,
we have derived the maximum slope and its location in the latent space
$`r LATENT_SPACE`$ independently of its basis $`r LS_BASIS`$.
If we assume that $`r LS_BASIS`$ is orthonormal, i.e., $`r BASIS_EQ`$,
$`r TRANSFORM_MATRIX`$ will also be orthonormal
and thus $`r INNER_PROD_MAT_STD_EQ`$.
In such a case the multidimensional item location and discrimination simplify
to the expressions derived by @reckase_difficulty_1985
and @reckase_discriminating_1991.
As these expressions implicitly assume the orthonormality of $`r LS_BASIS`$,
which bears no meaning besides its pure algebraic purpose,
we shall refer to them as the *agnostic* version of the parameters:

\begin{equation}
  \begin{split}
    `r MDISC_AG_PARAM_EQ` \\
    `r MIL_AG_PARAM_EQ`.
  \end{split}
\end{equation}

### Covariance-based version

On the other hand, we can let the geometrical representation of
the latent space $`r LATENT_SPACE`$ account for its covariance structure,
thus giving $`r LS_BASIS`$ a statistical meaning
in the context of our MIRT modeling approach.
To do that, we may assume that $`r TRAIT_COORDS`$ is
a random vector distributed with covariance $`r COV_MATRIX_EQ`$,
being $`r COV_MATRIX`$ an $`r N_DIMS`$-dimensional positive definite matrix,
$`r CORR_MATRIX`$ its corresponding correlation matrix
(also positive definite) and $`r SD_MATRIX`$ a scaling diagonal matrix
with $`r VAR_COV_EQ`$ (i.e., the variances).

To make the representation of $`r LATENT_SPACE`$ meaningful,
the axes in $`r LS_BASIS`$ must have
an interpretation in terms of the latent space structure.
Although we do not know, in principle, what
the interpretation of a latent space basis should be, we may assume that
it must be somewhat related to the covariance structure.
Therefore, an orthonormal basis should represent independent,
standard coordinates;
formally, if $`r COV_MATRIX_ORTH`$ is the covariance matrix of
$`r TRAIT_ORTH_COORDS`$, then $`r COV_MATRIX_ORTH_EQ`$.
Applying Equation \@ref(eq:change-basis), $`r COV_MATRIX_TRANSF_EQ`$, and thus,

\begin{equation}
  \begin{split}
    `r COV_MATRIX_TRANSF_ID_EQ` \\
    `r COV_MATRIX_RESULT_EQ`.
  \end{split}
\end{equation}

Therefore, we may define $`r TRANSFORM_MATRIX`$ such that
$`r INNER_PROD_COV_COND`$;
in this case, $`r DIAG_MATRIX_INNER_PROD_INV_VAR_EQ`$, resulting in a
*covariance-based* version of the parameters:

\begin{equation}
  \begin{split}
    `r MDISC_COV_PARAM_EQ` \\
    `r MIL_COV_PARAM_EQ`.
  \end{split}
\end{equation}

According to @reckase_discriminating_1991, the multidimensional parameters must
meet three conditions for being regarded as a valid generalization
of the (unidimensional) IRT parameters:

1. If an item measures only dimension $`r DIM_INDEX`$,
   then $`r MDISC_UNIDIM_EQ`$.

2. The distance $`r DISTANCE_PARAM`$ has the same relationship with the
   intercept as $b_i$ in the unidimensional case,
   i.e., $`r DIST_INTERCEPT_REL`$.

3. $`r MDISC_SYM`$ is four times the maximum slope $`r SLOPE_MAX_PARAM`$.

The second property derives straightforwardly from
Equation \@ref(eq:distance-param),
and the third one is implicit in the definition of the $`r MDISC_SYM`$.
However, when an item measures only dimension $`r DIM_INDEX`$,
simple arithmetic can show that $`r MDISC_UNIDIM_GENERALIZED_EQ`$.
In the covariance-based case, this means that $`r MDISC_COV_BASED_UNIDIM_EQ`$
and, as $`r COV_SYM_DIAG`$ is generally different from 1,
$`r MDISC_COV_PARAM`$ does not fulfill the first property
(we shall refer to this property as *scale invariance*, from now on).
Nevertheless, it can be interpreted as a *scaled version* of the parameter,
with the standard deviation of the corresponding dimension as scaling factor.
This transformation may be made in the unidimensional case as well,
thus finding a *standard-metric* discrimination parameter
(i.e., referred to a unitary variance space).

### Correlation-based version

Despite $`r MDISC_COV_PARAM`$ not being scale-invariant,
we have argued that it is still a valid generalization of the IRT parameters.
However, it may be convenient to find yet another generalization,
one that accounts for the latent space structure
while still fulfilling this property.
To do this, let's assume a matrix $`r TRANSFORM_MATRIX_AUX_EQ`$, such that
$`r TRANSFORM_MATRIX_AUX_RESULT`$.
Then, we have that

$$
  `r CORR_BASED_DERIVATION`.
$$

That is, defining $`r TRANSFORM_MATRIX`$ as
$`r TRANSFORM_MATRIX_CORR_BASED_DEF`$ can be interpreted as re-scaling
the transformed parameter back to its original metric.
In this case, we get that $`r INNER_PROD_CORR_COND`$.
Therefore, we can also define a *correlation-based* version of the
multidimensional parameters as

\begin{equation}
  \begin{split}
    `r MDISC_CORR_PARAM_EQ` \\
    `r MIL_CORR_PARAM_EQ`.
  \end{split}
\end{equation}

As we can see, this version also fulfills the scale invariance property,
as the $`r INNER_PROD_MAT_INV_DIAG_ELEMENT`$ elements are
the diagonal elements of $`r CORR_MATRIX`$, which are always equal to 1.
It is important to note though that this transformation does not have
an also desirable property,
which we have used to justify the covariance-based version:
The $`r TRAIT_ORTH_COORDS`$ coordinates resulting from transforming
$`r TRAIT_COORDS`$ to the orthonormal basis $`r LS_ORTH_BASIS`$
are not generally uncorrelated, as one would expect
(save some exceptions, e.g., when all the variances are equal).
Nevertheless, it is important to note how the $`r MDISC_CORR_PARAM`$ accounts
for the correlation structure of the latent space
while still being scale invariant.
This will be very useful for the purpose of representing the items graphically,
as we will see later.

## Vector representation

According to @ackerman_creating_1994, the most appropriate way of representing
multidimensional items is in vector form.
This allows representing several items altogether and analyzing them
in terms of their multidimensional parameters [@ackerman_graphical_1996].
An item vector will be applied at a direction and (signed) distance
from the origin of the coordinate system given by its $`r MIL_PARAM`$ parameter,
and will have a length equal to its $`r MDISC_SYM`$ parameter.
Its direction will also be given by the direction component of $`r MIL_PARAM`$,
so its orientation will always pass through the origin.
Based on these conditions, we need to compute
the origin and end coordinates of the vector,
denoted by $`r ORIGIN_ITEM_BASIS`$ and $`r END_ITEM_BASIS`$, respectively.

In an orthogonal basis it is easy to compute the coordinates,
as they are simply the orthogonal projections onto the corresponding axis.
In the oblique case it is not so simple though, but we can take advantage of the
orthornormalization introduced above.
In basis $`r TEST_SPACE_STD_BASIS`$, the origin and end coordinates are
$`r ORIGIN_ITEM_ORTH`$ and $`r END_ITEM_ORTH`$, respectively,
and they correspond to the coordinates in $`r TEST_SPACE_BASIS`$ transformed
according to Equation \@ref(eq:discr-vector-std),
i.e., $`r ORIGIN_TRANSF_EQ`$ and $`r END_TRANSF_EQ`$.
Given the mapping of $`r TEST_SPACE_STD_BASIS`$ onto $`r LS_ORTH_BASIS`$,
the direction cosines in $`r TEST_SPACE_STD_BASIS`$
are $`r DIR_COS_ITEM_VEC_STD`$.
Therefore,

\begin{align}
  `r ORIGIN_ORTH_EQ` \\
  `r END_ORTH_EQ`.
\end{align}

Applying Equation \@ref(eq:dir-cos-vec-std),
for the origin coordinates we have that

\begin{equation}
  `r ORIGIN_ORTH_DISCR_EQ`                              (\#eq:origin-orth-discr)
\end{equation}

which, operating, results in

\begin{align}
  `r ORIGIN_ITEM_COORDS_EQ`    \\
  `r ORIGIN_TRANSF_SQUARED_EQ` \\
  `r ORIGIN_INNERPROD_EQ`      \\
  `r ORIGIN_INNERPROD_DIAG_EQ`
\end{align}

and, finally,

\begin{equation}
  `r ORIGIN_INNERPROD_DIR_COS_EQ`.                   (\#eq:origin-coords-result)
\end{equation}

The end coordinates can be computed by a similar procedure, which results in

\begin{equation}
  `r END_ITEM_COORDS_EQ`.                               (\#eq:end-coords-result)
\end{equation}

Equations \@ref(eq:origin-coords-result) and \@ref(eq:end-coords-result)
express the items coordinates in terms of their multidimensional parameters.
However, it is worth noting that the latent space Gram matrix
also plays a role in computing these coordinates,
and will therefore have an effect on the representation.
Of course,
from Equations \@ref(eq:origin-orth-discr) and \@ref(eq:end-coords-result)
it is easy to express the coordinates in terms of the parameters in the original
model formulation, namely,

\begin{align}
  `r ORIGIN_ITEM_PARAMS_EQ`                  &(\#eq:origin-from-model-params) \\
  `r END_ITEM_PARAMS_EQ`.                    &(\#eq:end-from-model-params)
\end{align}

However, formulating them in terms of the multidimensional parameters allows us
studying the items in terms of their multidimensional properties.
In the following, we see how to geometrically interpret
these multidimensional parameters.

### Geometric properties of the items

We have seen how the two new versions of the parameters fulfill
the three conditions proposed by @reckase_discriminating_1991
for valid generalizations of the IRT parameters to the multidimensional case,
being the scale invariance of the $`r MDISC_COV_PARAM`$ the only exception.
From a geometrical perspective, the item $`r MDISC_SYM`$ is simply interpreted
as the length of its corresponding vector.
Regarding the $`r MIL_PARAM`$ parameter,
we will consider now two more properties, related to each of its two components:
the distance to the origin, and the measurement direction.

The (signed) distance to the origin,
given by the first component of $`r MIL_PARAM`$, has an interesting property:
By the definition of the inner product,
$`r MDISC_SYM`$ is strictly positive for non-null vectors.
Therefore, the sign of $`r DISTANCE_PARAM`$ is equal to
the sign of $`r INTERCEPT_PARAM`$.
The implication is that the item is displaced along
its measurement direction forwards or backwards with respect to the origin.
As the $`r MDISC_SYM`$ appears in the denominator of $`r DISTANCE_PARAM`$,
that displacement will be inversely proportional
to the discrimination of the item.
That is, the less discriminating the item is,
the further away it will be from the origin.

The measurement direction is given by the signs of $`r DIR_COS_ITEM_VEC`$ which,
also due to the strict positiveness of $`r MDISC_SYM`$,
are equal to the signs of $`r DISCR_VECTOR`$.
This property leads directly to generalizing
the parameters to the monotonically non-increasing case:
The measurement direction relative to dimension $`r DIM_INDEX`$
will be negative when $`r IRF_ABBR`$ is monotonically decreasing
with respect to variations along $`r TRAIT_COMPONENT`$.
When $`r IRF_ABBR`$ is constant
with respect to variations along $`r TRAIT_COMPONENT`$,
$`r SIGN_COS_VEC_ITEM_EQ`$,
which means that $`r MIL_PARAM`$ is orthogonal to the $`r DIM_INDEX`$-th axis.
However, note that this does not necessarily imply that
the vector is parallel to any other axis
(or strictly contained in the hyperplane formed by other axes, for that matter);
this will always be true for the $`r MIL_AG_PARAM`$ due to
the orthogonality assumption,
but it will depend on the correlation matrix $`r CORR_MATRIX`$
for the other two versions of the parameter.

## Graphical representation

The geometric properties are easier to apprehend with
a visual representation of the items.
To do that, we need to choose a version of the parameters
that yields their most faithful representation.
We have already seen that
the agnostic version disregards the latent space structure.
Therefore, it will only be useful in the unlikely case that
we have no information of such structure.
When we have an estimate of the latent space covariances,
we may use either the covariance-based or the correlation-based version.
However, we have already seen how
the distance to the origin depends on the $`r MDISC_SYM`$.
Also, in the covariance-based,
the term $`r INNER_PROD_MAT_DIAG_INV_PROD`$ becomes $`r COV_BASED_SCALE_MAT_EQ`$
which, as we have argued before, implies a change of scale.
In the case of the correlation-based version however,
$`r INNER_PROD_MAT_DIAG_INV_PROD`$ is simply $`r CORR_MATRIX_INV`$,
being the coordinates scale-invariant.
Therefore, for graphical representation purposes,
the correlation-based version will be preferred.

Plotting on a display usually requires providing
coordinates in a rectangular Cartesian system.
Therefore, we will need to compute the equivalent rectangular coordinates
of our general, possibly non-rectangular coordinates.
This implies pre-multiplying its coordinates in basis $`r TEST_SPACE_BASIS`$
by $`r TRANSFORM_MATRIX_TRANSP_INV`$, to find
the equivalent, orthonormalized coordinates.
Therefore, we will need to define
a convenient value for $`r TRANSFORM_MATRIX_TRANSP_INV`$
that allows us to represent the test space structure and the items
in a rectangular Cartesian system.

The general procedure for plotting an item $`r ITEM_INDEX`$ consists of
the following steps:

1. Compute $`r MDISC_CORR_PARAM_ITEM`$ and $`r MIL_CORR_PARAM_ITEM`$,
   the item $`r MDISC_CORR_PARAM`$ and $`r MIL_CORR_PARAM`$ parameters,
   respectively.

1. Compute the origin coordinates $`r ORIGIN_ITEM_EQ`$, with
   $`r DISTANCE_CORR_PARAM`$ and $`r DIR_CORR_PARAM`$ the distance and direction
   component, respectively, of $`r MIL_CORR_PARAM_ITEM`$.

1. Compute the end coordinates $`r END_ITEM_EQ`$.

1. Define $`r TRANSFORM_MATRIX_TRANSP_INV`$ such that
   $`r TRANSF_MATRIX_INV_SQ_CORR_EQ`$.

1. Compute the rectangular coordinates
   $`r ORIGIN_ITEM_ORTH`$ and $`r END_ITEM_ORTH`$ by pre-multiplying
   $`r ORIGIN_ITEM_BASIS`$ and $`r END_ITEM_BASIS`$, respectively, by
   the transformation matrix $`r TRANSFORM_MATRIX`$
   ($`r ORIGIN_TRANSF_EQ`$; $`r END_TRANSF_EQ`$).

Instead of steps 1 to 3,
one can alternatively compute $`r ORIGIN_ITEM_BASIS`$ and $`r END_ITEM_BASIS`$
directly from Equations \@ref(eq:origin-from-model-params)
and \@ref(eq:end-from-model-params).
Note though that these equations are not independent of $`r CORR_MATRIX`$,
as it is used in the computation of $`r DISCR_VECTOR_MODULE`$,
and thus this matrix is necessary in any case to compute the coordinates.

### Graphical representation example

```{r read-chunks-graphical-example}
read_chunk("src/Graphical_example_paper.R")
```

```{r libraries}
#| warning: false
```

```{r sources}
```

```{r constants}
```

```{r graphical-output-conf}
```

```{r compute-example-items}
```

As plotting more than two dimensions is difficult on a bidimensional display
(and hardly possible at all for more than three dimensions with the
currently available technology),
we will limit ourselves to the bidimensional case here.

The example in Figure \@ref(fig:item-plot-out) showcases a representation of
a set of items in two different bidimensional spaces.
In the first case, the two latent dimensions are independent,
with $`r TRANSFORM_MATRIX_TRANSP_INV`$ being an identity matrix.
In the second one, the correlation between the dimensions is $`r CORR_OBL_OUT`$.
To keep the horizontal axis invariant,
we define $`r TRANSFORM_MATRIX_TRANSP_INV`$ such that
it only rotates the vertical axis [@harman_modern_1970]:

\begin{equation}
  `r TRANSFORM_MATRIX_EXAMPLE_EQ`.
\end{equation}

The sample items have the same `r MODEL_ACRONYM` parameters
in both latent spaces.
However, their multidimensional parameters differ due to the correlation among
the latent dimensions.
The parameters in both spaces,
along with their corresponding `r MODEL_ACRONYM` parameters,
are shown in Table \@ref(tab:example-items-table-out).

```{r compose-example-items-table}
```

```{r set-flextable-wd}
#| cache: false

# Necessary for officedown to find the template file:
opts_knit$set(root.dir = DOC_DIR)
```

```{r example-items-table-out}
#| cache:   false
#| tab_id:  example-items-table-out
#| tab.cap: Item parameters for the graphical example
item_params_output
```

<br>

```{r compose-oblique-plot}
```

```{r compose-orthogonal-plot}
```

```{r item-plot-out}
#| cache:      false
#| fig.cap:   Item plots in an orthogonal (left) and oblique (right) test space.
#| fig.height: 3.1
#| fig.width:  6.65
plot_orth + plot_oblique
```

Figure \@ref(fig:item-plot-out) illustrates the effect of
a positive correlation on the $`r MDISC_CORR_PARAM`$
(and consequently the $`r MIL_CORR_PARAM`$) value:
The components of the discrimination parameters tend to
*sum up* in the same direction as they get *more aligned*.
Thus, if the discrimination parameters have the same sign, as in items 1 and 5,
the $`r MDISC_CORR_PARAM_ITEM`$ value tends to increase;
on the contrary, when the discrimination parameters have opposite signs
they tend to cancel each other out and $`r MDISC_CORR_PARAM_ITEM`$ decreases,
as happens with items 2 and 3.
This effect can be especially noticed by comparing
the $`r MDISC_CORR_PARAM_ITEM`$ of items 3 and 5:
Their discrimination parameters are equal in absolute value,
so their $`r MDISC_CORR_PARAM_ITEM`$ values are equal in the orthogonal space;
however, the effect of the correlation shrinks the former and
stretches the latter.
Finally, note that item 4 has one discrimination parameter equal to 0,
so its $`r MDISC_CORR_PARAM_ITEM`$ is unaffected by the correlation.

We can also see how the sign of $`r INTERCEPT_PARAM`$ affects
the distance to the origin:
The item vector is applied at the origin when its value is null (item 1),
and tends to be shifted
*against* the item direction when its value is positive (items 2 and 4) and
*towards* the item direction when it is negative (items 3 and 5).
Finally, we can see how the direction is determined by the discrimination
parameters *and* the correlation; the sign of the discrimination parameters
determines where the item will point at;
thus, an item with two positive (negative) discrimination parameters will point
at the first (third) quadrant,
while an item with opposite-sign discrimination parameters will be in the
second or fourth quadrant.
The direction relative to the axes will be given by the relative absolute value
of the two parameters, but also by the correlation:
Items 2 and 3 especially illustrate this effect,
as each of them is orthogonal to one of the axes in the oblique space,
even when they are not parallel to the other axis.

## Application to examples from the literature

In the following, we apply our results to two instances taken from
the MIRT literature.
The first one draws from a classic example by @reckase_multidimensional_2009,
who presents a three-dimensional instrument made up by 30 items.
It is supposed to be a cognitive test
(it is unclear whether the item parameters are estimated from
actual empirical data, or they are a fictional example,
created ad-hoc for simulation purposes),
so all the discrimination parameters are positive.
The second example, taken from @tezza_modelo_2018, is an application of
the `r MODEL_ACRONYM` to assess the quality of e-commerce websites.
Besides being an empirical example, applied to actual test data,
it pertains to the non-cognitive domain and contains
items with  negative discrimination parameters.
This is very convenient for exemplifying our results with
items with opposite-sign discrimination parameters.

The two examples are represented in Tables
\@ref(tab:empirical-table-Reckase2009-out)
and \@ref(tab:empirical-table-Tezza2018-out), respectively.
Each of these tables shows the original `r MODEL_ACRONYM` parameters
[the estimation errors in @tezza_modelo_2018 are omitted for clarity],
along with the agnostic version of the multidimensional parameters;
the $`r MDISC_AG_PARAM`$ and the $`r DISTANCE_AG_PARAM`$
are coincident with their respective original sources
[see Table 6.1 in @reckase_multidimensional_2009, p. 153, and
Table 5 in @tezza_modelo_2018, p. 926].
In the first case, the direction angles are also coincident with the values
provided in the original [@tezza_modelo_2018 do not provide these results].
In order to facilitate the comparison,
the covariance-based version of each parameter is provided in
the column next to its corresponding agnostic version.

### @reckase_multidimensional_2009

```{r reset-root-wd}
#| cache: false

# Necessary for knitr to find the source file:
opts_knit$set(root.dir = ROOT_DIR)
```

```{r read-chunks-empirical-example}
read_chunk("src/Empirical_example.R")
```

```{r libraries}
```

```{r sources}
```

```{r general-constants}
```

```{r set-cov-matrix-Reckase2009}
# Version in Reckase, 2009, p. 153; this represents a usual 3D case
COV_MATRIX_VALUE <- matrix(
  c(
    1.210, .297, 1.232,
     .297, .810,  .252,
    1.232, .252, 1.960
  ),
  nrow = 3
)
# (see also Reckase, 2009, p. 183 for a "close-to-1D" case).
```

```{r read-items}
```

```{r compute-orthogonal-params}
```

```{r compute-oblique-params}
```

```{r format-covariance-matrix}
```

```{r create-params-table}
```

```{r compose-output-table}
```

Table \@ref(tab:empirical-table-Reckase2009-out) shows the parameters of the
30-item test in the @reckase_multidimensional_2009 example:
There are three ten-item blocks assumed to approximate
a simple structure: Items 1-10 measure dimension 1, items 11-20 dimension 3,
and items 21-30 dimension 2
[see Table 6.1, p. 153, in @reckase_multidimensional_2009].
In his example, Reckase tests the estimation of these items
on two simulated datasets: In the first one, the latent trait distribution
is standard with null correlations; however, the second one has

\begin{equation}
  `r cov_matrix_value_out`,
\end{equation}

as covariance matrix, which corresponds to a correlation matrix of

\begin{equation}
  `r cor_matrix_value_out`.
\end{equation}

As we can see in Table \@ref(tab:empirical-table-Reckase2009-out),
the agnostic versions of the multidimensional parameters
(columns labelled _`r AGNOSTIC_ABBR`_)
coincide with Reckase's columns A, B, and $\alpha_1$ to $\alpha_3$.
However, when we consider the latent space structure
(columns labelled $`r COV_MATRIX`$), we notice several differences.
First, the $`r MDISC_ITEM`$ parameters tend to be larger
than their agnostic counterparts.
This happens because all the discrimination parameters and the latent space
correlations are positive; in this respect, these items are similar to item 1
in Table \@ref(tab:example-items-table-out) and Figure \@ref(fig:item-plot-out).
However, for the items mostly aligned with dimension 2 (items 21-30), as the
$`r MDISC_ITEM`$ parameters are also scaled by a standard deviation smaller
than 1, their values decrease, being some of them smaller than the agnostic
versions. The $`r DISTANCE_PARAM`$ parameters decrease or increase, consequently
(see Equation \@ref(eq:M2PL-formula-polar)), thus being most of them smaller
than the agnostic ones (again, with the exceptions only happening
in the items aligned with dimension 2).
Finally, we see that the direction angles are not as clearly separated as one
might expect from the agnostic versions. This effect is more pronounced among
the items aligned with dimensions 1 and 3, which are strongly correlated.
In this situation, we notice that the angles tend to be much smaller, as the
high correlation induces an alignment between the two dimensions
(compare item 1 in the two test spaces in
Table \@ref(tab:example-items-table-out) and Figure \@ref(fig:item-plot-out)).

```{r set-flextable-wd}
#| cache: false
```

```{r empirical-table-Reckase2009-out}
#| cache:   false
#| tab.id:  empirical-table-Reckase2009-out
#| tab.cap: |
#|   Agnostic and covariance-based multidimensional item parameters in
#|   Reckase (2009), with a (rank-complete) three-dimensional covariance matrix.
item_params_output
```

The latter effect can be better noticed in another case, in which Reckase
uses the same item parameters to demonstrate the over-specification of
dimensions (see p. 183).
In that example, he uses a covariance matrix that practically has rank 1,
implying that the (alleged) three dimensions collapse into a single one, being
thus the test practically unidimensional.
In such a case, the covariance-based direction angles
are extremely close to 0 for all items, as one would expect
from a unidimensional test, where all items are strictly aligned with the one
dimension they measure. The reader interested can check out the parameters
computed in that case in the table in the Supplementary Material.

### @tezza_modelo_2018

```{r set-cov-matrix-Tezza2018}
# See Tezza et al., 2018, p. 927:
# > No presente estudo, a correlação entre a dimensão 1 e a dimensão 4
# > foi de aproximadamente 0,4.
COV_MATRIX_VALUE <- diag(4)
DIMS_CORR        <- .4
COV_MATRIX_VALUE[1, 4] <- COV_MATRIX_VALUE[4, 1] <- DIMS_CORR
```

```{r reset-root-wd}
#| cache: false
```

```{r read-items-Tezza2018}
items <- read_csv2(ITEM_PARAMS_TEZZA_2018_FILEPATH)
```

```{r compute-orthogonal-params}
```

```{r compare-orthogonal-params}
#| results: hide
```

<!-- Computed item parameters in the orthonormal space match the ones in given
in Tezza 2018, table 5 -->

```{r compute-oblique-params}
```

```{r format-covariance-matrix}
```

```{r create-params-table}
```

```{r compose-output-table}
```

@tezza_modelo_2018 provide us with another example that allows showcasing
the effect of correlated dimensions on negative discrimination parameters.
In their model, dimensions 1 and 4 are estimated to have a correlation of
`r DIMS_CORR`.
With such latent space structure, items that strongly discriminate in any or
both of these two dimensions change drastically their multidimensional
parameters from the agnostic to the covariance-based version
(see Table \@ref(tab:empirical-table-Tezza2018-out)) ---this happens,
for example, with items 57 and 60. These items also have different-sign
discrimination parameters in dimensions 1 and 4, so their situation is
similar to items 2 and 3 in Table \@ref(tab:example-items-table-out) and
Figure \@ref(fig:item-plot-out),
thus the decrease in their $`r MDISC_ITEM`$ parameter.
The effect on the direction is also patent in the angles these items make with
these dimensions, which change up to almost 20º, compared with at most 7º in the
other two dimensions, and around 1º in most cases.

On the other hand, in cases such as items 12 or 59, which have
relatively low discrimination parameters on those two dimensions
(compared with the ones in dimension 2 and 3, respectively) are barely affected.
The most extreme change in these two examples occurs in
the direction of item 12 with respect to dimension 4. We see how the large
discrimination in dimension 1, when combined with the correlation
with dimension 4, induces a decrease in the angular direction
with that dimension.
In this case, being the discrimination in dimension 4 so low, the situation is
very similar to item 4 in Table \@ref(tab:example-items-table-out)
and Figure \@ref(fig:item-plot-out).

<!---BLOCK_LANDSCAPE_START--->

```{r set-flextable-wd}
#| cache: false
```

```{r empirical-table-Tezza2018-out}
#| cache:   false
#| tab.id:  empirical-table-Tezza2018-out
#| tab.cap: |
#|   Agnostic and covariance-based multidimensional item parameters in
#|   Tezza et al. (2018).
item_params_output
```

<!---BLOCK_LANDSCAPE_STOP--->

# Discussion

@reckase_difficulty_1985 proposed a definition of the $`r MIL_PARAM`$ parameter
general enough to "be used with any model that yields probabilities that
increase monotonically with an increase in ability on any dimension" (p. 411).
However, we have discussed that this definition implicitly assumes
the orthonormality of the latent space.
Applying the general procedure outlined by @reckase_difficulty_1985 and later
extended by @reckase_discriminating_1991
we have obtained a set of multidimensional parameters
that generalize the original results in two aspects:
(1) to a non-orthonormal space that accounts for the covariance structure of
the multivariate latent variable, and (2) to any case of unchanging monotony.
It is worth noting that @zhang_theoretical_1999 already proposed a formulation
of the $`r MDISC_SYM`$ equivalent to our covariance-based version,
but for a more general, semi-parametric formulation
(although constrained to non-decreasing monotony).
However, they provided no formal proof for this result.
Our results provide this formal derivation, although only for the specific
case of the `r MODEL_ACRONYM` model.

As we have seen, taking the latent space structure into account may have a
substantial effect in the resulting values and interpretation of the parameters,
something not contemplated by the original
(hereby referred to as agnostic) version.
Moreover, in applications where the latent space structure is critical
(e.g., multi-group IRT, equating, and linking),
the agnostic version may overlook relevant differences in
the values of the multidimensional parameters.
Thus, the covariance structure of the latent space should be considered
when computing the multidimensional parameters of a MIRT model.

With that purpose in mind, we have defined two versions of these parameters:
One that takes into account the whole covariance structure,
and another one that only considers the correlations among latent dimensions.
The covariance-based parameter has the drawback of not fulfilling
the first of @reckase_discriminating_1991's properties, namely,
the scale invariance property.
However, we have provided a rationale for the violation of this property,
arguing that it is also paralleled in unidimensional IRT.
Given its equivalence to the inner product definition in a space with
an identity covariance matrix,
it also refers the item parameters to an orthonormal metric,
comparable among different latent spaces.
More importantly, the covariance-based version
of the multidimensional parameters is invariant to
changes of latent-space basis.
Once again, this is important in applications where
potentially different bases are involved.
Given these considerations,
we propose to adopt the covariance-based version of
the multidimensional parameters as the most general definition,
and refer to them simply as
the $`r MDISC_SYM`$ and $`r MIL_PARAM`$ parameters.
The correlation-based version, on the other hand,
has the advantage of being scale-invariant,
which is convenient for fidelity when representing the items.
However, it is worth noting that it has been defined ad-hoc
to fulfill this specific property.
Therefore, it lacks other desirable properties of the covariance-based version,
like being invariant to a latent space scaling transformation,
or yielding a diagonal latent covariance matrix in the corresponding
orthogonal latent space basis.
Therefore, we recommend its use for graphical representation purposes only.
As for the agnostic version, we recommend it only in
the (unlikely) case of having no information about the latent space structure.

One may consider that the formal derivations presented here are
overly complicated,
when one might simply transform the space to rectangular Cartesian coordinates
with Equations \@ref(eq:change-basis) and \@ref(eq:discr-vector-std) and
then compute the parameters in the transformed space using
@reckase_difficulty_1985's and @reckase_discriminating_1991's procedure.
After all, interpreting the axes of this space as substantive latent traits
depends on "the distinction between coordinate axes and the constructs that are
the target of the instrument"
(Reckase, M.D., personal communication, 28 April 2015).
Such a procedure will of course yield identical results,
given the invariance of the parameters to a change of basis.
However, the computations in the oblique space allow us to obtain
the direction cosines relative to the original latent dimensions,
which may give relevant information if these are actual, substantive traits.
Computing the parameters in the orthonormalized (or just orthogonalized) space,
by contrast, would require computing the cosines in
the original basis afterwards.

On the other hand, establishing the relationship between the original and
the orthonormalized basis has provided us with an important insight:
The distinction between the latent space and the test space.
Our derivations show that, to properly represent the items,
we need to define a vector space that has
the latent covariance (or correlation) matrix as its inner product matrix.
On the other hand, the latent space,
where the person parameters are represented,
has the inverse of the covariance (correlation) matrix as
its inner product matrix.
This result is both surprising and counterintuitive;
The MIRT literature suggests in several instances that
"the angle between the axes [of latent ability space] represents
the degree of correlation" [@ackerman_multidimensional_2005-1, p. 16].
We hope our results help shed a clearer light on
this often misunderstood topic.

It is worth noting that a similar relationship exists in
the common factor literature
between the *primary factors* and Thurstone's *reference axes*
[@harman_modern_1970, see p. 291].
Nevertheless, one must not make the mistake of assimilating the test space and
the *common-factor space*.
Their parameterization, albeit related [@mcdonald_test_1999], is different,
and we do not mean to imply that the derivations made here are directly
generalizable to the common factor model.
Those relationships deserve further exploration on their own.

Finally, we must also warn against generalizing these results
to other IRT models without further consideration.
First, we have only considered the case of dichotomous items,
whereas non-cognitive applications often (and sometimes cognitive ones as well)
require modeling several response categories.
Although proposals such as the graded-scale [@samejima_estimation_1968] and
the nominal response [@bock_estimating_1972] model are closely related to
the 2-parameter logistic model, it would be hasty to assume that these
parameters generalize in any way to
the multidimensional versions of those models without a formal proof.
The same can be said from the multidimensional parameters of other models
for dichotomous items different from the `r MODEL_ACRONYM` model.
We expect this work to help set the foundations for investigating those
generalizations.
Hopefully, this will contribute to a better understanding
of the underlying multidimensional measurement theory, its implications,
and its interpretation.

\newpage

# References