### 2024-09-03 output/paper_draft/Supplementary_material.Rmd
Remove author and date from supplementary material

---
title: "Supplementary material"
output:
  officedown::rdocx_document:
    tables:
      caption:
        pre: "Table S."
      style: APA_Like_style
    mapstyles:
      Normal: First Paragraph
    base_format: bookdown::word_document2
    reference_docx: ../../www/Template.docx
    number_sections: no
    keep_md:         no
    fig_width:       4
    fig_height:      4
bibliography: [../../www/Multidimensional-parameters-MCLM.bib, ../../www/packages.bib]
csl:          ../../www/apa-old-doi-prefix.csl
editor_options: 
  chunk_output_type: console
---

```{r setup}
#| include: false

# Includes:

library(knitr)
library(patchwork)
library(rstudioapi)

ROOT_DIR <- getActiveProject()
DOC_DIR  <- getwd()

opts_knit$set(root.dir = ROOT_DIR)
opts_chunk$set(
  echo     = FALSE,
  message  = FALSE,
  warning  = FALSE,
  cache    = FALSE
)

# Output configuration options:
options(digits = 3)
```

```{r reset-root-wd}
#| cache: false

# Necessary for knitr to find the source file:
opts_knit$set(root.dir = ROOT_DIR)
```

```{r read-chunks-empirical-example}
read_chunk("src/Empirical_example.R")
```

```{r libraries}
```

```{r sources}
```

```{r general-constants}
```

```{r set-cov-matrix}
```

```{r read-items}
```

```{r compute-orthogonal-params}
```

```{r compute-oblique-params}
```

```{r format-covariance-matrix}
```

```{r create-params-table}
```

```{r compose-output-table}
```

Table S.\@ref(tab:empirical-table-Reckase2009-out) is similar to Table 2,
comparing the agnostic version of the multidimensional parameters drawn from
@reckase_multidimensional_2009 with the covariance-based version.
In this case though, we use the almost rank-1 covariance matrix Reckase defines
for representing a *close-to-unidimensional* case, that is,

\begin{equation}
  `r cov_matrix_value_out`,
\end{equation}

which yields the correlation matrix

\begin{equation}
  `r cor_matrix_value_out`.
\end{equation}

As in Table 2, the agnostic versions (columns labelled _`r AGNOSTIC_ABBR`_)
coincide with Reckase's columns A, B, and $\alpha_1$ to $\alpha_3$.
However, the covariance-based versions (columns labelled $`r COV_MATRIX`$)
have much smaller values for the $`r MDISC_ITEM`$ parameters.
This is primarily due to the scaling effect induced by the variances,
which are all smaller than 1.
Even though the discrimination parameters almost sum up to yield
the $`r MDISC_ITEM`$ value, due to the very high alignment of the dimensions,
the scaling effect of the variances has a stronger weight.
The $`r DISTANCE_PARAM`$ parameter values decrease consequently,
as we have seen before.

More interesting is the effect on the direction angles:
We can see how their values are all very close to 0.
This result is as expected,
given that all the three dimensions are highly aligned,
and therefore they almost represent a single dimension.
The items thus all measure in a direction very closely aligned to the dominant
dimension made up by those three, being almost parallel to all the three
dimensions.
Therefore, we see how computing the more appropriate covariance-based version of
the multidimensional parameters can help diagnose correctly
the dimensionality of a test.

```{r set-flextable-wd}
#| cache: false
```

# References

::: {#refs}
:::

\newpage

```{r empirical-table-Reckase2009-out}
#| cache:   false
#| tab.id:  empirical-table-Reckase2009-out
#| tab.cap: |
#|   Agnostic and covariance-based multidimensional item parameters in
#|   Reckase (2009), with an almost-rank-1 covariance matrix.
item_params_output
```