### 2024-09-05 output/Supplementary_material.Rmd
Updates build script
The build script now generates all output files. Also, the files can be rendered with the automating script, not depending on the Rstudio active session which would cause an exception in a non-interactive session

---
title: "Supplementary material"
output:
  officedown::rdocx_document:
    tables:
      caption:
        pre: "Table S."
      style: APA_Like_style
    mapstyles:
      Body Text: First Paragraph
    base_format: bookdown::word_document2
    reference_docx: ../www/Template.docx
    number_sections: no
    keep_md:         no
    fig_width:       4
    fig_height:      4
bibliography: [../www/Multidimensional-parameters-MCLM.bib, ../www/packages.bib]
csl:          ../www/apa-old-doi-prefix.csl
editor_options: 
  chunk_output_type: console
---

```{r setup}
#| include: false

# Includes:

library(here)
library(knitr)
library(patchwork)
library(rstudioapi)

ROOT_DIR <- here()

opts_knit$set(root.dir = ROOT_DIR)
opts_chunk$set(
  echo     = FALSE,
  message  = FALSE,
  warning  = FALSE,
  cache    = FALSE
)

# Output configuration options:
options(digits = 3)
```

```{r reset-root-wd}
#| cache: false

# Necessary for knitr to find the source file:
opts_knit$set(root.dir = ROOT_DIR)
```

```{r read-chunks-empirical-example}
read_chunk("src/Empirical_example.R")
```

```{r libraries}
```

```{r sources}
```

```{r general-constants}
```

```{r set-cov-matrix}
```

```{r read-items}
```

```{r compute-orthogonal-params}
```

```{r compute-oblique-params}
```

```{r format-covariance-matrix}
```

```{r create-params-table}
```

```{r compose-output-table}
```

Table S.\@ref(tab:empirical-table-Reckase2009-out) is similar to Table 2,
comparing the agnostic version of the multidimensional parameters drawn from
@reckase_multidimensional_2009 [p. 183] with the covariance-based version.
In this case though, we use the almost rank-1 covariance matrix Reckase defines
for representing a *close-to-unidimensional* case, that is,

\begin{equation}
  `r cov_matrix_value_out`,
\end{equation}

which yields the correlation matrix

\begin{equation}
  `r cor_matrix_value_out`.
\end{equation}

As in Table 2, the agnostic versions (columns labelled _`r AGNOSTIC_ABBR`_)
coincide with Reckase's columns A, B, and $\alpha_1$ to $\alpha_3$.
However, the covariance-based versions (columns labelled $`r COV_MATRIX`$)
have much smaller values for the $`r MDISC_ITEM`$ parameters.
This is primarily due to the scaling effect induced by the variances,
which are all smaller than 1.
Even though the discrimination parameters almost sum up to yield
the $`r MDISC_ITEM`$ value, due to the very high alignment of the dimensions,
the scaling effect of the variances has a stronger weight.
The $`r DISTANCE_PARAM`$ parameter values decrease consequently,
as we have seen before.

More interesting is the effect on the direction angles:
We can see how their values are all very close to 0.
This result is as expected,
given that all the three dimensions are highly aligned,
and therefore they almost represent a single dimension.
The items thus all measure in a direction very closely aligned to the dominant
dimension made up by those three, being almost parallel to all the three
dimensions.
Therefore, we see how computing the more appropriate covariance-based version of
the multidimensional parameters can help diagnose correctly
the dimensionality of a test.

# References

::: {#refs}
:::

\newpage

```{r empirical-table-Reckase2009-out}
#| cache:   false
#| tab.id:  empirical-table-Reckase2009-out
#| tab.cap: |
#|   Agnostic and covariance-based multidimensional item parameters in
#|   Reckase (2009), with an almost-rank-1 covariance matrix.
item_params_output
```