### 2024-09-16 output/Title_page.Rmd

---
output:
  officedown::rdocx_document:
    mapstyles:
      First Paragraph: Body Text
    base_format: bookdown::word_document2
    reference_docx: ../www/Template.docx
    number_sections: no
    keep_md:         no
editor_options: 
  chunk_output_type: console
---

**Article title:**
A Generalized Definition of Multidimensional Item Response Theory Parameters

**Running title:** Generalized MIRT parameters

# List of authors

- Daniel Morillo-Cuadrado. Department of Behavioral Science Methodology,
  Faculty of Psychology, Universidad Nacional de Educación a Distancia (UNED),
  Spain.

- Mario Luzardo-Verde. Department of Mathematics and Applications,
  Universidad de La República, Uruguay.

## Corresponding author:

Daniel Morillo-Cuadrado \
Faculty of Psychology, Office 2.74. \
C. de Juan del Rosal 10, 28040 Madrid

Phone nº: +34 91398 6584 \
Email address: dvmorillo@psi.uned.es

# Financial Support

This research received no specific grant funding form any funding agency,
commercial or not-for-profit sectors.

# Competing Interests Statement

Competing interests: The author(s) declare none.