### 2025-07-09 output/sandbox_slideshow.qmd
Update sandbox slideshow to use with IMPS slides

---
bibliography:  ../www/Multidimensional-parameters-MCLM.bib
csl:           ../www/apa-old-doi-prefix.csl
date-meta:     "2024-11-27"
editor:
  mode:        source
  markdown: 
    wrap:      sentence
    canonical: true
lang:          es
knitr:
  opts_chunk: 
    results:   asis
    cache:     false
    autodep:   true
format:
  revealjs:
    auto-stretch:            true
    code-annotations:        hover
    df-print:                paged
    fig-cap-location:        bottom
    fig-format:              svg
    fig-asp:                 1
    fig-width:               6
    keep-md:                 true
    link-external-newwindow: true
    margin:                  0
    self-contained:          true
    slide-number:            false
    theme:                   ../www/slideshow-assets/extra-styles.scss
editor_options: 
  chunk_output_type: console
---

```{r libraries}
#| cache:   false
#| include: false

library(tibble)
library(dplyr)
library(knitr)
library(lavaan)
library(semPlot)
library(extrafont)
library(tidyr)
library(ggplot2)
library(units)
library(ggtext)
library(ggridges)
library(qgraph)
library(stringr)
library(purrr)
library(plotly)
library(metR)
library(scales)
library(fontawesome)
library(mvtnorm)
library(patchwork)
```

```{r setup}
#| cache:   false
#| include: false
opts_knit$set(root.dir = here::here()) # See lines 43-44

# Graphical configuration options:
loadfonts(device = "win") # Load fonts (if necessary)
par(family = "Arial", adj = 0.5) # Set graphical device params
```

```{r sources}
source("R/Formulae.R",                  encoding = 'UTF-8')
source("R/Output.R",                    encoding = 'UTF-8')
source("R/Mirt_toolbox.R",              encoding = 'UTF-8')
source("src/Graphical_example_paper.R", encoding = 'UTF-8')
```

```{r graphical-constants}
# Color palettes:

PALETTE_UNED <- c(
  normal = "#00533e",
  medium = "#427562",
  light  = "#86a699",
  white  = "#cddad5"
)
PALETTE_APPLE <- c(
  normal = "#749f4c",
  medium = "#8aac5d",
  light  = "#a2bc7e",
  white  = "#cddbb8"
)
PALETTE_BLUE <- c(
  normal = "#5c6eb1",
  medium = "#7683bd",
  light  = "#929acb",
  white  = "#d4d6eb"
)
PALETTE_TANGERINE <- c(
  normal = "#d76f47",
  medium = "#dd8964",
  light  = "#e4a482",
  white  = "#f0cfb9"
)
PALETTE_STRAWBERRY <- c(
  normal = "#da5268",
  medium = "#c87384",
  light  = "#daa6b1",
  white  = "#efdbdf"
)
PALETTE_RASPBERRY <- c(
  normal = "#90214a",
  medium = "#8e4d60",
  light  = "#af848b",
  white  = "#dac7c7"
)

# Factor axis colors:
F1_COLOR <- unname(PALETTE_BLUE["normal"])
F2_COLOR <- unname(PALETTE_RASPBERRY["normal"])

# Grid breaks:
X_BREAKS <- seq(-3, 3, length.out = 5)

# Probability axes:
PROB_AXIS <- seq(-3.3, 3.3, by = .01)
PROB_AXIS_2D <- seq(-3, 3, by = 0.05) # Axis for the probability contour plots
```

```{r graphical-output-conf}
#| cache: false

# Graphical output constants:
GRAPH_FONT   <- "Arial"  # Replace default font to comply with UNED identity
FONT_SIZE    <- 24       # Replace default font size for slideshow
AXIS_LAB_POS <-  0.5     # Replace default axis label position for slideshow
LINE_WIDTH   <-  0.75    # Replace default line width for slideshow
GRID_WIDTH   <-  0.5     # Width for grid and axis lines
AXIS_COLOR   <- "gray70" # Replace default grid line color for slideshow
VECTOR_WIDTH <- 1.5      # Replace default vector width for slideshow

theme_set( # `ggplot` output configuration
  theme_classic(
    base_size      = FONT_SIZE,
    base_family    = GRAPH_FONT,
    base_line_size = LINE_WIDTH
  ) %+replace%
    theme(
      axis.title.x       = element_text(hjust = AXIS_LAB_POS),
      axis.title.y       = element_text(vjust = AXIS_LAB_POS, angle = 0),
      axis.title.y.right = element_text(vjust = AXIS_LAB_POS, angle = 0),
      panel.grid         = element_line(color = AXIS_COLOR),
      panel.grid.minor   = element_line(linetype = "17", color = AXIS_COLOR)
    )
)

MARGINS <- c(12, 1.3, 8, 1.3) # margins for the sem path plots
```

## References