### 2025-07-14 output/IMPS-2025-slideshow.qmd
Delete old "orthogonality assumption" slides

---
title:     |
  A generalization of multidimensional item response theory parameters
author:    |
  <table>
    <tr>
      <td>Daniel Morillo, Ph.D.</td>
      <td>Mario Luzardo, Ph.D.</td>
    </tr>
    <tr>
      <td>
        <center>
          [![](../www/slideshow-assets/github-logo.png){height="50"}](https://github.com/DaniMori/)&emsp;
          [![](../www/slideshow-assets/orcid.png){height="50"}](https://orcid.org/0000-0003-3021-3878)
        </center>
      </td>
      <td>
        <center>
          [![](../www/slideshow-assets/github-logo.png){height="50"}](https://github.com/mluzardov/)&emsp;
          [![](../www/slideshow-assets/orcid.png){height="50"}](https://orcid.org/0000-0002-9360-2806)
        </center>
      </td>
    </tr>
  </table>
institute: |
  <table>
    <!-- # TODO: Fix vertical alignment in logos-->
    <tr>
      <td style="width: 400px">
        [![](../www/slideshow-assets/logos_fac_psicologia.svg){height="60"}](https://www.uned.es/universidad/facultades/psicologia.html)
      </td>
      <td style="width: 100px">
        [![](../www/slideshow-assets/logo_udelar.svg){height="120"}](https://udelar.edu.uy)
      </td>
    </tr>
  </table>
bibliography:  ../www/Multidimensional-parameters-MCLM.bib
csl:           ../www/apa-old-doi-prefix.csl
date:          "2024-07-17"
date-meta:     "2024-11-27"
date-format:   long
editor:
  mode:        source
  markdown: 
    wrap:      sentence
    canonical: true # TODO: Restore
execute: 
  cache:       true
lang:          en
knitr:
  # opts_knit: # Does not work, see line 102
  #   root_dir:  here::here()
  opts_chunk: 
    results:   asis
    autodep:   true
format:
  revealjs:
    auto-stretch:            true
    code-annotations:        hover
    df-print:                paged
    fig-cap-location:        bottom
    fig-format:              svg
    fig-asp:                 1
    fig-width:               6
    incremental:             false
    keep-md:                 false
    link-external-newwindow: true
    margin:                  0
    self-contained:          true
    slide-number:            false
    theme:                   ../www/slideshow-assets/extra-styles.scss
    transition:              none
    view-distance:           3
    template-partials:
      - ../www/slideshow-assets/title-slide.html
editor_options:
  chunk_output_type: console
---

```{r libraries}
#| cache:   false
#| include: false

library(tibble)
library(dplyr)
library(knitr)
library(lavaan)
library(semPlot)
library(extrafont)
library(tidyr)
library(ggplot2)
library(units)
library(ggtext)
library(ggridges)
library(qgraph)
library(stringr)
library(purrr)
library(plotly)
library(metR)
library(scales)
library(fontawesome)
library(mvtnorm)
library(patchwork)
```

```{r setup}
#| cache:   false
#| include: false
opts_knit$set(root.dir = here::here()) # See lines 43-44

# Graphical configuration options:
loadfonts(device = "win") # Load fonts (if necessary)
par(family = "Arial", adj = 0.5) # Set graphical device params
```

```{r sources}
source("R/Formulae.R",                  encoding = 'UTF-8')
source("R/Output.R",                    encoding = 'UTF-8')
source("R/Mirt_toolbox.R",              encoding = 'UTF-8')
source("src/Graphical_example_paper.R", encoding = 'UTF-8')
```

```{r graphical-constants}
# Color palettes:

PALETTE_UNED <- c(
  normal = "#00533e",
  medium = "#427562",
  light  = "#86a699",
  white  = "#cddad5"
)
PALETTE_APPLE <- c(
  normal = "#749f4c",
  medium = "#8aac5d",
  light  = "#a2bc7e",
  white  = "#cddbb8"
)
PALETTE_BLUE <- c(
  normal = "#5c6eb1",
  medium = "#7683bd",
  light  = "#929acb",
  white  = "#d4d6eb"
)
PALETTE_TANGERINE <- c(
  normal = "#d76f47",
  medium = "#dd8964",
  light  = "#e4a482",
  white  = "#f0cfb9"
)
PALETTE_STRAWBERRY <- c(
  normal = "#da5268",
  medium = "#c87384",
  light  = "#daa6b1",
  white  = "#efdbdf"
)
PALETTE_RASPBERRY <- c(
  normal = "#90214a",
  medium = "#8e4d60",
  light  = "#af848b",
  white  = "#dac7c7"
)

# Factor axis colors:
F1_COLOR <- unname(PALETTE_BLUE["normal"])
F2_COLOR <- unname(PALETTE_RASPBERRY["normal"])

# Item colors:
ITEMS_PALETTE <- c(
  PALETTE_BLUE["normal"],
  PALETTE_RASPBERRY["normal"],
  PALETTE_APPLE["normal"],
  PALETTE_TANGERINE["normal"],
  PALETTE_UNED["medium"]
) |>
  unname() |>
  setNames(1:5)

# Grid breaks:
X_BREAKS <- seq(-3, 3, length.out = 5)

# Probability axes:
PROB_AXIS <- seq(-3.3, 3.3, by = .01)
PROB_AXIS_2D <- seq(-3, 3, by = 0.05) # Axis for the probability contour plots
```

```{r graphical-output-conf}
#| cache: false

# Graphical output constants:
GRAPH_FONT   <- "Arial"  # Replace default font to comply with UNED identity
FONT_SIZE    <- 24       # Replace default font size for slideshow
AXIS_LAB_POS <-  0.5     # Replace default axis label position for slideshow
LINE_WIDTH   <-  0.75    # Replace default line width for slideshow
GRID_WIDTH   <-  0.5     # Width for grid and axis lines
AXIS_COLOR   <- "gray70" # Replace default grid line color for slideshow
VECTOR_WIDTH <- 1.5      # Replace default vector width for slideshow

theme_set( # `ggplot` output configuration
  theme_classic(
    base_size      = FONT_SIZE,
    base_family    = GRAPH_FONT,
    base_line_size = LINE_WIDTH
  ) %+replace%
    theme(
      axis.title.x       = element_text(hjust = AXIS_LAB_POS),
      axis.title.y       = element_text(vjust = AXIS_LAB_POS, angle = 0),
      axis.title.y.right = element_text(vjust = AXIS_LAB_POS, angle = 0),
      panel.grid         = element_line(color = AXIS_COLOR),
      panel.grid.minor   = element_line(linetype = "17", color = AXIS_COLOR)
    )
)

MARGINS <- c(12, 1.3, 8, 1.3) # margins for the sem path plots
```

## Multidimensional IRT {.smaller}

-   Multidimensional 2PL (M2PL) model [@mckinley_extension_1983]:

:::::: columns
:::: column
$$
  P(Y_i = 1) = \Phi_L (\mathbf{a}_i^T \mathbf{\theta} + d_i)
$$

::: {.fragment fragment-index="1"}
```{r mirt-model-irs-3d}
#| fig-width:  5
#| fig-height: 5
mirt_item <- tribble(
  ~item, ~a_1, ~a_2, ~d,
  1,     1,    0.5,  -0.5
)

irs2d <- function(a_1, a_2, d, theta_1, theta_2) {

  logit(a_1 * theta_1 + a_2 * theta_2 + d)
}

mirt_item_probs <- expand_grid(
  trait_2 = PROB_AXIS_2D,
  trait_1 = PROB_AXIS_2D
) |>
  mutate(
    mirt_item |>
      select(-item) |>
      pmap(irs2d, theta_1 = trait_1, theta_2 = trait_2) |>
      bind_cols()
  ) |>
    rename(prob = `...1`)

mirt_prob_surface <- mirt_item_probs |>
  pivot_wider(names_from = trait_1, values_from = prob) |>
    select(-trait_2) |>
    as.matrix()

plot_ly(
  x = ~PROB_AXIS_2D,
  y = ~PROB_AXIS_2D,
  z = ~mirt_prob_surface,
  colorscale = list(c(0, 1), PALETTE_BLUE["normal"] |> rep(2)),
  type = 'surface',
  hovertemplate = 'P(I<sub>i</sub> = 1) = %{z:.2f}<extra></extra>',
  showscale = FALSE,
  hoverinfo = 'Z'
) |>
  layout(
    scene = list(
      xaxis = list(title = "<i>θ</i><sub>1</sub>"),
      yaxis = list(title = "<i>θ</i><sub>2</sub>"),
      zaxis = list(title = "P(I<sub>i</sub> = 1)"),
      camera = list(
        center = list(x = -.1, z = -.2),
        eye = list(x = -.12, y = -1.8, z = 1)
      )
    )
  ) |>
  config(displayModeBar = FALSE, scrollZoom = FALSE)
```
:::
::::

::: {.column .fragment fragment-index="1"}
<br>

```{r mirt-model-irs-contour}
contour_plot <- mirt_item_probs |>
  ggplot(mapping = aes(trait_1, trait_2, z = prob, label = after_stat(level))) +
  geom_vline(xintercept = 0, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0, linewidth = GRID_WIDTH) +
  geom_contour2(
    color = F1_COLOR,
    linewidth = LINE_WIDTH,
    breaks = c(.05, .1, .2, .3, .5, .7, .8, .9, .95),
    label.placer = label_placer_fraction(frac = .90, rot_adjuster = 0)
  ) +
  scale_x_continuous(
    minor_breaks = NULL,
    limits = range(PROB_AXIS_2D),
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "*&theta;*~2~", labels = NULL)
  ) +
  scale_y_continuous(
    minor_breaks = NULL,
    limits = range(PROB_AXIS_2D),
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "*&theta;*~1~", labels = NULL)
  ) +
  coord_fixed() +
  theme(
    axis.ticks       = element_blank(),
    axis.line        = element_blank(),
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(),
    panel.grid.major = element_line(linetype = "17", linewidth = GRID_WIDTH)
  )

contour_plot
```
:::
::::::

::: notes
The Multidimensional 2PL model is one of many extensions of the 2PL model.

It assumes that the probability of a correct response to an item is a function of the linear combination of the latent traits weighted by item discrimination parameters.
(This here is a vector inner product; the "T" means that the discrimination vector is transposed).

And it has this additional "intercept parameter"; so it is basically a logistic multivariate regression with latent predictors.

\[NEXT\]

This results in a response probability function where equal probability points form a straight line in the latent space.
:::

## Multidimensional parameters {.smaller}

<br>

-   Unidimensional 2PL model: $P(Y_i = 1) = \Phi_L (a_i (\theta - b_i))$

    -   $a_i$ = discrimination

    -   $b_i$ = difficulty (location)

-   Multidimensional 2PL model: $P(Y_i = 1) = \Phi_L (\mathbf{a}_i^T \mathbf{\theta} + d_i)$

    -   $\mathbf{a}_i$ = ?

    -   $d_i$ = ?

::: fragment
@reckase_difficulty_1985 [p. 402]:

> \[...\] means of describing the characteristics of an item that takes into account the dimensionality of the skills \[...\] can then be used to determine how or if it is possible to compare items that measure different combinations of abilities.
:::

::: notes
The problem with the M2PL is that its parameters do not have a clear interpretation in IRT terms.

\[NEXT\]

That is why Reckase proposes to find a way of describing the characteristics of multidimensional items, taking their dimensionality into account.

He does it in this 1985 paper, and in a subsequent one in 1991...
:::

## Multidimensional parameters {.smaller}

<br>

::::: columns
::: column
-   @reckase_discriminating_1991:

$$
  MDISC_i = \sqrt{\mathbf{a}_i^T \mathbf{a}_i}
$$

<br>

-   @reckase_difficulty_1985:

$$
  MID_i = \begin{Bmatrix}
    D_i &= \frac{-d_i}{MDISC_i} \\
    \cos \mathbf{\alpha}_i &= \frac{\mathbf{a}_i}{MDISC_i}
  \end{Bmatrix}
$$
:::

::: {.column .fragment}
-   @ackerman_graphical_1996:

```{r m2pl-representation}
#| fig-align:  center
#| fig-height: 6.2
#| fig-width:  6.2

mirt_item <- tribble(
  ~item, ~a_1, ~a_2, ~d,
  1,     1,    0.5,  -0.5
)

mirt_item_params <- mirt_item |>
  compute_mirt_params(d, starts_with('a'))

mirt_item_coords <- mirt_item_params |>
  compute_mirt_coords(D, MDISC, starts_with("cos"), original_coords = FALSE) |>
    rename_with(str_remove, pattern = "transf_")

d_i_spike <- mirt_item_coords |>
  select(item, trait_1 = origin_1, trait_2 = origin_2) |>
  bind_rows(tibble(item = 1, trait_1 = 0, trait_2 = 0)) |>
  add_column(type = c("end", "start")) |>
  pivot_wider(names_from = type, values_from = starts_with("trait"))

mirt_item_coords |>
  ggplot(
    aes(
      origin_1, origin_2,
      xend = end_1,
      yend = end_2,
      color = factor(item), fill = factor(item)
    ),
  )                                                  +
  geom_vline(xintercept = 0, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0, linewidth = GRID_WIDTH) +
  geom_segment(
    data    = d_i_spike,
    mapping = aes(
      trait_1_start, trait_2_start,
      xend = trait_1_end,
      yend = trait_2_end
    ),
    inherit.aes = FALSE,
    linewidth = LINE_WIDTH,
    linetype  = '43',
    color = PALETTE_BLUE["light"]
  ) +
  geom_point(
    data = mirt_item_coords,
    mapping = aes(origin_1, origin_2),
    inherit.aes = FALSE,
    size = I(4)
  ) +
  geom_segment(
    arrow     = arrow(angle = 20, length = unit(10, "points"), type = "closed"),
    linejoin  = "mitre",
    linewidth = VECTOR_WIDTH
  ) +
  scale_x_continuous(
    limits       = c(-1, 1.5),
    breaks       = -1:2,
    minor_breaks = NULL,
    name         = NULL,
    sec.axis     = dup_axis(name = "*&theta;*~2~", labels = NULL)
  ) +
  scale_y_continuous(
    limits       = c(-1, 1.5),
    breaks       = -1:2,
    minor_breaks = NULL,
    name         = NULL,
    sec.axis     = dup_axis(name = "*&theta;*~1~", labels = NULL)
  ) +
  scale_color_manual(values = unname(PALETTE_BLUE["normal"]), guide = NULL) +
  annotate(
    geom = "richtext",
    x = 0.16, y = 0.20,
    label = "*D~i~*",
    size = 9,
    angle = mirt_item_params |> pull(deg_1),
    label.colour  = NA,
    fill = NA
  ) +
  annotate(
    "richtext",
    x = mirt_item_coords |> pull(origin_1) + 0.40,
    y = mirt_item_coords |> pull(origin_2) + 0.32,
    label = "*MDISC~i~*",
    size  = 9,
    angle = mirt_item_params |> pull(deg_1),
    label.colour  = NA,
    fill = NA
  ) +
  coord_fixed(clip = "on") +
  theme(
    axis.line          = element_blank(),
    axis.title         = element_markdown(size = 20, padding = unit(0, "lines")),
    axis.title.x       = element_markdown(hjust = .4),
    axis.title.y.right = element_markdown(vjust = .4),
    axis.ticks         = element_blank(),
    panel.grid.major   = element_line(
      linewidth = GRID_WIDTH,
      linetype  = "17"
    ),
  )
```
:::
:::::

::: notes
...Reckase & McKinley define the item discrimination, which is an extension of the Pythagorean theorem.

This discrimination can be used to define the item difficulty, finding a parameterization that parallels the unidimensional 2PL model.

Note though that, in this case, the difficulty has two components: a "signed distance" from the origin, and a direction.

\[NEXT\]

It is easier to appreciate this with the graphical representation, that was proposed by Ackerman a few years later, where we can see that the item vector is applied at a point given by the difficulty (with its distance and direction), and has a length equal to the discrimination.
The vector direction always goes through the origin, and the sign of the distance (positive in this case), which is the opposite of the intercept sign, will determine whether the vector is displaced forward or backward along that direction.
:::

## Assumptions {.smaller}

<br>

@reckase_difficulty_1985:

-   Increasing monotony (p. 402):

> the probability of answering an item correctly increases monotonically with an increase in each dimension being measured

-   Orthogonality (p. 404):

> \[...\] before performing the differentiation, the constraint that Σ cos^2^**α** = 1 is added to the expression for the slope \[...\]

::: notes
However, there are two assumptions for defining the multidimensional parameters.

First, that the probability of a correct answer increases monotonically with each dimension.

And second, this constrain you see there, which implies that the item is represented in an orthogonal basis.
This is necessary for applying the "general Pythagorean theorem" as we saw before.

So, our purpose is **to relax these two assumptions**.
:::

## Increasing monotony {.smaller}

<br>

::::: columns
::: column
> "*I neglect my duties*"

-- [International Personality Item Pool](https://ipip.ori.org/newAB5CKey.htm#Conscientiousness)
:::

::: {.column .fragment fragment-index="1"}
```{r inverse-item-curve}
#| fig-align:  center
#| fig-asp:    0.5344
#| fig-width:  9

ITEMS <- tribble(
  ~item, ~a,   ~b,
   1,    -1.5, 0
)

logit_curve_data <- tibble(
  latent_trait = PROB_AXIS,
  prob         = ITEMS |>
    select(-item) |>
    pmap(irf, theta = latent_trait) |>
    bind_cols()
) |>
  unnest(
    prob,
    names_sep = '_'
  ) |>
  rename_with(str_remove, pattern = 'prob_\\.{3}')

logit_curve_data_long <- logit_curve_data |>
  pivot_longer(matches('\\d'), names_to = "item", values_to = "prob")

logit_curve_data_long |>
  ggplot(
    mapping = aes(
      latent_trait, prob,
      color = item,
      group = after_scale(color)
    )
  )                                                        +
  geom_line(linewidth = LINE_WIDTH)                        +
  labs(x = "Conscientiousness", y = "P(Y~i~ = 1)")           +
  scale_y_continuous(
    breaks       = 0:1,
    minor_breaks = NULL,
    limits       = 0:1,
    expand       = expansion(mult = c(0, .05))
  )                                                        +
  scale_x_continuous(
    minor_breaks = NULL,
    expand = expansion()
  )                                                        +
  scale_color_manual(values = ITEMS_PALETTE, guide = NULL) +
  theme(
    axis.title       = element_markdown(),
    axis.title.x     = element_text(margin = margin(t = 1, unit = "lines")),
    axis.title.y     = element_markdown(angle = 90),
    panel.grid.major = element_line(linetype = "17", linewidth = GRID_WIDTH),
    # The following is necessary to avoid showing the axis breaks, as dropping
    #   them prevents the axis line from being drawn (see
    #   https://github.com/tidyverse/ggplot2/issues/2983)
    axis.text.x.bottom = element_blank(),
    axis.ticks.length.x.bottom = unit(0, units = "cm"),
    panel.grid.major.x = element_blank()
  )
```
:::
:::::

::::: {.columns .fragment fragment-index="2"}
::: column
$$
  P(Y_i = 1) = \Phi_L (a_i (\theta - b_i))
$$
:::

::: {.column style="color:#5c6eb1;"}
$$
  a_i < 0
$$
:::
:::::

::: {.fragment .r-center}
### Generalization: "Constant monotony"
:::

::: notes
Because they derive from the traditional application of the model to the cognitive domain, but we want to apply multidimensional IRT to non-cognitive items as well, e.g. to personality measures.

\[NEXT\]

Here, we can have items with "decreasing" probability.

\[NEXT\]

Formally, what this would imply is a negative discrimination parameter.

In the multidimensional case, this may imply a mix of positive and negative discrimination parameters, even in the same item.
:::

## Constant monotony {.smaller}

<br>

:::::::::: columns
:::::::: column
$$
  MDISC_i = \sqrt{a_{1i}^2 + ... + a_{ni}^2}
$$

$$
  D_i = \frac{-d_i}{MDISC_i}
$$

$$
  \cos \, \mathbf{\alpha}_i = \frac{\mathbf{a}_i}{MDISC_i}
$$

::::::: {.fragment fragment-index="1"}
::: {style="color:#5c6eb1; padding-left:300px;"}
$d_1 = 0$
:::

::: {style="color:#90214a; padding-left:300px;"}
$d_2 = -0.5$
:::

::: {style="color:#749f4c; padding-left:300px;"}
$d_3 = -0.25$
:::

::: {style="color:#d76f47; padding-left:300px;"}
$d_4 = 0.25$
:::
:::::::
::::::::

::: {.column .fragment fragment-index="1"}
```{r inverse-item-representation-distance}
#| fig-align:  center
#| fig-height: 8
#| fig-width:  8

mirt_item <- tribble(
  ~item, ~a_1, ~a_2, ~d,
  1,      1,    1,    0,
  2,      1,   -1,   -0.5,
  3,     -1,   -1,   -0.25,
  4,     -1,    1,    0.25
)

mirt_item_params <- mirt_item |>
  compute_mirt_params(d, starts_with('a'))

mirt_item_coords <- mirt_item_params |>
  compute_mirt_coords(D, MDISC, starts_with("cos"), original_coords = FALSE) |>
  rename_with(str_remove, pattern = "transf_")

mirt_item_coords |>
  ggplot(
    aes(
      origin_1, origin_2,
      xend = end_1,
      yend = end_2,
      color = factor(item), fill = factor(item)
    ),
  )                                                  +
  geom_vline(xintercept = 0, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0, linewidth = GRID_WIDTH) +
  geom_segment(
    arrow     = arrow(angle = 20, length = unit(10, "points"), type = "closed"),
    linejoin  = "mitre",
    linewidth = VECTOR_WIDTH
  ) +
  scale_x_continuous(
    limits       = c(-1.5, 1.5),
    breaks       = -1:2,
    minor_breaks = NULL,
    name         = NULL,
    sec.axis     = dup_axis(name = "*&theta;*~2~", labels = NULL)
  ) +
  scale_y_continuous(
    limits       = c(-1.5, 1.5),
    breaks       = -1:2,
    minor_breaks = NULL,
    name         = NULL,
    sec.axis     = dup_axis(name = "*&theta;*~1~", labels = NULL)
  ) +
  scale_color_manual(values = ITEMS_PALETTE, guide = NULL) +
  coord_fixed(clip = "on") +
  theme(
    axis.line          = element_blank(),
    axis.title        = element_markdown(size = 20, padding = unit(0, "lines")),
    axis.title.x       = element_markdown(hjust = .5),
    axis.title.y.right = element_markdown(vjust = .5),
    axis.ticks         = element_blank(),
    panel.grid.major   = element_line(
      linewidth = GRID_WIDTH,
      linetype  = "17"
    ),
  )
```
:::
::::::::::

::: notes
The generalization is straightforward: The MDISC implies squared values of the discrimination components (this is the inner product expanded), so it is always positive regardless of the sign of the discrimination components.

The signed distance is also the same, and the direction only needs to take into account the sign of each component when computing that corresponding cosine.

\[NEXT\]

Graphically, we can see that each item points in a direction given by its discrimination components, including their signs (so, 1 1, 1 -1, etc.).

And the displacement from the origin is again with respect to the pointing direction.

This seems obvious enough, but we must explicitly state that this definitions are valid in these cases, and that they have the same interpretation.

\[PAUSE\]

Now, to the "not so obvious"...
:::

## Orthogonality {.smaller}

<br>

:::::: columns
::: column
```{r item-oblique-axes-1}
#| fig-align:  center
#| fig-height: 8
#| fig-width:  8
mirt_item <- tribble(
  ~item, ~a_1, ~a_2, ~d,
  1,     0.5,  1,    0
)

mirt_item_params <- mirt_item |>
  compute_mirt_params(d, starts_with('a'))

mirt_item_coords <- mirt_item_params |>
  compute_mirt_coords(D, MDISC, starts_with("cos"), original_coords = FALSE) |>
    rename_with(str_remove, pattern = "transf_")

item_spikes <- mirt_item_coords        |>
  slice(1 |> rep(2))                   |>
  add_column(dim = 'F' |> paste0(1:2)) |>
  mutate(
    origin_1 = if_else(dim == 'F1', end_1, origin_1),
    origin_2 = if_else(dim == 'F2', end_2, origin_2),
  )

mirt_item_coords |>
  ggplot(
    aes(
      origin_1, origin_2,
      xend = end_1,
      yend = end_2,
      color = factor(item), fill = factor(item)
    ),
  )                                                  +
  geom_vline(xintercept = 0, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0, linewidth = GRID_WIDTH) +
  geom_segment(
    data = item_spikes,
    mapping = aes(origin_1, origin_2, xend = end_1, yend = end_2),
    linewidth = LINE_WIDTH,
    linetype  = '43',
    color     = PALETTE_BLUE["light"]
  ) +
  geom_segment(
    arrow     = arrow(angle = 20, length = unit(10, "points"), type = "closed"),
    linejoin  = "mitre",
    linewidth = VECTOR_WIDTH
  ) +
  scale_x_continuous(
    limits       = c(-0.75, 1.5),
    breaks       = (-1:3)/2,
    minor_breaks = NULL,
    name         = NULL,
    expand       = expansion(),
    sec.axis     = dup_axis(name = "*&theta;*^*^~2~", labels = NULL)
  ) +
  scale_y_continuous(
    limits       = c(-0.75, 1.5),
    breaks       = (-1:2)/2,
    minor_breaks = NULL,
    name         = NULL,
    expand       = expansion(),
    sec.axis     = dup_axis(name = "*&theta;*^*^~1~", labels = NULL)
  ) +
  scale_color_manual(values = unname(PALETTE_BLUE["normal"]), guide = NULL) +
  annotate(
    "richtext",
    x = c(0.5, -0.1), y = c(-0.1, 1),
    label         = c("*a*^*^~11~", "*a*^*^~21~"),
    size          = 8,
    label.colour  = NA,
    label.padding = unit(0, "lines"),
    fill = NA
  ) +
  coord_fixed(clip = "on") +
  theme(
    axis.line          = element_blank(),
    axis.title         = element_markdown(size = 20, padding = unit(0, "lines")),
    axis.title.x       = element_markdown(hjust = 1/3),
    axis.title.y.right = element_markdown(vjust = 1/3),
    axis.ticks         = element_blank(),
    panel.grid.major   = element_line(
      linewidth = GRID_WIDTH,
      linetype  = "17"
    )
  )
```
:::

:::: column
$$
  \begin{align}
    \mathbf{a}^*_1 &=
      \begin{bmatrix}
        0.5 \\
        1
      \end{bmatrix}\\
      d^*_1 &= 0
  \end{align}
$$

<br>

::: fragment
$$
  \begin{align}
    MDISC^*_1 &= \sqrt{
        [0.5, 1]
        \begin{bmatrix}
          0.5 \\
          1
        \end{bmatrix}
      } \\
      &= \sqrt{ 0.5 · 0.5 + 1 · 1 } \\
      &= \sqrt{ 0.25 + 1 } \\
      &= \sqrt{ 1.25 } \approx 1.12
  \end{align}
$$
:::
::::
::::::

::: notes
You'll recall that we needed orthogonal axes to apply Reckase's formulas.

So let's say we have these "theta star" axes,...

\[NEXT\]

...where we have this item, with discrimination (.5, 1) in this coordinate system, where we may compute its MDISC.

We can transform these axes to another coordinate system, for example, applying an oblique rotation...
:::

## Orthogonality {.smaller}

<br>

::::: columns
::: column
```{r item-oblique-axes-2}
#| fig-align:  center
#| fig-height: 8
#| fig-width:  8
BASIS_VEC_1 <- c(1, 0)
BASIS_VEC_2 <- c(-.5, 1)

axis_slope <- BASIS_VEC_2[2] / BASIS_VEC_2[1]
transf_matrix <- cbind(BASIS_VEC_1, BASIS_VEC_2) |> unname()

item_coords_transf <- mirt_item_coords |>
  pivot_longer(-item) |>
  separate(name, into = c("coord", "dim")) |>
  group_by(coord) |>
  mutate(
    dim   = paste0('F', dim),
    value = solve(transf_matrix) %*% value |> drop()
  ) |>
  ungroup() |>
  pivot_wider(names_from = dim)

item_spikes_transf <- item_coords_transf         |>
  select(-item)                                  |>
  filter(coord == "end")                         |>
  bind_rows(tibble(F1 = 0, F2 = 0))              |>
  complete(F1, F2, fill = list(coord = "start")) |>
  filter(F1 != 0 | F2 != 0)

item_spikes_transf <- item_spikes_transf               |>
  bind_rows(item_spikes_transf |> slice(3))            |>
  mutate(n = paste0('F', (row_number() + 1) %% 2 + 1)) |>
  pivot_wider(names_from = coord, values_from = F1:F2)

item_spikes_transf_ob <- item_spikes_transf |>
  pivot_longer(starts_with("F")) |>
  separate(name, into = c("dim", "coord")) |>
  pivot_wider(names_from = coord, values_from = value) |>
  group_by(n) |>
  mutate(
    start = (transf_matrix %*% start |> drop()),
    end   = (transf_matrix %*% end   |> drop())
  ) |>
  pivot_wider(names_from = dim, values_from = start:end)

mirt_item_coords |>
  ggplot(
    aes(
      origin_1, origin_2,
      xend = end_1,
      yend = end_2,
      color = factor(item), fill = factor(item)
    ),
  ) +
  geom_abline(
    intercept = -2:3,
    slope     = -2,
    linewidth = GRID_WIDTH,
    linetype  = "17",
    color     = AXIS_COLOR
  ) +
  geom_abline(slope = -2, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0,  linewidth = GRID_WIDTH) +
  geom_segment(
    data = item_spikes_transf_ob,
    mapping = aes(start_F1, start_F2, xend = end_F1, yend = end_F2),
    inherit.aes = FALSE,
    linewidth = LINE_WIDTH,
    linetype  = '43',
    color     = PALETTE_BLUE["light"]
  ) +
  geom_segment(
    arrow     = arrow(angle = 20, length = unit(10, "points"), type = "closed"),
    linejoin  = "mitre",
    linewidth = VECTOR_WIDTH
  ) +
  scale_x_continuous(
    limits       = c(-0.75, 1.5),
    breaks       = (-2:2)/2 + .375,
    labels       = ~number(. - .375, .1),
    minor_breaks = NULL,
    name         = NULL,
    expand       = expansion(),
    sec.axis     = dup_axis(name = "*&theta;*~2~", labels = NULL)
  ) +
  scale_y_continuous(
    limits       = c(-0.75, 1.5),
    breaks       = (-1:2)/2,
    minor_breaks = NULL,
    name         = NULL,
    expand       = expansion(),
    sec.axis     = dup_axis(name = "*&theta;*~1~", labels = NULL)
  ) +
  scale_color_manual(values = unname(PALETTE_BLUE["normal"]), guide = NULL) +
  annotate(
    "richtext",
    x = c(1, -0.59), y = c(-0.07, 1),
    label         = c("*a*~11~", "*a*~21~"),
    size          = 8,
    label.colour  = NA,
    label.padding = unit(0, "lines"),
    fill = NA
  ) +
  coord_fixed(clip = "on") +
  theme(
    axis.line          = element_blank(),
    axis.title         = element_markdown(size = 20, padding = unit(0, "lines")),
    axis.title.x       = element_markdown(hjust = 0),
    axis.title.y.right = element_markdown(vjust = 1/3),
    axis.ticks         = element_blank(),
    panel.grid.major.y = element_line(
      linewidth = GRID_WIDTH,
      linetype  = "17"
    )
  )
```
:::

::: {.column .fragment}
<br> <br>

> \[...\] before performing the differentiation, the constraint that Σ cos^2^**α** = 1 is added to the expression for the slope \[...\]

--- @reckase_difficulty_1985 [p. 404]
:::
:::::

::: notes
...to obtain these "theta axes" (let's say, because they align with personality traits of our interest).
We have now different discrimination components than in the "theta star" axes.

\[NEXT\]

If we applied here Reckase's restriction...
:::

## Orthogonality {.smaller}

<br>

::::: columns
::: column
```{r item-oblique-axes-4}
#| fig-align:  center
#| fig-height: 8
#| fig-width:  8
mirt_item_transf <- tribble(
  ~item, ~a_1, ~a_2, ~d,
  1,     1,    1,    0
)

mirt_item_params_transf <- mirt_item_transf |>
  compute_mirt_params(d, starts_with('a'))

mirt_item_coords_transf <- mirt_item_params_transf |>
  compute_mirt_coords(D, MDISC, starts_with("cos"), original_coords = FALSE) |>
    rename_with(str_remove, pattern = "transf_")

item_spikes <- mirt_item_coords_transf |>
  slice(1 |> rep(2))                   |>
  add_column(dim = 'F' |> paste0(1:2)) |>
  mutate(
    origin_1 = if_else(dim == 'F1', end_1, origin_1),
    origin_2 = if_else(dim == 'F2', end_2, origin_2),
  )

mirt_item_coords_transf |>
  ggplot(
    aes(
      origin_1, origin_2,
      xend = end_1,
      yend = end_2,
      color = factor(item), fill = factor(item)
    ),
  )                                                  +
  geom_vline(xintercept = 0, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0, linewidth = GRID_WIDTH) +
  geom_segment(
    data = item_spikes,
    mapping = aes(origin_1, origin_2, xend = end_1, yend = end_2),
    linewidth = LINE_WIDTH,
    linetype  = '43',
    color     = PALETTE_BLUE["light"]
  ) +
  geom_segment(
    arrow     = arrow(angle = 20, length = unit(10, "points"), type = "closed"),
    linejoin  = "mitre",
    linewidth = VECTOR_WIDTH
  ) +
  scale_x_continuous(
    limits       = c(-0.75, 1.5),
    breaks       = (-1:3)/2,
    minor_breaks = NULL,
    name         = NULL,
    expand       = expansion(),
    sec.axis     = dup_axis(name = "*&theta;*~2~", labels = NULL)
  ) +
  scale_y_continuous(
    limits       = c(-0.75, 1.5),
    breaks       = (-1:2)/2,
    minor_breaks = NULL,
    name         = NULL,
    expand       = expansion(),
    sec.axis     = dup_axis(name = "*&theta;*~1~", labels = NULL)
  ) +
  scale_color_manual(values = unname(PALETTE_BLUE["normal"]), guide = NULL) +
  annotate(
    "richtext",
    x = c(1, -0.1), y = c(-0.1, 1),
    label         = c("*a*~11~", "*a*~21~"),
    size          = 8,
    label.colour  = NA,
    label.padding = unit(0, "lines"),
    fill = NA
  ) +
  coord_fixed(clip = "on") +
  theme(
    axis.line          = element_blank(),
    axis.title         = element_markdown(size = 20, padding = unit(0, "lines")),
    axis.title.x       = element_markdown(hjust = 1/3),
    axis.title.y.right = element_markdown(vjust = 1/3),
    axis.ticks         = element_blank(),
    panel.grid.major   = element_line(
      linewidth = GRID_WIDTH,
      linetype  = "17"
    )
  )
```
:::

::: column
$$
  \begin{align}
    \mathbf{a}_1 &=
      \begin{bmatrix}
        1 \\
        1
      \end{bmatrix}\\
      d_1 &= 0
  \end{align}
$$

<br>

$$
  \begin{align}
    MDISC_1 &= \sqrt{
        [1, 1]
        \begin{bmatrix}
          1 \\
          1
        \end{bmatrix}
      } \\
      &= \sqrt{ 1 · 1 + 1 · 1 } \\
      &= \sqrt{ 1 + 1 } \\
      &= \sqrt{ 2 } \approx 1.41
  \end{align}
$$
:::
:::::

::: notes
...we would be assuming that the axes are actually orthogonal, implying that the MDISC is not invariant.

As you can see, does not seem intuitively right to assume that both coordinate systems are orthogonal, ¿right?
:::

## Orthogonality {.smaller}

<br>

::::: columns
::: column
```{r item-oblique-axes-2}
```
:::

::: column
<br> <br>

> ~~\[...\] before performing the differentiation, the constraint that Σ cos^2^**α** = 1 is added to the expression for the slope \[...\]~~

--- @reckase_difficulty_1985 [p. 404]
:::
:::::

::: notes
What we actually need is to compute the multidimensional parameters in this general space, which is not necessarily orthogonal (or, we should say more generally, not necessarily orthonormal).
:::

## Generalization: Oblique axes {.smaller}

> \[...\] before performing the differentiation, the constraint that Σ cos^2^**α** = 1 is added to the expression for the slope \[...\]

---@reckase_difficulty_1985 [p. 404]

:::::::::::: columns
:::::: column
::::: r-stack
::: {.fragment fragment-index="1"}
```{r corr-traits-example}
#| fig-width:  5.5
#| fig-height: 5.5

set.seed(7117)

corr_matrix <- matrix(c(1, .5, .5, 1), nrow = 2)

thetas <- rmvnorm(10^4, sigma = corr_matrix) |>
  as_tibble(.name_repair = ~paste0("trait_", 1:2)) |>
    rownames_to_column("i")

thetas |>
  ggplot(mapping = aes(trait_1, trait_2)) +
  geom_vline(xintercept = 0, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0, linewidth = GRID_WIDTH) +
  geom_point(alpha = .33, shape = 16, colour = I(PALETTE_BLUE["normal"])) +
  scale_x_continuous(
    minor_breaks = NULL,
    limits = c(-4, 4),
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "*&theta;*~2~", labels = NULL)
  ) +
  scale_y_continuous(
    minor_breaks = NULL,
    limits = c(-4, 4),
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "*&theta;*~1~", labels = NULL)
  ) +
  coord_fixed() +
  theme(
    axis.ticks       = element_blank(),
    axis.line        = element_blank(),
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(),
    panel.grid.major = element_line(linetype = "17", linewidth = GRID_WIDTH)
  )
```
:::

::: {.fragment .fade-in fragment-index="4"}
```{r corr-traits-transf-example}
#| fig-width:  5.5
#| fig-height: 5.5

# This is a square root of the correlation matrix, then inverted and transposed:
orth_matrix <- c(1, 0, .5, sqrt(.75)) |> matrix(nrow = 2) |> solve() |> t()

thetas_transf <- thetas |>
  pivot_longer(starts_with("trait")) |>
  group_by(i) |>
  mutate(value = orth_matrix %*% value |> drop()) |>
  pivot_wider() |>
  ungroup()

thetas_transf |>
  ggplot(mapping = aes(trait_1, trait_2)) +
  geom_vline(xintercept = 0, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0, linewidth = GRID_WIDTH) +
  geom_point(alpha = .33, shape = 16, colour = I(PALETTE_BLUE["normal"])) +
  scale_x_continuous(
    minor_breaks = NULL,
    limits = c(-4, 4),
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "*&theta;*^*^~2~", labels = NULL)
  ) +
  scale_y_continuous(
    minor_breaks = NULL,
    limits = c(-4, 4),
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "*&theta;*^*^~1~", labels = NULL)
  ) +
  coord_fixed() +
  theme(
    axis.ticks       = element_blank(),
    axis.line        = element_blank(),
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(),
    panel.grid.major = element_line(linetype = "17", linewidth = GRID_WIDTH)
  )
```
:::
:::::
::::::

::::::: column
::: {.fragment fragment-index="1"}
$$
  cor (\mathbf{\theta}) = \mathbf{R}
$$
:::

::: {.fragment fragment-index="2"}
$$
  \mathbf{R} = \begin{bmatrix}
      1 & 0.5 \\
      0.5 & 1
    \end{bmatrix}
$$
:::

::: {.fragment fragment-index="3"}
$$
  \mathbf{\theta}^* = \mathbf{P} \mathbf{\theta}
$$
:::

::: {.fragment fragment-index="4"}
$$
  cor (\mathbf{\theta}^*) =
    \mathbf{R}^* =
    \begin{bmatrix}
      1 & 0 \\
      0 & 1
    \end{bmatrix}
$$
:::
:::::::
::::::::::::

::: notes
Hacer esta generalización no es algo tan directo como ocurría con la probabilidad "monótonamente decreciente".
Si os acordáis, Reckase aplicaba esta restricción de que la suma de cosenos directores era 1 (que se deriva del teorema de Pitágoras).
Pero para aplicarla hace falta asumir que esos cosenos directores son todos ortogonales entre sí, como ya hemos visto.
Si no tenemos ángulos rectos, esa condición no se cumple.

\[NEXT\]

Y lo que nosotros estamos asumiendo es que tenemos una distribución de vectores de rasgo latente ($\theta$) con una matriz de correlaciones $R$ (la que sea).

\[NEXT\]

(En este ejemplo, vamos a usar una distribución normal estandarizada con correlación de 0,5).

Normalmente representamos esos vectores en coordenadas rectangulares (ejes ortogonales), pero para obtener los parámetros multidimensionales generalizados a ejes oblicuos tenemos que representados en una base no ortogonal (aquí también hablaría de "unidades no estandarizadas", es decir, que la distancia entre el 0 y el 1 pueda ser "distinta de la base estándar, pero digamos solamente "no ortogonal" por simplificar).

¿Cómo podríamos solucionar esto?

\[NEXT\]

Lo que podemos hacer es aplicar una "transformación" a los rasgos latentes de forma que pasemos de esa base oblicua (la que sea, aún no sabemos cuál es) a una base ortogonal.
Esto es una transformación de las coordenadas, como la que aplicamos antes para calcular las coordenadas del ítem tras aplicar la rotación oblicua, pero aplicada en este caso al vector de rasgo latente.

Ahora bien, como recordaréis, queremos representar los ítems en unos ejes que sean, por los ángulos relativos entre ellos, una representación fidedigna, de alguna manera, de las correlaciones entre las dimensiones de rasgo latente.

Estas nuevas coordenadas (transformadas), que ahora están en una base ortogonal, queremos que representen una distribución en la que las correlaciones sean nulas, es decir, que la matriz de correlaciones transformada ($R^*$ la vamos a llamar) sea una matriz identidad (unos en la diagonal, ceros en el resto).
Pero hay que tener en cuenta que esto es un "artefacto matemático".
Estamos haciendo esto por conveniencia, pero las coordenadas que realmente nos interesan como investigadores, porque son las que verdaderamente representan algo que nos interesa investigar, son esas coordenadas originales, que pueden representar cualquier fenómeno que sea de nuestro interés (rasgos de personalidad, síntomas, actitudes, inteligencia, conocimientos... lo que estemos investigando).

Llegados a este punto, la incógnita es: ¿Qué matriz es $\mathbf{P}$, es decir, la matriz que nos permite pasar de las coordenadas originales, $\mathbf{\theta}$, a las transformadas $\mathbf{\theta^*}$?
Esto es equivalente a decir, ¿cuáles son los ejes originales en los que tenemos que representar $\mathbf{\theta}$?
(Recordad, para que $\mathbf{R^*}$ sea una matriz identidad, 1's en la diagonal y 0's fuera de ella).
:::

## Generalization: Oblique axes {.smaller}

<center>

<br>

[bit.ly/vote-coords](https://bit.ly/vote-coords) ([slido.com](https://slido.com) : #1904926)

<br>

![](../www/slideshow-assets/bit.ly_vote-coords.svg){width="40%" height="40%"}

</center>

::: notes
En base a lo planteado anteriormente, me gustaría hacer una prueba con vosotras y vosotros.
Aquí tenéis un enlace y QR para participar en una breve "encuesta".
:::

## Generalization: Oblique axes {.smaller}

<br>

<!-- # TODO: Faltan dos primeros paneles -->

:::::: columns
::: {.column width="33%"}
```{r}
#| fig-width:  5.5
#| fig-height: 5.5
thetas_transf |>
  ggplot(mapping = aes(trait_1, trait_2)) +
  geom_point(alpha = .33, shape = 16, colour = I(PALETTE_BLUE["normal"])) +
  scale_x_continuous(
    minor_breaks = NULL,
    labels = \(breaks) '' |> rep(length(breaks)),
    limits = c(-4, 4),
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "{*&theta;*~1~, *&theta;*~2~}", labels = NULL)
  ) +
  scale_y_continuous(
    minor_breaks = NULL,
    labels = \(breaks) '' |> rep(length(breaks)),
    limits = c(-4, 4),
    expand = expansion(),
    name   = NULL
  ) +
  coord_fixed() +
  theme(
    axis.ticks   = element_blank(),
    axis.line    = element_blank(),
    axis.title   = element_markdown(),
    axis.title.y = element_markdown(),
    panel.grid   = element_blank()
  )
```
:::

::: {.column .fragment width="33%"}
```{r question-mark}
fa(
  "question",
  fill           = "#929acb",
  height         = "200px",
  vertical_align = "-10em",
  margin_left    = "4em"
)
```
:::

::: {.column .fragment width="33%"}
```{r corr-traits-transf-example}
#| fig-width:  5.5
#| fig-height: 5.5
```
:::
::::::

¿What is the coordinate system of *θ* to make *θ\** "spherical"?

::: notes
Supongamos una distribución bivariada de rasgos latentes, con una correlación de .5, como la mostrada anteriormente.
Queremos saber cuál es la matriz $\mathbf{P}$ que nos permite pasar de las coordenadas originales, $\mathbf{\theta}$, a las transformadas $\mathbf{\theta^*}$.

Esto es equivalente a preguntar:

\[NEXT\]

¿Cuál es el sistema de coordenadas que representaría adecuadamente los ejes de una distribución bivariada,...

\[NEXT\]

de forma que al transformarlos a coordenadas rectangulares (mediante la matriz $\mathbf{P}$) obtuviésemos una distribución con correlaciones de 0 y varianzas de 1, es decir, una distribución "esférica"?
:::

## Generalization: Oblique axes

```{r create-grid-function}
create_grid <- function(basis,
                        x_limits = c(-4, 4),
                        y_limits = c(-4, 4),
                        break_step = 2) {
  # Argument checking and formatting: ----

  ## Basis vectors tangents and norms
  vec_tan <- basis[2, ] / basis[1, ]

  # Main: ----

  # Define grid limits as a function of the axes limits (in canonical metric):
  grid_box <- expand_grid(x = x_limits, y = y_limits) |> mutate(
    x_lim = if_else(x == x_limits[1], "inf", "sup"),
    y_lim = if_else(y == y_limits[1], "inf", "sup")
  )

  # Select delimiting grid vertices, based on the basis vector directions
  grid_bounds <- bind_rows(
    h = grid_box |> mutate(slope = vec_tan[1], limit = y_lim),
    v = grid_box |> mutate(slope = vec_tan[2], limit = x_lim),
    .id = "grid"
  ) |>
    filter(xor(slope > 0, x_lim == y_lim)) |>
    mutate(limit_intercept = y - slope * x) # Compute their intercept

  # Transformed (1, 1), expressed in the canonical basis
  unit_point <- basis %*% c(1, 1) |> drop() |>
    as_tibble(.name_repair = "minimal") |>
    add_column(name = c('x', 'y') |> paste0('_unit')) |>
    pivot_wider(values_from = value)

  # Grid specifications, in canonical basis
  grid_specs <- grid_bounds |>
    bind_cols(unit_point) |>
    mutate(unit_intercept = y_unit - slope * x_unit) |>
    select(-ends_with("unit")) |>
    mutate(
      n_units  = trunc(limit_intercept / unit_intercept),
      n_breaks = n_units %/% break_step * break_step
    )

  # Create vertical grid data, if the second basis vector is completely vertical
  if (is.infinite(vec_tan[2])) {

    # Create vertical grid data
    v_grid <- grid_bounds |>
      filter(grid == "v") |>
      mutate(
        unit     = basis[1, 1],
        n_units  = trunc(x / unit),
        n_breaks = n_units %/% break_step * break_step
      ) |>
      select(grid, limit, n_breaks, unit) |>
      pivot_wider(names_from = "limit", values_from = n_breaks) |>
      reframe(coord = seq(from = inf, to = sup, by = break_step) * unit) |>
      mutate(
        grid      = 'v',
        slope     = Inf,
        intercept = NA_real_,
        label_pos = coord,
        ref       = as.character(!coord)
      )

    # Filter out non-valid grid values
    grid_specs <- grid_specs |> filter(!is.infinite(slope))
  }

  # Create grid data
  grid <- grid_specs |>
    select(grid, limit, slope, unit_intercept, n_breaks) |>
    pivot_wider(names_from = limit, values_from = n_breaks) |>
    group_by(grid) |>
    reframe(
      slope,
      coord = seq(from = inf, to = sup, by = break_step),
      intercept = coord * unit_intercept
    ) |>
    mutate(
      label_pos = if_else(
        grid == "h",
         intercept   + x_limits[1] * vec_tan[1],
        (y_limits[1] - intercept)  / vec_tan[2]
      ),
      ref = as.character(!intercept)
    )

    if (is.infinite(vec_tan[2])) return(grid |> bind_rows(v_grid))

    grid
}

plot_grid <- function(grid,
                      x_limits = c(-4, 4),
                      y_limits = c(-4, 4),
                      break_step = 2,
                      linetype_axis = 'solid',
                      linetype_grid = 'dashed') {

  # Argument checking and formatting: ----

  ## Grid plot linetypes
  linetypes <- c(`TRUE` = linetype_axis, `FALSE` = linetype_grid)

  ## Basis vectors tangents and norms
  vec_tan <- grid |> distinct(grid, slope) |> deframe()

  ## Axis title positions
  y_axis_titpos <- (y_limits[2] / vec_tan[2] - x_limits[1]) /
    (x_limits[2] - x_limits[1])
  x_axis_titpos <- (x_limits[2] * vec_tan[1] - y_limits[1]) /
    (y_limits[2] - y_limits[1])

  # Main: ----

  # Assign x axis label values and coordinates (to account for the vertical
  #   special case)
  x_labels <- grid |> filter(grid == "v") |> select(coord, label_pos)

  # Create vertical grid geometry if the second vector is completely vertical
  if (is.infinite(vec_tan[2])) {

    geom_vgrid <- geom_vline(
      mapping = aes(xintercept = coord, linetype = ref),
      data = grid |> filter(grid == "v")
    )

    grid <- grid |> filter(grid == "h")
  }

  grid_plot <- grid |>
    ggplot() +
    geom_abline(
      mapping = aes(
        intercept = intercept,
        slope     = slope,
        linetype  = ref
      )
    )

  if (is.infinite(vec_tan[2])) grid_plot <- grid_plot + geom_vgrid

  grid_plot +
    scale_x_continuous(
      minor_breaks = NULL,
      breaks = x_labels |> pull(label_pos),
      labels = x_labels |> pull(coord),
      limits = x_limits,
      expand = expansion(),
      name   = NULL,
      sec.axis = dup_axis(name = "*&theta;*~2~", labels = NULL)
    ) +
    scale_y_continuous(
      minor_breaks = NULL,
      breaks = grid |> filter(grid == "h") |> pull(label_pos),
      labels = grid |> filter(grid == "h") |> pull(coord),
      limits = y_limits,
      expand = expansion(),
      name   = NULL,
      sec.axis = dup_axis(name = "*&theta;*~1~", labels = NULL)
    ) +
    scale_linetype_manual(values = linetypes, guide = NULL) +
    coord_fixed() +
    theme(
      axis.ticks   = element_blank(),
      axis.line    = element_blank(),
      axis.title.x       = element_markdown(hjust = y_axis_titpos),
      axis.title.y.right = element_markdown(vjust = x_axis_titpos)
    )
}

transform_grid <- function(basis,
                           x_limits = c(-4, 4),
                           y_limits = c(-4, 4),
                           break_step = 2,
                           linetype_axis = 'solid',
                           linetype_grid = 'dashed') {

  # Argument checking and formatting: ----

  # Main: ----
  basis |>
    create_grid(
      x_limits   = x_limits,
      y_limits   = y_limits,
      break_step = break_step
    ) |>
    plot_grid(
      x_limits   = x_limits,
      y_limits   = y_limits,
      break_step = break_step,
      linetype_axis = 'solid',
      linetype_grid = 'dashed'
    )
}
```

<br>

:::::: columns
::: {.column width="33%"}
<center>A</center>

```{r grid-orthogonal-rotation}
#| fig-width: 5
rot_basis <- cbind(
  c(sqrt(.75), -sqrt(.25)),
  c(sqrt(.25), sqrt(.75))
)

rot_grid <- transform_grid(rot_basis)

rot_grid
```
:::

::: {.column .fragment width="33%"}
<center>B</center>

```{r grid-ts-basis}
#| fig-width: 5
ts_basis <- cbind(
  c(1, 0),
  c(sqrt(.25), sqrt(.75))
)

ts_grid <- transform_grid(ts_basis)

ts_grid
```
:::

::: {.column .fragment width="33%"}
<center>C</center>

```{r grid-ls-basis}
#| fig-width: 5
ls_basis <- t(solve(ts_basis))

ls_grid <- transform_grid(ls_basis)

ls_grid
```
:::
::::::

::: fragment
[Answers in 3, 2, 1...](https://wall.sli.do/event/dPunB4BMx3S8hfSjaFSuFh?section=ae77dd00-34fb-4600-9677-337a65b04b83)
:::

::: notes
-   ¿Será el A?

\[NEXT\]

-   ¿O puede ser el B?

\[NEXT\]

-   ¿O es el C?

\[NEXT\]

Tomáos vuestro tiempo y en seguida vemos las respuestas.
:::

## Generalization: Oblique axes

<br>

:::::::: columns
::: {.column width="33%"}
<center>A</center>

```{r scatter-orthogonal-rotation}
#| fig-width: 5
thetas_transf <- thetas |>
  group_by(i) |>
  pivot_longer(starts_with("trait")) |>
  mutate(value = rot_basis %*% value |> drop()) |>
  pivot_wider() |>
  ungroup()

rot_grid +
  geom_point(
    data    = thetas_transf,
    mapping = aes(trait_1, trait_2),
    alpha   = .33,
    shape   = 16,
    colour = I(PALETTE_BLUE["normal"])
  )
```
:::

::::: {.column .fragment width="33%"}
<center>B</center>

:::: r-stack
```{r grid-ts-basis}
#| fig-width: 5
```

::: fragment
```{r scatter-ts-basis}
#| fig-width: 5
thetas_transf <- thetas |>
  group_by(i) |>
  pivot_longer(starts_with("trait")) |>
  mutate(value = ts_basis %*% value |> drop()) |>
  pivot_wider() |>
  ungroup()

ts_grid +
  geom_point(
    data    = thetas_transf,
    mapping = aes(trait_1, trait_2),
    alpha   = .33,
    shape   = 16,
    colour = I(PALETTE_BLUE["normal"])
  )
```
:::
::::
:::::

::: {.column .fragment width="33%"}
<center>C</center>

```{r scatter-ls-basis}
#| fig-width: 5
thetas_transf <- thetas |>
  group_by(i) |>
  pivot_longer(starts_with("trait")) |>
  mutate(value = ls_basis %*% value |> drop()) |>
  pivot_wider() |>
  ungroup()

ls_scatter <- ls_grid +
  geom_point(
    data    = thetas_transf,
    mapping = aes(trait_1, trait_2),
    alpha   = .33,
    shape   = 16,
    colour = I(PALETTE_BLUE["normal"])
  )

ls_scatter
```
:::
::::::::

::: notes
El primero yo creo que estamos de acuerdo en que no, ya que solamente es una "rotación ortogonal" (como la llamaríamos en el ámbito psicométrico).
La transformación que se aplica podría dar lugar a correlaciones nulas (es lo que haría la transformación de los ejes mediante Análisis de Componentes Principales), pero las varianzas resultantes no serían estandarizadas.

\[NEXT\]

El segundo caso rotaría el eje vertical de manera que ambos ejes formarían un ángulo de 60º, que tiene un coseno igual a 0.5.
La gente que ha pensado que estos son los ejes correctos, pensáis como yo: que el coseno del ángulo entre los ejes representa la correlación entre las dimensiones.
Por lo tanto, cuanto más alta la correlación, más alineados deben estar los ejes.

\[NEXT\]

Sin embargo, mirad lo que ocurre cuando "proyectamos" esa nube de puntos sobre esos ejes.
La "nube de puntos" se achata y se alarga aún más.

\[NEXT\]

La respuesta correcta era la C, que "compensa" de alguna manera el achatamiento de la nube de puntos, de forma que las coordenadas transformadas que nos quedan (en los ejes ortogonales, los "normales" del gráfico) tienen una distribución "esférica" (con matriz de correlaciones "identidad").

Esto, que visto así parece ya evidente, es algo que al menos a mí me ha llevado varias decenas (puede que "cientos") de horas entender, durante al menos dos años trabajando en este problema.
:::

## Latent space vs. test space {.smaller}

> It has been suggested by some researchers that the angle between the axes represents the degree of correlation.
> However, for sake of clarity, an orthogonal axes system is used in which distance measures and vectors can be easily calculated, understood, and interpreted.

---@ackerman_multidimensional_2005-1 [p. 16]

::::::::: columns
::::: column
::: {.fragment fragment-index="1"}
$$
  P(Y_i = 1) = \Phi_L (\mathbf{a}_i^T \mathbf{\theta} + d_i)
$$
:::

::: {.fragment fragment-index="3"}
$$
  P(Y_i = 1) = \Phi_L ({\mathbf{a}^*_i}^T \mathbf{\theta}^* + d_i)
$$
:::
:::::

::::: column
::: {.fragment fragment-index="2"}
$$
  \mathbf{\theta}^* = \mathbf{P} \mathbf{\theta}
$$
:::

::: {.fragment fragment-index="4"}
$$
  \mathbf{a}^*_i = \mathbf{Q} \mathbf{a}_i \\
$$
:::
:::::
:::::::::

::: fragment
$$
  \mathbf{a}_i^T \mathbf{\theta} =
    {\mathbf{a}^*_i}^T \mathbf{\theta}^* =
    \mathbf{a}_i^T \mathbf{Q}^T \mathbf{P} \mathbf{\theta}
$$
:::

::: fragment
$$
  \mathbf{Q} = (\mathbf{P}^{-1})^T
$$
:::

::: notes
Viendo que en la literatura hay afirmaciones como esta, no me resulta totalmente descorazonador.
Parece que no soy el único al que le ha supuesto un esfuerzo entenderlo (resultados de la "encuesta"?).

Pues bien, lo que ocurre en realidad es lo siguiente:

\[NEXT\]

Este es el modelo de TRI multidimensional, recordaréis (ya lo presentamos, es como un modelo de regresión multivariada, donde los predcitores, $\theta$ son variables latentes).
El modelo tiene que ser invariante; es decir, tiene que dar igual qué sistema de coordenadas usemos para representarlo (y esto por cierto también tiene su paralelismo en análisis factorial).

\[NEXT\]

Así que si transformamos los rasgos latentes de las personas a un sistema de coordenadas ortogonal, los ítems tienen que transformarse también.

\[NEXT\]

Pero recordaréis que llamé Q a su matriz de transformación, mientras que a la de los rasgos latentes la llamé P.

\[NEXT\]

Claro, para que se cumpla la invarianza, se tiene que cumplir esta igualdad: Este término debe ser el mismo, da igual el sistema de coordenadas que utilice, y por lo tanto P y Q no pueden ser la misma matriz.

\[NEXT\]

En realidad, para transformar los ítems hay que usar la matriz inversa (inversa traspuesta, más concretamente) de la matriz de transformación de los rasgos latentes.
:::

## Latent space vs. test space {.smaller}

<br>

::::::: columns
::::: column
<center>Espacio latente</center>

:::: r-stack
```{r grid-latent-space}
#| fig-width: 6.5
ls_grid
```

::: fragment
```{r scatter-latent-space}
#| fig-width: 6.5
ls_scatter
```
:::
::::
:::::

::: {.column .fragment}
<center>Espacio test</center>

```{r grid-test-space}
#| fig-width: 6.5
ts_grid
```
:::
:::::::

::: notes
Y esto nos ha permitido "descubrir" (o darnos cuenta, más bien) de que hay dos "espacios vectoriales" distintos, uno en el que se encuentran las coordenadas de rasgo latente (las puntuaciones factoriales),...

\[NEXT\]

...
y otro distinto, que es el espacio en el que se encuentran las coordenadas de los ítems, y que es donde en realidad habría que representarlos.

Porque en este espacio es donde los ángulos entre los ejes son los que verdaderamente son una representación geométrica de las correlaciones entre las variables latentes.
:::

## Latent space vs. test space {.smaller}

```{r space-inner-products}
# Mahalanobis distance:
TRAIT_VECTOR        <- latex_bf("\\theta")
TRAIT_NORM          <- latex_norm(TRAIT_VECTOR)
TRAIT_VECTOR_T      <- latex_transp(TRAIT_VECTOR)
TRAIT_COV_INV_TRAIT <- latex(TRAIT_VECTOR_T, COV_MATRIX_INV, TRAIT_VECTOR)
MAH_DIST_DEF        <- latex_sqrt(TRAIT_COV_INV_TRAIT)
MAH_DIST_EQ         <- latex_eq(TRAIT_NORM, MAH_DIST_DEF)

INNER_PROD_INV_COV_EQ <- latex_eq(COV_MATRIX_INV, INNER_PROD_TRANSF_DEF)

# Latent space:
TRAIT_VECTOR_1    <- latex_sub(TRAIT_VECTOR, 1)
TRAIT_VECTOR_2    <- latex_sub(TRAIT_VECTOR, 2)
TRAIT_VECTOR_1_T  <- latex_transp(TRAIT_VECTOR_1)

INNER_PROD     <- latex_innerprod(TRAIT_VECTOR_1, TRAIT_VECTOR_2)
INNER_PROD_DEF <- latex(TRAIT_VECTOR_1_T, COV_MATRIX_INV, TRAIT_VECTOR_2)
INNER_PROD_EQ  <- latex_eq(INNER_PROD, INNER_PROD_DEF)

# Test space:
DISCR_VECTOR_J <- latex_sub(DISCR_VECTOR_ANY, AUX_INDEX)
DISCR_VECTOR_T <- latex_transp(DISCR_VECTOR)

INNER_PROD_DISCR     <- latex_innerprod(DISCR_VECTOR, DISCR_VECTOR_J)
INNER_PROD_DISCR_DEF <- latex(DISCR_VECTOR_T, COV_MATRIX, DISCR_VECTOR_J)
INNER_PROD_DISCR_EQ  <- latex_eq(INNER_PROD_DISCR, INNER_PROD_DISCR_DEF)
```

<br>

@mahalanobis_reprint_2018:

::::: columns
::: {.column width="50%"}
<center>$`r MAH_DIST_EQ`$</center>
:::

::: {.column .fragment width="50%"}
$`r INNER_PROD_INV_COV_EQ`$
:::
:::::

<br>

::: fragment
> \[...\] the latent space should be regarded as a general Euclidean space with its *inner product* defined by $`r INNER_PROD_EQ`$.

---@zhang_theoretical_1999 [p. 221]
:::

::: fragment
> \[...\] the *inner product* in \[the test space\] is consistently defined by
>
> $`r INNER_PROD_DISCR_EQ`$

---@zhang_theoretical_1999 [p. 221]
:::

::: notes
Esto da lugar a algo muy curioso, y es que la "norma" de un punto en el espacio latente (la longitud de un vector aplicado en el origen que termina en ese punto) viene determinada por la distancia de Mahalanobis (esto es una reimpresión de 1936, no penséis que es una cosa de anteayer).

\[NEXT\]

Y eso se debe a que, para que la transformación en el espacio latente se cumpla, esa matriz de transformación (la que llamababmos P) tiene que ser "raíz cuadrada" de la inversa de la matriz de covarianzas.

\[NEXT\]

Y, curiosamente también, aparece en la literatura, en el año 1999, en un artículo que no tiene nada que ver con esto.
Lo menciona de soslayo y no profundiza en su deducción como hemos hecho aquí.
Solamente lo utiliza para resolver un problema "parcial" que se encuentra en la deducción de un índice de dimensionalidad, distinto de los parámetros de los que estamos hablando.
Pero podía haber obviado este problema y haber utilizado coordenadas ortogonales, sin importarle si había o no correlaciones entre los rasgos latentes, como ha hecho todo el mundo (antes y después que ellos).

Así que, todos mis respetos por estos autores.

Sin embargo claro, ellos estaban hablando de otra cosa, y sólo tocaban este tema de paso.
Así que al final eso se quedó ahí, y nadie ha hecho caso en 25 años a lo que dijeron Zhang y Stout.
La gente ha usado los modelos, hecho simulaciones, etc., usando distribuciones correlacionadas, y ha usado las fórmulas de Reckase para transformar entre parámetros multidimensionales y los parámetros "lineales" (podríamos llamarlos, los de la formulación original) del modelo.

\[NEXT\]

Y lo más importante: Esto nos demuestra efectivamente que tenemos dos espacios distintos: Uno para los parámetros de las personas (espacio latente) y otro para los ítems (espacio test).
Y esto es así porque la operación del producto interno está definida de manera distinta en los dos espacios; si la matriz del producto interno es la inversa de las covarianzas en el espacio latente, entonces en el espacio test, donde la transformación del parámetro de discriminación es inversa (como recordaréis), pues la matriz del producto interno es la matriz de covarianzas.
:::

## Generalized multidimensional parameters

<br>

$$
`r latex_eq(MDISC_COV_PARAM, DISCR_VECTOR_COV_MODULE)`
$$

<br>

::: fragment
$$
`r latex_eq(
     MIL_COV_PARAM,
     latex_curlybraces("$DISTANCE_COV_DEF$, $DIR_COS_ITEM_VEC_COV_DEF$")
   )`
$$
:::

::: notes
Si tenemos en cuenta las correlaciones (más generalmente, covarianzas) de la distribución de puntuaciones factoriales, lo que obtenemos son estas expresiones para los parámetros multidimensionales.

Muy importante, la distancia con signo y la discriminación multidimensional son iguales, no importa en qué espacio (latente, o test) queramos obtenerlas y representarlas.
Pero los cosenos directores es necesario calcularlos en el espacio test, no en el espacio de puntuaciones latentes, como decía Reckase, porque este es el espacio que va a representar en realidad la relación entre las variables latentes.
:::

## Complete graphical representation

```{r item-plot-out}
#| fig-cap:   Orthogonal (left) and oblique (right) test space
#| fig-align: center
#| fig-asp:     .5
#| fig-width: 13

plot_orth <- plot_orth +
  scale_color_manual(values = ITEMS_PALETTE) +
  theme(
    axis.ticks       = element_blank(),
    legend.title     = element_blank(),
    panel.grid.minor = element_blank()
  )

plot_oblique <- plot_oblique +
  scale_color_manual(values = ITEMS_PALETTE, guide = NULL) +
  theme(axis.ticks = element_blank())

plot_orth + plot_oblique ## TODO: Tema ggplot
```

## Conclusions

<br>

Generalization of multidimensional parameters

-   To constant-monotony items

-   To oblique axes

<br>

::: fragment
Distinction latent space - test space
:::

::: notes
En conclusión, lo que hemos propuesto en este artículo consiste en una generalización:

-   A ítems con monotonía constante, creciente, decreciente o nula.
    Esto se venía haciendo ya como hemos visto, pero creemos que hacía falta formalizarlo.
    No basta con hacerlo y punto, si Reckase dijo que las discriminaciones tiene que ser positivas, tenemos que demostrar que también vale para discriminaciones negativas o nulas (monotonía decreciente, o probabilidad constante).

-   A Ejes oblicuos, las cuales consideramos que son necesarias para representar distribuciones de puntuaciones de rasgo correlacionadas, en general.
    Hemos visto que esto ya lo mencionaron Zhang y Stout, hace 25 años, pero no proporcionaron una "deducción" de estos parámetros multidimensionales, cosa que nosotros hacemos.

\[NEXT\]

Por último, es imortante recalcar este "descubrimiento" (o "redescubrimiento", más bien) del espacio test, ya que tenemos que ser muy conscientes de que los ítems y las personas están en espacios vectoriales distintos.
Aunque esto parezca solamente un "artefacto matemático" (y en el fondo lo es), hemos visto que es verdaderamente crucial tenerlo en cuenta para entender bien los modelos que estamos utilizando.
Además, esta "distinción" se halla también en los modelos factoriales: Existe lo que se denominan "ejes primarios" y "ejes de referencia", propuestos por Thurstone hace años (pero que prácticamente no se utilizan ni tienen en consideración).
<!-- # TODO: Hace cuántos años? (Thurstone) --> Sin embargo, cuidado con esto, porque no está claro que esos espacios en TRI multidimensional se puedan asimilar directamente a los ejes en un modelo de AF.
:::

## To be continued...

<br>

-   Multidimensional IRT - FA parallelism

<br>

::: fragment
-   M2PL model extensions: Graded scale, nominal response, 3PL...
:::

<br>

::: fragment
-   Other models
:::

::: notes
Esta de hecho puede ser una futura línea de investigación.
Puede parecer evidente el paralelismo (o al menos a mí me lo parece), pero si una cosa me ha enseñado el estar trabajando en este problema de los parámetros (y toparme con el problema de los dos espacios), es que no hay que fiarse de las apariencias.
Por eso aplicamos el método científico (el método puramente deductivo, en este caso)

\[NEXT\]

El modelo multidimensional logístico de 2 parámetros se puede considerar como "anidado", o un caso más simple de varios modelos.
Así que partiendo de este como base, lo interesante es aplicarlo a otros más complejos, pero más útiles en la práctica.
Por cierto, que esto también lo hace mucha gente, aplica las fórmulas de Reckase a otros modelos y a correr, se quedan "tan anchos".

El modelo de escala graduada, por ejemplo, sería el equivalente al modelo de AF para respuestas ordinales, con varios umbrales o "posiciones" en TRI.

El de crédito parcial, por ejemplo, puede servir también para formatos con varias opciones de respuesta que miden o pesan en factores diferentes cada una de ella.

El logístico de 3 parámetros es como el de 2, pero añadiendo un parámetro de pseudoazar, para modelar el que un participante acierte un ítem por pura suerte, respondiendo "a ciegas" (esto sólo valdría para ítems cognitivos).

Estos modelos, que son "extensiones" de este más básico, deberían partir de estos parámetros como fundamento.
Pero un investigador no debería dar los parámetros sin más de ese modelo, sin proporcionar una prueba formal de que se aplica esa fórmula y se cumplen los supuestos.

\[NEXT\]

Y por supuesto, existen muchos otros modelos psicométricos, muy útiles y aplicados en la investigación en Psicología y Educación, en los cuales no está claro aún cómo generalizar estos resultados al ámbito multidimensional.
:::

## References {.smaller}