### 2025-07-14 output/IMPS-2025-slideshow.qmd
Add "latent space vs. test space" slides

---
title:     |
  A generalization of multidimensional item response theory parameters
author:    |
  <table>
    <tr>
      <td>Daniel Morillo, Ph.D.</td>
      <td>Mario Luzardo, Ph.D.</td>
    </tr>
    <tr>
      <td>
        <center>
          [![](../www/slideshow-assets/github-logo.png){height="50"}](https://github.com/DaniMori/)&emsp;
          [![](../www/slideshow-assets/orcid.png){height="50"}](https://orcid.org/0000-0003-3021-3878)
        </center>
      </td>
      <td>
        <center>
          [![](../www/slideshow-assets/github-logo.png){height="50"}](https://github.com/mluzardov/)&emsp;
          [![](../www/slideshow-assets/orcid.png){height="50"}](https://orcid.org/0000-0002-9360-2806)
        </center>
      </td>
    </tr>
  </table>
institute: |
  <table>
    <!-- # TODO: Fix vertical alignment in logos-->
    <tr>
      <td style="width: 400px">
        [![](../www/slideshow-assets/logos_fac_psicologia.svg){height="60"}](https://www.uned.es/universidad/facultades/psicologia.html)
      </td>
      <td style="width: 100px">
        [![](../www/slideshow-assets/logo_udelar.svg){height="120"}](https://udelar.edu.uy)
      </td>
    </tr>
  </table>
bibliography:  ../www/Multidimensional-parameters-MCLM.bib
csl:           ../www/apa-old-doi-prefix.csl
date:          "2024-07-17"
date-meta:     "2024-11-27"
date-format:   long
editor:
  mode:        source
  markdown: 
    wrap:      sentence
    canonical: true # TODO: Restore
execute: 
  cache:       true
lang:          en
knitr:
  # opts_knit: # Does not work, see line 102
  #   root_dir:  here::here()
  opts_chunk: 
    results:   asis
    autodep:   true
format:
  revealjs:
    auto-stretch:            true
    code-annotations:        hover
    df-print:                paged
    fig-cap-location:        bottom
    fig-format:              svg
    fig-asp:                 1
    fig-width:               6
    incremental:             false
    keep-md:                 false
    link-external-newwindow: true
    margin:                  0
    self-contained:          true
    slide-number:            false
    theme:                   ../www/slideshow-assets/extra-styles.scss
    transition:              none
    view-distance:           3
    template-partials:
      - ../www/slideshow-assets/title-slide.html
editor_options:
  chunk_output_type: console
---

```{r libraries}
#| cache:   false
#| include: false

library(tibble)
library(dplyr)
library(knitr)
library(lavaan)
library(semPlot)
library(extrafont)
library(tidyr)
library(ggplot2)
library(units)
library(ggtext)
library(ggridges)
library(qgraph)
library(stringr)
library(purrr)
library(plotly)
library(metR)
library(scales)
library(fontawesome)
library(mvtnorm)
library(patchwork)
```

```{r setup}
#| cache:   false
#| include: false
opts_knit$set(root.dir = here::here()) # See lines 43-44

# Graphical configuration options:
loadfonts(device = "win") # Load fonts (if necessary)
par(family = "Arial", adj = 0.5) # Set graphical device params
```

```{r sources}
source("R/Formulae.R",                  encoding = 'UTF-8')
source("R/Output.R",                    encoding = 'UTF-8')
source("R/Mirt_toolbox.R",              encoding = 'UTF-8')
source("src/Graphical_example_paper.R", encoding = 'UTF-8')
```

```{r graphical-constants}
# Color palettes:

PALETTE_UNED <- c(
  normal = "#00533e",
  medium = "#427562",
  light  = "#86a699",
  white  = "#cddad5"
)
PALETTE_APPLE <- c(
  normal = "#749f4c",
  medium = "#8aac5d",
  light  = "#a2bc7e",
  white  = "#cddbb8"
)
PALETTE_BLUE <- c(
  normal = "#5c6eb1",
  medium = "#7683bd",
  light  = "#929acb",
  white  = "#d4d6eb"
)
PALETTE_TANGERINE <- c(
  normal = "#d76f47",
  medium = "#dd8964",
  light  = "#e4a482",
  white  = "#f0cfb9"
)
PALETTE_STRAWBERRY <- c(
  normal = "#da5268",
  medium = "#c87384",
  light  = "#daa6b1",
  white  = "#efdbdf"
)
PALETTE_RASPBERRY <- c(
  normal = "#90214a",
  medium = "#8e4d60",
  light  = "#af848b",
  white  = "#dac7c7"
)

# Factor axis colors:
F1_COLOR <- unname(PALETTE_BLUE["normal"])
F2_COLOR <- unname(PALETTE_RASPBERRY["normal"])

# Item colors:
ITEMS_PALETTE <- c(
  PALETTE_BLUE["normal"],
  PALETTE_RASPBERRY["normal"],
  PALETTE_APPLE["normal"],
  PALETTE_TANGERINE["normal"],
  PALETTE_UNED["medium"]
) |>
  unname() |>
  setNames(1:5)

# Grid breaks:
X_BREAKS <- seq(-3, 3, length.out = 5)

# Probability axes:
PROB_AXIS <- seq(-3.3, 3.3, by = .01)
PROB_AXIS_2D <- seq(-3, 3, by = 0.05) # Axis for the probability contour plots
```

```{r graphical-output-conf}
#| cache: false

# Graphical output constants:
GRAPH_FONT   <- "Arial"  # Replace default font to comply with UNED identity
FONT_SIZE    <- 24       # Replace default font size for slideshow
AXIS_LAB_POS <-  0.5     # Replace default axis label position for slideshow
LINE_WIDTH   <-  0.75    # Replace default line width for slideshow
GRID_WIDTH   <-  0.5     # Width for grid and axis lines
AXIS_COLOR   <- "gray70" # Replace default grid line color for slideshow
VECTOR_WIDTH <- 1.5      # Replace default vector width for slideshow

theme_set( # `ggplot` output configuration
  theme_classic(
    base_size      = FONT_SIZE,
    base_family    = GRAPH_FONT,
    base_line_size = LINE_WIDTH
  ) %+replace%
    theme(
      axis.title.x       = element_text(hjust = AXIS_LAB_POS),
      axis.title.y       = element_text(vjust = AXIS_LAB_POS, angle = 0),
      axis.title.y.right = element_text(vjust = AXIS_LAB_POS, angle = 0),
      panel.grid         = element_line(color = AXIS_COLOR),
      panel.grid.minor   = element_line(linetype = "17", color = AXIS_COLOR)
    )
)

MARGINS <- c(12, 1.3, 8, 1.3) # margins for the sem path plots
```

## Multidimensional IRT {.smaller}

-   Multidimensional 2PL (M2PL) model [@mckinley_extension_1983]:

:::::: columns
:::: column
$$
  P(Y_i = 1) = \Phi_L (\mathbf{a}_i^T \mathbf{\theta} + d_i)
$$

::: {.fragment fragment-index="1"}
```{r mirt-model-irs-3d}
#| fig-width:  5
#| fig-height: 5
mirt_item <- tribble(
  ~item, ~a_1, ~a_2, ~d,
  1,     1,    0.5,  -0.5
)

irs2d <- function(a_1, a_2, d, theta_1, theta_2) {

  logit(a_1 * theta_1 + a_2 * theta_2 + d)
}

mirt_item_probs <- expand_grid(
  trait_2 = PROB_AXIS_2D,
  trait_1 = PROB_AXIS_2D
) |>
  mutate(
    mirt_item |>
      select(-item) |>
      pmap(irs2d, theta_1 = trait_1, theta_2 = trait_2) |>
      bind_cols()
  ) |>
    rename(prob = `...1`)

mirt_prob_surface <- mirt_item_probs |>
  pivot_wider(names_from = trait_1, values_from = prob) |>
    select(-trait_2) |>
    as.matrix()

plot_ly(
  x = ~PROB_AXIS_2D,
  y = ~PROB_AXIS_2D,
  z = ~mirt_prob_surface,
  colorscale = list(c(0, 1), PALETTE_BLUE["normal"] |> rep(2)),
  type = 'surface',
  hovertemplate = 'P(I<sub>i</sub> = 1) = %{z:.2f}<extra></extra>',
  showscale = FALSE,
  hoverinfo = 'Z'
) |>
  layout(
    scene = list(
      xaxis = list(title = "<i>θ</i><sub>1</sub>"),
      yaxis = list(title = "<i>θ</i><sub>2</sub>"),
      zaxis = list(title = "P(I<sub>i</sub> = 1)"),
      camera = list(
        center = list(x = -.1, z = -.2),
        eye = list(x = -.12, y = -1.8, z = 1)
      )
    )
  ) |>
  config(displayModeBar = FALSE, scrollZoom = FALSE)
```
:::
::::

::: {.column .fragment fragment-index="1"}
<br>

```{r mirt-model-irs-contour}
contour_plot <- mirt_item_probs |>
  ggplot(mapping = aes(trait_1, trait_2, z = prob, label = after_stat(level))) +
  geom_vline(xintercept = 0, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0, linewidth = GRID_WIDTH) +
  geom_contour2(
    color = F1_COLOR,
    linewidth = LINE_WIDTH,
    breaks = c(.05, .1, .2, .3, .5, .7, .8, .9, .95),
    label.placer = label_placer_fraction(frac = .90, rot_adjuster = 0)
  ) +
  scale_x_continuous(
    minor_breaks = NULL,
    limits = range(PROB_AXIS_2D),
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "*&theta;*~2~", labels = NULL)
  ) +
  scale_y_continuous(
    minor_breaks = NULL,
    limits = range(PROB_AXIS_2D),
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "*&theta;*~1~", labels = NULL)
  ) +
  coord_fixed() +
  theme(
    axis.ticks       = element_blank(),
    axis.line        = element_blank(),
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(),
    panel.grid.major = element_line(linetype = "17", linewidth = GRID_WIDTH)
  )

contour_plot
```
:::
::::::

::: notes
The Multidimensional 2PL model is one of many extensions of the 2PL model.

It assumes that the probability of a correct response to an item is a function of the linear combination of the latent traits weighted by item discrimination parameters.
(This here is a vector inner product; the "T" means that the discrimination vector is transposed).

And it has this additional "intercept parameter"; so it is basically a logistic multivariate regression with latent predictors.

\[NEXT\]

This results in a response probability function where equal probability points form a straight line in the latent space.
:::

## Multidimensional parameters {.smaller}

<br>

-   Unidimensional 2PL model: $P(Y_i = 1) = \Phi_L (a_i (\theta - b_i))$

    -   $a_i$ = discrimination

    -   $b_i$ = difficulty (location)

-   Multidimensional 2PL model: $P(Y_i = 1) = \Phi_L (\mathbf{a}_i^T \mathbf{\theta} + d_i)$

    -   $\mathbf{a}_i$ = ?

    -   $d_i$ = ?

::: fragment
@reckase_difficulty_1985 [p. 402]:

> \[...\] means of describing the characteristics of an item that takes into account the dimensionality of the skills \[...\] can then be used to determine how or if it is possible to compare items that measure different combinations of abilities.
:::

::: notes
The problem with the M2PL is that its parameters do not have a clear interpretation in IRT terms.

\[NEXT\]

That is why Reckase proposes to find a way of describing the characteristics of multidimensional items, taking their dimensionality into account.

He does it in this 1985 paper, and in a subsequent one in 1991...
:::

## Multidimensional parameters {.smaller}

<br>

::::: columns
::: column
-   @reckase_discriminating_1991:

$$
  MDISC_i = \sqrt{\mathbf{a}_i^T \mathbf{a}_i}
$$

<br>

-   @reckase_difficulty_1985:

$$
  MID_i = \begin{Bmatrix}
    D_i &= \frac{-d_i}{MDISC_i} \\
    \cos \mathbf{\alpha}_i &= \frac{\mathbf{a}_i}{MDISC_i}
  \end{Bmatrix}
$$
:::

::: {.column .fragment}
-   @ackerman_graphical_1996:

```{r m2pl-representation}
#| fig-align:  center
#| fig-height: 6.2
#| fig-width:  6.2

mirt_item <- tribble(
  ~item, ~a_1, ~a_2, ~d,
  1,     1,    0.5,  -0.5
)

mirt_item_params <- mirt_item |>
  compute_mirt_params(d, starts_with('a'))

mirt_item_coords <- mirt_item_params |>
  compute_mirt_coords(D, MDISC, starts_with("cos"), original_coords = FALSE) |>
    rename_with(str_remove, pattern = "transf_")

d_i_spike <- mirt_item_coords |>
  select(item, trait_1 = origin_1, trait_2 = origin_2) |>
  bind_rows(tibble(item = 1, trait_1 = 0, trait_2 = 0)) |>
  add_column(type = c("end", "start")) |>
  pivot_wider(names_from = type, values_from = starts_with("trait"))

mirt_item_coords |>
  ggplot(
    aes(
      origin_1, origin_2,
      xend = end_1,
      yend = end_2,
      color = factor(item), fill = factor(item)
    ),
  )                                                  +
  geom_vline(xintercept = 0, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0, linewidth = GRID_WIDTH) +
  geom_segment(
    data    = d_i_spike,
    mapping = aes(
      trait_1_start, trait_2_start,
      xend = trait_1_end,
      yend = trait_2_end
    ),
    inherit.aes = FALSE,
    linewidth = LINE_WIDTH,
    linetype  = '43',
    color = PALETTE_BLUE["light"]
  ) +
  geom_point(
    data = mirt_item_coords,
    mapping = aes(origin_1, origin_2),
    inherit.aes = FALSE,
    size = I(4)
  ) +
  geom_segment(
    arrow     = arrow(angle = 20, length = unit(10, "points"), type = "closed"),
    linejoin  = "mitre",
    linewidth = VECTOR_WIDTH
  ) +
  scale_x_continuous(
    limits       = c(-1, 1.5),
    breaks       = -1:2,
    minor_breaks = NULL,
    name         = NULL,
    sec.axis     = dup_axis(name = "*&theta;*~2~", labels = NULL)
  ) +
  scale_y_continuous(
    limits       = c(-1, 1.5),
    breaks       = -1:2,
    minor_breaks = NULL,
    name         = NULL,
    sec.axis     = dup_axis(name = "*&theta;*~1~", labels = NULL)
  ) +
  scale_color_manual(values = unname(PALETTE_BLUE["normal"]), guide = NULL) +
  annotate(
    geom = "richtext",
    x = 0.16, y = 0.20,
    label = "*D~i~*",
    size = 9,
    angle = mirt_item_params |> pull(deg_1),
    label.colour  = NA,
    fill = NA
  ) +
  annotate(
    "richtext",
    x = mirt_item_coords |> pull(origin_1) + 0.40,
    y = mirt_item_coords |> pull(origin_2) + 0.32,
    label = "*MDISC~i~*",
    size  = 9,
    angle = mirt_item_params |> pull(deg_1),
    label.colour  = NA,
    fill = NA
  ) +
  coord_fixed(clip = "on") +
  theme(
    axis.line          = element_blank(),
    axis.title         = element_markdown(size = 20, padding = unit(0, "lines")),
    axis.title.x       = element_markdown(hjust = .4),
    axis.title.y.right = element_markdown(vjust = .4),
    axis.ticks         = element_blank(),
    panel.grid.major   = element_line(
      linewidth = GRID_WIDTH,
      linetype  = "17"
    ),
  )
```
:::
:::::

::: notes
...Reckase & McKinley define the item discrimination, which is an extension of the Pythagorean theorem.

This discrimination can be used to define the item difficulty, finding a parameterization that parallels the unidimensional 2PL model.

Note though that, in this case, the difficulty has two components: a "signed distance" from the origin, and a direction.

\[NEXT\]

It is easier to appreciate this with the graphical representation, that was proposed by Ackerman a few years later, where we can see that the item vector is applied at a point given by the difficulty (with its distance and direction), and has a length equal to the discrimination.
The vector direction always goes through the origin, and the sign of the distance (positive in this case), which is the opposite of the intercept sign, will determine whether the vector is displaced forward or backward along that direction.
:::

## Assumptions {.smaller}

<br>

@reckase_difficulty_1985:

-   Increasing monotony (p. 402):

> the probability of answering an item correctly increases monotonically with an increase in each dimension being measured

-   Orthogonality (p. 404):

> \[...\] before performing the differentiation, the constraint that Σ cos^2^**α** = 1 is added to the expression for the slope \[...\]

::: notes
However, there are two assumptions for defining the multidimensional parameters.

First, that the probability of a correct answer increases monotonically with each dimension.

And second, this constrain you see there, which implies that the item is represented in an orthogonal basis.
This is necessary for applying the "general Pythagorean theorem" as we saw before.

So, our purpose is **to relax these two assumptions**.
:::

## Increasing monotony {.smaller}

<br>

::::: columns
::: column
> "*I neglect my duties*"

-- [International Personality Item Pool](https://ipip.ori.org/newAB5CKey.htm#Conscientiousness)
:::

::: {.column .fragment fragment-index="1"}
```{r inverse-item-curve}
#| fig-align:  center
#| fig-asp:    0.5344
#| fig-width:  9

ITEMS <- tribble(
  ~item, ~a,   ~b,
   1,    -1.5, 0
)

logit_curve_data <- tibble(
  latent_trait = PROB_AXIS,
  prob         = ITEMS |>
    select(-item) |>
    pmap(irf, theta = latent_trait) |>
    bind_cols()
) |>
  unnest(
    prob,
    names_sep = '_'
  ) |>
  rename_with(str_remove, pattern = 'prob_\\.{3}')

logit_curve_data_long <- logit_curve_data |>
  pivot_longer(matches('\\d'), names_to = "item", values_to = "prob")

logit_curve_data_long |>
  ggplot(
    mapping = aes(
      latent_trait, prob,
      color = item,
      group = after_scale(color)
    )
  )                                                        +
  geom_line(linewidth = LINE_WIDTH)                        +
  labs(x = "Conscientiousness", y = "P(Y~i~ = 1)")           +
  scale_y_continuous(
    breaks       = 0:1,
    minor_breaks = NULL,
    limits       = 0:1,
    expand       = expansion(mult = c(0, .05))
  )                                                        +
  scale_x_continuous(
    minor_breaks = NULL,
    expand = expansion()
  )                                                        +
  scale_color_manual(values = ITEMS_PALETTE, guide = NULL) +
  theme(
    axis.title       = element_markdown(),
    axis.title.x     = element_text(margin = margin(t = 1, unit = "lines")),
    axis.title.y     = element_markdown(angle = 90),
    panel.grid.major = element_line(linetype = "17", linewidth = GRID_WIDTH),
    # The following is necessary to avoid showing the axis breaks, as dropping
    #   them prevents the axis line from being drawn (see
    #   https://github.com/tidyverse/ggplot2/issues/2983)
    axis.text.x.bottom = element_blank(),
    axis.ticks.length.x.bottom = unit(0, units = "cm"),
    panel.grid.major.x = element_blank()
  )
```
:::
:::::

::::: {.columns .fragment fragment-index="2"}
::: column
$$
  P(Y_i = 1) = \Phi_L (a_i (\theta - b_i))
$$
:::

::: {.column style="color:#5c6eb1;"}
$$
  a_i < 0
$$
:::
:::::

::: {.fragment .r-center}
### Generalization: "Constant monotony"
:::

::: notes
Because they derive from the traditional application of the model to the cognitive domain, but we want to apply multidimensional IRT to non-cognitive items as well, e.g. to personality measures.

\[NEXT\]

Here, we can have items with "decreasing" probability.

\[NEXT\]

Formally, what this would imply is a negative discrimination parameter.

In the multidimensional case, this may imply a mix of positive and negative discrimination parameters, even in the same item.
:::

## Constant monotony {.smaller}

<br>

:::::::::: columns
:::::::: column
$$
  MDISC_i = \sqrt{a_{1i}^2 + ... + a_{ni}^2}
$$

$$
  D_i = \frac{-d_i}{MDISC_i}
$$

$$
  \cos \, \mathbf{\alpha}_i = \frac{\mathbf{a}_i}{MDISC_i}
$$

::::::: {.fragment fragment-index="1"}
::: {style="color:#5c6eb1; padding-left:300px;"}
$d_1 = 0$
:::

::: {style="color:#90214a; padding-left:300px;"}
$d_2 = -0.5$
:::

::: {style="color:#749f4c; padding-left:300px;"}
$d_3 = -0.25$
:::

::: {style="color:#d76f47; padding-left:300px;"}
$d_4 = 0.25$
:::
:::::::
::::::::

::: {.column .fragment fragment-index="1"}
```{r inverse-item-representation-distance}
#| fig-align:  center
#| fig-height: 8
#| fig-width:  8

mirt_item <- tribble(
  ~item, ~a_1, ~a_2, ~d,
  1,      1,    1,    0,
  2,      1,   -1,   -0.5,
  3,     -1,   -1,   -0.25,
  4,     -1,    1,    0.25
)

mirt_item_params <- mirt_item |>
  compute_mirt_params(d, starts_with('a'))

mirt_item_coords <- mirt_item_params |>
  compute_mirt_coords(D, MDISC, starts_with("cos"), original_coords = FALSE) |>
  rename_with(str_remove, pattern = "transf_")

mirt_item_coords |>
  ggplot(
    aes(
      origin_1, origin_2,
      xend = end_1,
      yend = end_2,
      color = factor(item), fill = factor(item)
    ),
  )                                                  +
  geom_vline(xintercept = 0, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0, linewidth = GRID_WIDTH) +
  geom_segment(
    arrow     = arrow(angle = 20, length = unit(10, "points"), type = "closed"),
    linejoin  = "mitre",
    linewidth = VECTOR_WIDTH
  ) +
  scale_x_continuous(
    limits       = c(-1.5, 1.5),
    breaks       = -1:2,
    minor_breaks = NULL,
    name         = NULL,
    sec.axis     = dup_axis(name = "*&theta;*~2~", labels = NULL)
  ) +
  scale_y_continuous(
    limits       = c(-1.5, 1.5),
    breaks       = -1:2,
    minor_breaks = NULL,
    name         = NULL,
    sec.axis     = dup_axis(name = "*&theta;*~1~", labels = NULL)
  ) +
  scale_color_manual(values = ITEMS_PALETTE, guide = NULL) +
  coord_fixed(clip = "on") +
  theme(
    axis.line          = element_blank(),
    axis.title        = element_markdown(size = 20, padding = unit(0, "lines")),
    axis.title.x       = element_markdown(hjust = .5),
    axis.title.y.right = element_markdown(vjust = .5),
    axis.ticks         = element_blank(),
    panel.grid.major   = element_line(
      linewidth = GRID_WIDTH,
      linetype  = "17"
    ),
  )
```
:::
::::::::::

::: notes
The generalization is straightforward: The MDISC implies squared values of the discrimination components (this is the inner product expanded), so it is always positive regardless of the sign of the discrimination components.

The signed distance is also the same, and the direction only needs to take into account the sign of each component when computing that corresponding cosine.

\[NEXT\]

Graphically, we can see that each item points in a direction given by its discrimination components, including their signs (so, 1 1, 1 -1, etc.).

And the displacement from the origin is again with respect to the pointing direction.

This seems obvious enough, but we must explicitly state that this definitions are valid in these cases, and that they have the same interpretation.

\[PAUSE\]

Now, to the "not so obvious"...
:::

## Orthogonality {.smaller}

<br>

:::::: columns
::: column
```{r item-oblique-axes-1}
#| fig-align:  center
#| fig-height: 8
#| fig-width:  8
mirt_item <- tribble(
  ~item, ~a_1, ~a_2, ~d,
  1,     0.5,  1,    0
)

mirt_item_params <- mirt_item |>
  compute_mirt_params(d, starts_with('a'))

mirt_item_coords <- mirt_item_params |>
  compute_mirt_coords(D, MDISC, starts_with("cos"), original_coords = FALSE) |>
    rename_with(str_remove, pattern = "transf_")

item_spikes <- mirt_item_coords        |>
  slice(1 |> rep(2))                   |>
  add_column(dim = 'F' |> paste0(1:2)) |>
  mutate(
    origin_1 = if_else(dim == 'F1', end_1, origin_1),
    origin_2 = if_else(dim == 'F2', end_2, origin_2),
  )

mirt_item_coords |>
  ggplot(
    aes(
      origin_1, origin_2,
      xend = end_1,
      yend = end_2,
      color = factor(item), fill = factor(item)
    ),
  )                                                  +
  geom_vline(xintercept = 0, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0, linewidth = GRID_WIDTH) +
  geom_segment(
    data = item_spikes,
    mapping = aes(origin_1, origin_2, xend = end_1, yend = end_2),
    linewidth = LINE_WIDTH,
    linetype  = '43',
    color     = PALETTE_BLUE["light"]
  ) +
  geom_segment(
    arrow     = arrow(angle = 20, length = unit(10, "points"), type = "closed"),
    linejoin  = "mitre",
    linewidth = VECTOR_WIDTH
  ) +
  scale_x_continuous(
    limits       = c(-0.75, 1.5),
    breaks       = (-1:3)/2,
    minor_breaks = NULL,
    name         = NULL,
    expand       = expansion(),
    sec.axis     = dup_axis(name = "*&theta;*^*^~2~", labels = NULL)
  ) +
  scale_y_continuous(
    limits       = c(-0.75, 1.5),
    breaks       = (-1:2)/2,
    minor_breaks = NULL,
    name         = NULL,
    expand       = expansion(),
    sec.axis     = dup_axis(name = "*&theta;*^*^~1~", labels = NULL)
  ) +
  scale_color_manual(values = unname(PALETTE_BLUE["normal"]), guide = NULL) +
  annotate(
    "richtext",
    x = c(0.5, -0.1), y = c(-0.1, 1),
    label         = c("*a*^*^~11~", "*a*^*^~21~"),
    size          = 8,
    label.colour  = NA,
    label.padding = unit(0, "lines"),
    fill = NA
  ) +
  coord_fixed(clip = "on") +
  theme(
    axis.line          = element_blank(),
    axis.title         = element_markdown(size = 20, padding = unit(0, "lines")),
    axis.title.x       = element_markdown(hjust = 1/3),
    axis.title.y.right = element_markdown(vjust = 1/3),
    axis.ticks         = element_blank(),
    panel.grid.major   = element_line(
      linewidth = GRID_WIDTH,
      linetype  = "17"
    )
  )
```
:::

:::: column
$$
  \begin{align}
    \mathbf{a}^*_1 &=
      \begin{bmatrix}
        0.5 \\
        1
      \end{bmatrix}\\
      d^*_1 &= 0
  \end{align}
$$

<br>

::: fragment
$$
  \begin{align}
    MDISC^*_1 &= \sqrt{
        [0.5, 1]
        \begin{bmatrix}
          0.5 \\
          1
        \end{bmatrix}
      } \\
      &= \sqrt{ 0.5 · 0.5 + 1 · 1 } \\
      &= \sqrt{ 0.25 + 1 } \\
      &= \sqrt{ 1.25 } \approx 1.12
  \end{align}
$$
:::
::::
::::::

::: notes
You'll recall that we needed orthogonal axes to apply Reckase's formulas.

So let's say we have these "theta star" axes,...

\[NEXT\]

...where we have this item, with discrimination (.5, 1) in this coordinate system, where we may compute its MDISC.

We can transform these axes to another coordinate system, for example, applying an oblique rotation...
:::

## Orthogonality {.smaller}

<br>

::::: columns
::: column
```{r item-oblique-axes-2}
#| fig-align:  center
#| fig-height: 8
#| fig-width:  8
BASIS_VEC_1 <- c(1, 0)
BASIS_VEC_2 <- c(-.5, 1)

axis_slope <- BASIS_VEC_2[2] / BASIS_VEC_2[1]
transf_matrix <- cbind(BASIS_VEC_1, BASIS_VEC_2) |> unname()

item_coords_transf <- mirt_item_coords |>
  pivot_longer(-item) |>
  separate(name, into = c("coord", "dim")) |>
  group_by(coord) |>
  mutate(
    dim   = paste0('F', dim),
    value = solve(transf_matrix) %*% value |> drop()
  ) |>
  ungroup() |>
  pivot_wider(names_from = dim)

item_spikes_transf <- item_coords_transf         |>
  select(-item)                                  |>
  filter(coord == "end")                         |>
  bind_rows(tibble(F1 = 0, F2 = 0))              |>
  complete(F1, F2, fill = list(coord = "start")) |>
  filter(F1 != 0 | F2 != 0)

item_spikes_transf <- item_spikes_transf               |>
  bind_rows(item_spikes_transf |> slice(3))            |>
  mutate(n = paste0('F', (row_number() + 1) %% 2 + 1)) |>
  pivot_wider(names_from = coord, values_from = F1:F2)

item_spikes_transf_ob <- item_spikes_transf |>
  pivot_longer(starts_with("F")) |>
  separate(name, into = c("dim", "coord")) |>
  pivot_wider(names_from = coord, values_from = value) |>
  group_by(n) |>
  mutate(
    start = (transf_matrix %*% start |> drop()),
    end   = (transf_matrix %*% end   |> drop())
  ) |>
  pivot_wider(names_from = dim, values_from = start:end)

mirt_item_coords |>
  ggplot(
    aes(
      origin_1, origin_2,
      xend = end_1,
      yend = end_2,
      color = factor(item), fill = factor(item)
    ),
  ) +
  geom_abline(
    intercept = -2:3,
    slope     = -2,
    linewidth = GRID_WIDTH,
    linetype  = "17",
    color     = AXIS_COLOR
  ) +
  geom_abline(slope = -2, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0,  linewidth = GRID_WIDTH) +
  geom_segment(
    data = item_spikes_transf_ob,
    mapping = aes(start_F1, start_F2, xend = end_F1, yend = end_F2),
    inherit.aes = FALSE,
    linewidth = LINE_WIDTH,
    linetype  = '43',
    color     = PALETTE_BLUE["light"]
  ) +
  geom_segment(
    arrow     = arrow(angle = 20, length = unit(10, "points"), type = "closed"),
    linejoin  = "mitre",
    linewidth = VECTOR_WIDTH
  ) +
  scale_x_continuous(
    limits       = c(-0.75, 1.5),
    breaks       = (-2:2)/2 + .375,
    labels       = ~number(. - .375, .1),
    minor_breaks = NULL,
    name         = NULL,
    expand       = expansion(),
    sec.axis     = dup_axis(name = "*&theta;*~2~", labels = NULL)
  ) +
  scale_y_continuous(
    limits       = c(-0.75, 1.5),
    breaks       = (-1:2)/2,
    minor_breaks = NULL,
    name         = NULL,
    expand       = expansion(),
    sec.axis     = dup_axis(name = "*&theta;*~1~", labels = NULL)
  ) +
  scale_color_manual(values = unname(PALETTE_BLUE["normal"]), guide = NULL) +
  annotate(
    "richtext",
    x = c(1, -0.59), y = c(-0.07, 1),
    label         = c("*a*~11~", "*a*~21~"),
    size          = 8,
    label.colour  = NA,
    label.padding = unit(0, "lines"),
    fill = NA
  ) +
  coord_fixed(clip = "on") +
  theme(
    axis.line          = element_blank(),
    axis.title         = element_markdown(size = 20, padding = unit(0, "lines")),
    axis.title.x       = element_markdown(hjust = 0),
    axis.title.y.right = element_markdown(vjust = 1/3),
    axis.ticks         = element_blank(),
    panel.grid.major.y = element_line(
      linewidth = GRID_WIDTH,
      linetype  = "17"
    )
  )
```
:::

::: {.column .fragment}
<br> <br>

> \[...\] before performing the differentiation, the constraint that Σ cos^2^**α** = 1 is added to the expression for the slope \[...\]

--- @reckase_difficulty_1985 [p. 404]
:::
:::::

::: notes
...to obtain these "theta axes" (let's say, because they align with personality traits of our interest).
We have now different discrimination components than in the "theta star" axes.

\[NEXT\]

If we applied here Reckase's restriction...
:::

## Orthogonality {.smaller}

<br>

::::: columns
::: column
```{r item-oblique-axes-4}
#| fig-align:  center
#| fig-height: 8
#| fig-width:  8
mirt_item_transf <- tribble(
  ~item, ~a_1, ~a_2, ~d,
  1,     1,    1,    0
)

mirt_item_params_transf <- mirt_item_transf |>
  compute_mirt_params(d, starts_with('a'))

mirt_item_coords_transf <- mirt_item_params_transf |>
  compute_mirt_coords(D, MDISC, starts_with("cos"), original_coords = FALSE) |>
    rename_with(str_remove, pattern = "transf_")

item_spikes <- mirt_item_coords_transf |>
  slice(1 |> rep(2))                   |>
  add_column(dim = 'F' |> paste0(1:2)) |>
  mutate(
    origin_1 = if_else(dim == 'F1', end_1, origin_1),
    origin_2 = if_else(dim == 'F2', end_2, origin_2),
  )

mirt_item_coords_transf |>
  ggplot(
    aes(
      origin_1, origin_2,
      xend = end_1,
      yend = end_2,
      color = factor(item), fill = factor(item)
    ),
  )                                                  +
  geom_vline(xintercept = 0, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0, linewidth = GRID_WIDTH) +
  geom_segment(
    data = item_spikes,
    mapping = aes(origin_1, origin_2, xend = end_1, yend = end_2),
    linewidth = LINE_WIDTH,
    linetype  = '43',
    color     = PALETTE_BLUE["light"]
  ) +
  geom_segment(
    arrow     = arrow(angle = 20, length = unit(10, "points"), type = "closed"),
    linejoin  = "mitre",
    linewidth = VECTOR_WIDTH
  ) +
  scale_x_continuous(
    limits       = c(-0.75, 1.5),
    breaks       = (-1:3)/2,
    minor_breaks = NULL,
    name         = NULL,
    expand       = expansion(),
    sec.axis     = dup_axis(name = "*&theta;*~2~", labels = NULL)
  ) +
  scale_y_continuous(
    limits       = c(-0.75, 1.5),
    breaks       = (-1:2)/2,
    minor_breaks = NULL,
    name         = NULL,
    expand       = expansion(),
    sec.axis     = dup_axis(name = "*&theta;*~1~", labels = NULL)
  ) +
  scale_color_manual(values = unname(PALETTE_BLUE["normal"]), guide = NULL) +
  annotate(
    "richtext",
    x = c(1, -0.1), y = c(-0.1, 1),
    label         = c("*a*~11~", "*a*~21~"),
    size          = 8,
    label.colour  = NA,
    label.padding = unit(0, "lines"),
    fill = NA
  ) +
  coord_fixed(clip = "on") +
  theme(
    axis.line          = element_blank(),
    axis.title         = element_markdown(size = 20, padding = unit(0, "lines")),
    axis.title.x       = element_markdown(hjust = 1/3),
    axis.title.y.right = element_markdown(vjust = 1/3),
    axis.ticks         = element_blank(),
    panel.grid.major   = element_line(
      linewidth = GRID_WIDTH,
      linetype  = "17"
    )
  )
```
:::

::: column
$$
  \begin{align}
    \mathbf{a}_1 &=
      \begin{bmatrix}
        1 \\
        1
      \end{bmatrix}\\
      d_1 &= 0
  \end{align}
$$

<br>

$$
  \begin{align}
    MDISC_1 &= \sqrt{
        [1, 1]
        \begin{bmatrix}
          1 \\
          1
        \end{bmatrix}
      } \\
      &= \sqrt{ 1 · 1 + 1 · 1 } \\
      &= \sqrt{ 1 + 1 } \\
      &= \sqrt{ 2 } \approx 1.41
  \end{align}
$$
:::
:::::

::: notes
...we would be assuming that the axes are actually orthogonal, implying that the MDISC is not invariant.

As you can see, does not seem intuitively right to assume that both coordinate systems are orthogonal, ¿right?
:::

## Orthogonality {.smaller}

<br>

::::: columns
::: column
```{r item-oblique-axes-2}
```
:::

::: column
<br> <br>

> ~~\[...\] before performing the differentiation, the constraint that Σ cos^2^**α** = 1 is added to the expression for the slope \[...\]~~

--- @reckase_difficulty_1985 [p. 404]
:::
:::::

::: {.fragment .r-center}
### Generalization: "Oblique axes"
:::

::: notes
What we actually need is to compute the multidimensional parameters in this general space,...

\[NEXT\]

...which is not necessarily orthogonal (or, we should say more generally, not necessarily orthonormal).

¿Why? Because we want this space to have a meaningful interpretation; we want it to be a geometrical representation of the probabilistic properties of our sample.
:::

## Oblique axes {.smaller}

<br>

:::::::::: columns
:::::: column
$$
  cor (\mathbf{\theta}) = \mathbf{\Sigma}
$$

::: {.fragment fragment-index="1"}
$$
  \mathbf{\Sigma} = \begin{bmatrix}
      1 & 0.5 \\
      0.5 & 1
    \end{bmatrix}
$$
:::

::: {.fragment fragment-index="2"}
$$
  \mathbf{\theta}^* = \mathbf{P} \mathbf{\theta}
$$
:::

::: {.fragment fragment-index="2"}
$$
  cor (\mathbf{\theta}^*) =
    \mathbf{\Sigma}^* =
    \begin{bmatrix}
      1 & 0 \\
      0 & 1
    \end{bmatrix}
$$
:::
::::::

::::: column
:::: {.r-stack .fragment .fade-in fragment-index="1"}
```{r corr-traits-example}
#| fig-width:  5.5
#| fig-height: 5.5

set.seed(7117)

corr_matrix <- matrix(c(1, .5, .5, 1), nrow = 2)

thetas <- rmvnorm(10^4, sigma = corr_matrix) |>
  as_tibble(.name_repair = ~paste0("trait_", 1:2)) |>
    rownames_to_column("i")

thetas |>
  ggplot(mapping = aes(trait_1, trait_2)) +
  geom_vline(xintercept = 0, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0, linewidth = GRID_WIDTH) +
  geom_point(alpha = .33, shape = 16, colour = I(PALETTE_BLUE["normal"])) +
  scale_x_continuous(
    minor_breaks = NULL,
    limits = c(-4, 4),
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "*&theta;*~2~", labels = NULL)
  ) +
  scale_y_continuous(
    minor_breaks = NULL,
    limits = c(-4, 4),
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "*&theta;*~1~", labels = NULL)
  ) +
  coord_fixed() +
  theme(
    axis.ticks       = element_blank(),
    axis.line        = element_blank(),
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(),
    panel.grid.major = element_line(linetype = "17", linewidth = GRID_WIDTH)
  )
```

::: {.fragment .fade-in fragment-index="4"}
```{r corr-traits-transf-example}
#| fig-width:  5.5
#| fig-height: 5.5

# This is a square root of the correlation matrix, then inverted and transposed:
orth_matrix <- c(1, 0, .5, sqrt(.75)) |> matrix(nrow = 2) |> solve() |> t()

thetas_transf <- thetas |>
  pivot_longer(starts_with("trait")) |>
  group_by(i) |>
  mutate(value = orth_matrix %*% value |> drop()) |>
  pivot_wider() |>
  ungroup()

thetas_transf |>
  ggplot(mapping = aes(trait_1, trait_2)) +
  geom_vline(xintercept = 0, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0, linewidth = GRID_WIDTH) +
  geom_point(alpha = .33, shape = 16, colour = I(PALETTE_BLUE["normal"])) +
  scale_x_continuous(
    minor_breaks = NULL,
    limits = c(-4, 4),
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "*&theta;*^*^~2~", labels = NULL)
  ) +
  scale_y_continuous(
    minor_breaks = NULL,
    limits = c(-4, 4),
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "*&theta;*^*^~1~", labels = NULL)
  ) +
  coord_fixed() +
  theme(
    axis.ticks       = element_blank(),
    axis.line        = element_blank(),
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(),
    panel.grid.major = element_line(linetype = "17", linewidth = GRID_WIDTH)
  )
```
:::
::::
:::::
::::::::::

::: notes
So, let's say $\theta$ is multivariate normal with covariane matrix $\mathbf{\Sigma}$.

\[NEXT\]

For the sake of the example, let's say it has unitary variances with a 0,5 correlation.

\[NEXT\]

Now, we can apply a "transformation" in the latent space that results in an identity covariance matrix.

\[NEXT\]

And if we want the geometrical space to be a representation of the latent space, the meaningful way is to represent these coordinates in an orthonormal basis (the covariane matrix is an identity matrix, the basis also forms an identity matrix).
:::

## Oblique axes {.smaller}

<br>

:::::: columns
::: {.column width="33%"}
```{r}
#| fig-width:  5.5
#| fig-height: 5.5

ts_basis <- cbind(
  c(1, 0),
  c(sqrt(.25), sqrt(.75))
)
ls_basis <- t(solve(ts_basis))

thetas_transf <- thetas |>
  group_by(i) |>
  pivot_longer(starts_with("trait")) |>
  mutate(value = ls_basis %*% value |> drop()) |>
  pivot_wider() |>
  ungroup()

thetas_transf |>
  ggplot(mapping = aes(trait_1, trait_2)) +
  geom_point(alpha = .33, shape = 16, colour = I(PALETTE_BLUE["normal"])) +
  scale_x_continuous(
    minor_breaks = NULL,
    labels = \(breaks) '' |> rep(length(breaks)),
    limits = c(-4, 4),
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "{*&theta;*~1~, *&theta;*~2~}", labels = NULL)
  ) +
  scale_y_continuous(
    minor_breaks = NULL,
    labels = \(breaks) '' |> rep(length(breaks)),
    limits = c(-4, 4),
    expand = expansion(),
    name   = NULL
  ) +
  coord_fixed() +
  theme(
    axis.ticks   = element_blank(),
    axis.line    = element_blank(),
    axis.title   = element_markdown(),
    axis.title.y = element_markdown(),
    panel.grid   = element_blank()
  )
```
:::

::: {.column .fragment width="33%" fragment-index="1"}
```{r question-mark}
fa(
  "question",
  fill           = "#929acb",
  height         = "200px",
  vertical_align = "-10em",
  margin_left    = "4em"
)
```
:::

::: {.column .fragment width="33%" fragment-index="2"}
```{r corr-traits-transf-example}
#| fig-width:  5.5
#| fig-height: 5.5
```
:::
::::::

<center>

::: {.fragment fragment-index="2"}
Which coordinate system *θ* makes *θ\** "spherical"?
:::

</center>

::: notes
So, if the basis matrix of $\mathbf{\theta^*}$, which has an identity covariance matrix, is an identity covariance matrix, what is $\mathbf{P}$, the basis matrix of $\mathbf{\theta}$.

In other words,...

\[NEXT\]

...what is the coordinate system of $\mathbf{\theta}$...

\[NEXT\]

...that makes $\mathbf{\theta^*}$ spherical?
:::

## Oblique axes

```{r create-grid-function}
create_grid <- function(basis,
                        x_limits = c(-4, 4),
                        y_limits = c(-4, 4),
                        break_step = 2) {
  # Argument checking and formatting: ----

  ## Basis vectors tangents and norms
  vec_tan <- basis[2, ] / basis[1, ]

  # Main: ----

  # Define grid limits as a function of the axes limits (in canonical metric):
  grid_box <- expand_grid(x = x_limits, y = y_limits) |> mutate(
    x_lim = if_else(x == x_limits[1], "inf", "sup"),
    y_lim = if_else(y == y_limits[1], "inf", "sup")
  )

  # Select delimiting grid vertices, based on the basis vector directions
  grid_bounds <- bind_rows(
    h = grid_box |> mutate(slope = vec_tan[1], limit = y_lim),
    v = grid_box |> mutate(slope = vec_tan[2], limit = x_lim),
    .id = "grid"
  ) |>
    filter(xor(slope > 0, x_lim == y_lim)) |>
    mutate(limit_intercept = y - slope * x) # Compute their intercept

  # Transformed (1, 1), expressed in the canonical basis
  unit_point <- basis %*% c(1, 1) |> drop() |>
    as_tibble(.name_repair = "minimal") |>
    add_column(name = c('x', 'y') |> paste0('_unit')) |>
    pivot_wider(values_from = value)

  # Grid specifications, in canonical basis
  grid_specs <- grid_bounds |>
    bind_cols(unit_point) |>
    mutate(unit_intercept = y_unit - slope * x_unit) |>
    select(-ends_with("unit")) |>
    mutate(
      n_units  = trunc(limit_intercept / unit_intercept),
      n_breaks = n_units %/% break_step * break_step
    )

  # Create vertical grid data, if the second basis vector is completely vertical
  if (is.infinite(vec_tan[2])) {

    # Create vertical grid data
    v_grid <- grid_bounds |>
      filter(grid == "v") |>
      mutate(
        unit     = basis[1, 1],
        n_units  = trunc(x / unit),
        n_breaks = n_units %/% break_step * break_step
      ) |>
      select(grid, limit, n_breaks, unit) |>
      pivot_wider(names_from = "limit", values_from = n_breaks) |>
      reframe(coord = seq(from = inf, to = sup, by = break_step) * unit) |>
      mutate(
        grid      = 'v',
        slope     = Inf,
        intercept = NA_real_,
        label_pos = coord,
        ref       = as.character(!coord)
      )

    # Filter out non-valid grid values
    grid_specs <- grid_specs |> filter(!is.infinite(slope))
  }

  # Create grid data
  grid <- grid_specs |>
    select(grid, limit, slope, unit_intercept, n_breaks) |>
    pivot_wider(names_from = limit, values_from = n_breaks) |>
    group_by(grid) |>
    reframe(
      slope,
      coord = seq(from = inf, to = sup, by = break_step),
      intercept = coord * unit_intercept
    ) |>
    mutate(
      label_pos = if_else(
        grid == "h",
         intercept   + x_limits[1] * vec_tan[1],
        (y_limits[1] - intercept)  / vec_tan[2]
      ),
      ref = as.character(!intercept)
    )

    if (is.infinite(vec_tan[2])) return(grid |> bind_rows(v_grid))

    grid
}

plot_grid <- function(grid,
                      x_limits = c(-4, 4),
                      y_limits = c(-4, 4),
                      break_step = 2,
                      linetype_axis = 'solid',
                      linetype_grid = 'dashed') {

  # Argument checking and formatting: ----

  ## Grid plot linetypes
  linetypes <- c(`TRUE` = linetype_axis, `FALSE` = linetype_grid)

  ## Basis vectors tangents and norms
  vec_tan <- grid |> distinct(grid, slope) |> deframe()

  ## Axis title positions
  y_axis_titpos <- (y_limits[2] / vec_tan[2] - x_limits[1]) /
    (x_limits[2] - x_limits[1])
  x_axis_titpos <- (x_limits[2] * vec_tan[1] - y_limits[1]) /
    (y_limits[2] - y_limits[1])

  # Main: ----

  # Assign x axis label values and coordinates (to account for the vertical
  #   special case)
  x_labels <- grid |> filter(grid == "v") |> select(coord, label_pos)

  # Create vertical grid geometry if the second vector is completely vertical
  if (is.infinite(vec_tan[2])) {

    geom_vgrid <- geom_vline(
      mapping = aes(xintercept = coord, linetype = ref),
      data = grid |> filter(grid == "v")
    )

    grid <- grid |> filter(grid == "h")
  }

  grid_plot <- grid |>
    ggplot() +
    geom_abline(
      mapping = aes(
        intercept = intercept,
        slope     = slope,
        linetype  = ref
      )
    )

  if (is.infinite(vec_tan[2])) grid_plot <- grid_plot + geom_vgrid

  grid_plot +
    scale_x_continuous(
      minor_breaks = NULL,
      breaks = x_labels |> pull(label_pos),
      labels = x_labels |> pull(coord),
      limits = x_limits,
      expand = expansion(),
      name   = NULL,
      sec.axis = dup_axis(name = "*&theta;*~2~", labels = NULL)
    ) +
    scale_y_continuous(
      minor_breaks = NULL,
      breaks = grid |> filter(grid == "h") |> pull(label_pos),
      labels = grid |> filter(grid == "h") |> pull(coord),
      limits = y_limits,
      expand = expansion(),
      name   = NULL,
      sec.axis = dup_axis(name = "*&theta;*~1~", labels = NULL)
    ) +
    scale_linetype_manual(values = linetypes, guide = NULL) +
    coord_fixed() +
    theme(
      axis.ticks   = element_blank(),
      axis.line    = element_blank(),
      axis.title.x       = element_markdown(hjust = y_axis_titpos),
      axis.title.y.right = element_markdown(vjust = x_axis_titpos)
    )
}

transform_grid <- function(basis,
                           x_limits = c(-4, 4),
                           y_limits = c(-4, 4),
                           break_step = 2,
                           linetype_axis = 'solid',
                           linetype_grid = 'dashed') {

  # Argument checking and formatting: ----

  # Main: ----
  basis |>
    create_grid(
      x_limits   = x_limits,
      y_limits   = y_limits,
      break_step = break_step
    ) |>
    plot_grid(
      x_limits   = x_limits,
      y_limits   = y_limits,
      break_step = break_step,
      linetype_axis = 'solid',
      linetype_grid = 'dashed'
    )
}
```

<br>

:::::: columns
::: {.column width="33%"}
<center>A</center>

```{r grid-orthogonal-rotation}
#| fig-width: 5
rot_basis <- cbind(
  c(sqrt(.75), -sqrt(.25)),
  c(sqrt(.25), sqrt(.75))
)

rot_grid <- transform_grid(rot_basis)

rot_grid
```
:::

::: {.column width="33%"}
<center>B</center>

```{r grid-ts-basis}
#| fig-width: 5
ts_grid <- transform_grid(ts_basis)

ts_grid
```
:::

::: {.column width="33%"}
<center>C</center>

```{r grid-ls-basis}
#| fig-width: 5
ls_grid <- transform_grid(ls_basis)

ls_grid
```
:::
::::::

[Answers in 3, 2, 1...](https://wall.sli.do/event/dPunB4BMx3S8hfSjaFSuFh?section=ae77dd00-34fb-4600-9677-337a65b04b83)

::: notes
So, here are three possible options; one of them is the correct one.

Take a few seconds to answer and we will see next which is the correct one.
:::

## Oblique axes

<br>

:::::::::: columns
::: {.column width="33%"}
<center>A</center>

```{r scatter-orthogonal-rotation}
#| fig-width: 5
thetas_transf_rot <- thetas |>
  group_by(i) |>
  pivot_longer(starts_with("trait")) |>
  mutate(value = rot_basis %*% value |> drop()) |>
  pivot_wider() |>
  ungroup()

rot_grid +
  geom_point(
    data    = thetas_transf_rot,
    mapping = aes(trait_1, trait_2),
    alpha   = .33,
    shape   = 16,
    colour = I(PALETTE_BLUE["normal"])
  )
```
:::

::::: {.column width="33%"}
<center>B</center>

:::: r-stack
```{r grid-ts-basis}
#| fig-width: 5
```

::: fragment
```{r scatter-ts-basis}
#| fig-width: 5
thetas_transf_wrong <- thetas |>
  group_by(i) |>
  pivot_longer(starts_with("trait")) |>
  mutate(value = ts_basis %*% value |> drop()) |>
  pivot_wider() |>
  ungroup()

ts_grid +
  geom_point(
    data    = thetas_transf_wrong,
    mapping = aes(trait_1, trait_2),
    alpha   = .33,
    shape   = 16,
    colour = I(PALETTE_BLUE["normal"])
  )
```
:::
::::
:::::

::::: {.column width="33%"}
<center>C</center>

:::: r-stack
```{r grid-ls-basis}
#| fig-width: 5
```

::: fragment
```{r scatter-ls-basis}
#| fig-width: 5
ls_scatter <- ls_grid +
  geom_point(
    data    = thetas_transf,
    mapping = aes(trait_1, trait_2),
    alpha   = .33,
    shape   = 16,
    colour = I(PALETTE_BLUE["normal"])
  )

ls_scatter
```
:::
::::
:::::
::::::::::

::: notes
A is pretty clearly not the answer, as it is just an orthogonal rotation.

Now, if you chose B, you probably thought that the axis units should not change (because the variances are unitary), and that the axes should form an angle which cosine is the correlation, that is, 0.5 (a 60º angle).

In that case, you think exactly like I do.

But if you picture this transformation in your mind, you will notice that when you make the axes closer, you will be squeezing the points in the distribution, making them look even more like a "sausage", so to speak.

\[NEXT\]

The resulting distribution is this; it took us hundreds of hours of work, during more than two years to realise this.

So if you want to actually "stretch" the cloud of points in the distribution, you need to pull the axes apart,...

\[NEXT\]

...as in the coordinate axes C.

But how is our intuition so wrong then?
Well, the truth is that it is not.
:::

## Latent space vs. test space {.smaller}

<br>

::::::: columns
:::: column
$$
  P(Y_i = 1) = \Phi_L (\mathbf{a}_i^T \mathbf{\theta} + d_i)
$$

::: {.fragment fragment-index="1"}
$$
  P(Y_i = 1) = \Phi_L ({\mathbf{a}^*_i}^T \mathbf{\theta}^* + d_i)
$$
:::
::::

:::: column
$$
  \mathbf{\theta}^* = \mathbf{P} \mathbf{\theta}
$$

::: {.fragment fragment-index="1"}
$$
  \mathbf{a}^*_i = \mathbf{Q} \mathbf{a}_i \\
$$
:::
::::
:::::::

::: {.fragment fragment-index="2"}
$$
  \mathbf{a}_i^T \mathbf{\theta} =
    {\mathbf{a}^*_i}^T \mathbf{\theta}^* =
    \mathbf{a}_i^T \mathbf{Q}^T \mathbf{P} \mathbf{\theta}
$$
:::

::: {.fragment fragment-index="2"}
$$
  \mathbf{Q} = (\mathbf{P}^{-1})^T
$$
:::

::: notes
The model probability must be invariant to a change of basis like this.

\[NEXT\]

So, the item vector must be transformed as well.

\[NEXT\]

And its transformation must be the opposite of the latent trait transformation.
:::

## Latent space vs. test space {.smaller}

<br>

::::: columns
::: column
<center>Latent space</center>

```{r scatter-latent-space}
#| fig-width: 6.5
ls_scatter
```
:::

::: column
<center>Test space</center>

```{r grid-test-space}
#| fig-width: 6.5
ts_grid
```
:::
:::::

::: notes
As it turns out, this gives a space where the axes actually do represent faithfully the latent covariances.
:::

## Latent space vs. test space {.smaller}

```{r space-inner-products}
# Mahalanobis distance:
TRAIT_VECTOR        <- latex_bf("\\theta")
TRAIT_NORM          <- latex_norm(TRAIT_VECTOR)
TRAIT_VECTOR_T      <- latex_transp(TRAIT_VECTOR)
TRAIT_COV_INV_TRAIT <- latex(TRAIT_VECTOR_T, COV_MATRIX_INV, TRAIT_VECTOR)
MAH_DIST_DEF        <- latex_sqrt(TRAIT_COV_INV_TRAIT)
MAH_DIST_EQ         <- latex_eq(TRAIT_NORM, MAH_DIST_DEF)

INNER_PROD_INV_COV_EQ <- latex_eq(COV_MATRIX_INV, INNER_PROD_TRANSF_DEF)

# Latent space:
TRAIT_VECTOR_1    <- latex_sub(TRAIT_VECTOR, 1)
TRAIT_VECTOR_2    <- latex_sub(TRAIT_VECTOR, 2)
TRAIT_VECTOR_1_T  <- latex_transp(TRAIT_VECTOR_1)

INNER_PROD     <- latex_innerprod(TRAIT_VECTOR_1, TRAIT_VECTOR_2)
INNER_PROD_DEF <- latex(TRAIT_VECTOR_1_T, COV_MATRIX_INV, TRAIT_VECTOR_2)
INNER_PROD_EQ  <- latex_eq(INNER_PROD, INNER_PROD_DEF)

# Test space:
DISCR_VECTOR_J <- latex_sub(DISCR_VECTOR_ANY, AUX_INDEX)
DISCR_VECTOR_T <- latex_transp(DISCR_VECTOR)

INNER_PROD_DISCR     <- latex_innerprod(DISCR_VECTOR, DISCR_VECTOR_J)
INNER_PROD_DISCR_DEF <- latex(DISCR_VECTOR_T, COV_MATRIX, DISCR_VECTOR_J)
INNER_PROD_DISCR_EQ  <- latex_eq(INNER_PROD_DISCR, INNER_PROD_DISCR_DEF)
```

<br>

::: fragment
> \[...\] the latent space should be regarded as a general Euclidean space with its *inner product* defined by $`r INNER_PROD_EQ`$.

--- @zhang_theoretical_1999 [p. 221]
:::

::: fragment
> \[...\] the *inner product* in \[the test space\] is consistently defined by
>
> $`r INNER_PROD_DISCR_EQ`$

--- @zhang_theoretical_1999 [p. 221]
:::

::: notes
This is actually not new, it was already noted by Zhang & Stout, more than 25 years ago, to solve a different problem, but unfortunately they didn't provide a deduction as we do here.

But the important thing here, is that we have two different spaces: One for the latent traits and one for the items.

And knowing these...
:::

## Generalized multidimensional parameters

<br>

$$
`r latex_eq(MDISC_COV_PARAM, DISCR_VECTOR_COV_MODULE)`
$$

<br>

::: fragment
$$
`r latex_eq(
     MIL_COV_PARAM,
     latex_curlybraces("$DISTANCE_COV_DEF$, $DIR_COS_ITEM_VEC_COV_DEF$")
   )`
$$
:::

::: notes
Si tenemos en cuenta las correlaciones (más generalmente, covarianzas) de la distribución de puntuaciones factoriales, lo que obtenemos son estas expresiones para los parámetros multidimensionales.

Muy importante, la distancia con signo y la discriminación multidimensional son iguales, no importa en qué espacio (latente, o test) queramos obtenerlas y representarlas.
Pero los cosenos directores es necesario calcularlos en el espacio test, no en el espacio de puntuaciones latentes, como decía Reckase, porque este es el espacio que va a representar en realidad la relación entre las variables latentes.
:::

## Complete graphical representation

```{r item-plot-out}
#| fig-cap:   Orthogonal (left) and oblique (right) test space
#| fig-align: center
#| fig-asp:     .5
#| fig-width: 13

plot_orth <- plot_orth +
  scale_color_manual(values = ITEMS_PALETTE) +
  theme(
    axis.ticks       = element_blank(),
    legend.title     = element_blank(),
    panel.grid.minor = element_blank()
  )

plot_oblique <- plot_oblique +
  scale_color_manual(values = ITEMS_PALETTE, guide = NULL) +
  theme(axis.ticks = element_blank())

plot_orth + plot_oblique ## TODO: Tema ggplot
```

## Conclusions

<br>

Generalization of multidimensional parameters

-   To constant-monotony items

-   To oblique axes

<br>

::: fragment
Distinction latent space - test space
:::

::: notes
En conclusión, lo que hemos propuesto en este artículo consiste en una generalización:

-   A ítems con monotonía constante, creciente, decreciente o nula.
    Esto se venía haciendo ya como hemos visto, pero creemos que hacía falta formalizarlo.
    No basta con hacerlo y punto, si Reckase dijo que las discriminaciones tiene que ser positivas, tenemos que demostrar que también vale para discriminaciones negativas o nulas (monotonía decreciente, o probabilidad constante).

-   A Ejes oblicuos, las cuales consideramos que son necesarias para representar distribuciones de puntuaciones de rasgo correlacionadas, en general.
    Hemos visto que esto ya lo mencionaron Zhang y Stout, hace 25 años, pero no proporcionaron una "deducción" de estos parámetros multidimensionales, cosa que nosotros hacemos.

\[NEXT\]

Por último, es imortante recalcar este "descubrimiento" (o "redescubrimiento", más bien) del espacio test, ya que tenemos que ser muy conscientes de que los ítems y las personas están en espacios vectoriales distintos.
Aunque esto parezca solamente un "artefacto matemático" (y en el fondo lo es), hemos visto que es verdaderamente crucial tenerlo en cuenta para entender bien los modelos que estamos utilizando.
Además, esta "distinción" se halla también en los modelos factoriales: Existe lo que se denominan "ejes primarios" y "ejes de referencia", propuestos por Thurstone hace años (pero que prácticamente no se utilizan ni tienen en consideración).
<!-- # TODO: Hace cuántos años? (Thurstone) --> Sin embargo, cuidado con esto, porque no está claro que esos espacios en TRI multidimensional se puedan asimilar directamente a los ejes en un modelo de AF.
:::

## To be continued...

<br>

-   Multidimensional IRT - FA parallelism

<br>

::: fragment
-   M2PL model extensions: Graded scale, nominal response, 3PL...
:::

<br>

::: fragment
-   Other models
:::

::: notes
Esta de hecho puede ser una futura línea de investigación.
Puede parecer evidente el paralelismo (o al menos a mí me lo parece), pero si una cosa me ha enseñado el estar trabajando en este problema de los parámetros (y toparme con el problema de los dos espacios), es que no hay que fiarse de las apariencias.
Por eso aplicamos el método científico (el método puramente deductivo, en este caso)

\[NEXT\]

El modelo multidimensional logístico de 2 parámetros se puede considerar como "anidado", o un caso más simple de varios modelos.
Así que partiendo de este como base, lo interesante es aplicarlo a otros más complejos, pero más útiles en la práctica.
Por cierto, que esto también lo hace mucha gente, aplica las fórmulas de Reckase a otros modelos y a correr, se quedan "tan anchos".

El modelo de escala graduada, por ejemplo, sería el equivalente al modelo de AF para respuestas ordinales, con varios umbrales o "posiciones" en TRI.

El de crédito parcial, por ejemplo, puede servir también para formatos con varias opciones de respuesta que miden o pesan en factores diferentes cada una de ella.

El logístico de 3 parámetros es como el de 2, pero añadiendo un parámetro de pseudoazar, para modelar el que un participante acierte un ítem por pura suerte, respondiendo "a ciegas" (esto sólo valdría para ítems cognitivos).

Estos modelos, que son "extensiones" de este más básico, deberían partir de estos parámetros como fundamento.
Pero un investigador no debería dar los parámetros sin más de ese modelo, sin proporcionar una prueba formal de que se aplica esa fórmula y se cumplen los supuestos.

\[NEXT\]

Y por supuesto, existen muchos otros modelos psicométricos, muy útiles y aplicados en la investigación en Psicología y Educación, en los cuales no está claro aún cómo generalizar estos resultados al ámbito multidimensional.
:::

## References {.smaller}