### 2025-09-17 output/Letter_to_editor-2.Rmd
Add Review 2 "letter to editor"

---
output:
  officedown::rdocx_document:
    mapstyles:
      First Paragraph: Body Text
    base_format: bookdown::word_document2
    reference_docx: ../www/Template.docx
    number_sections: no
    keep_md:         no
bibliography: ../www/Multidimensional-parameters-MCLM.bib
csl:          ../www/apa-old-doi-prefix.csl
editor_options: 
  chunk_output_type: console
---

October XX <!-- # TODO: Complete -->, 2025.

<br>

Sandip Sinharay, PhD \
Editor, *Psychometrika* \
Educational Testing Service, USA

<br>

Dear Dr. Sinharay:

Please find attached our second review of the manuscript
*"A generalized definition of multidimensional item response theory
parameters"*.
we have carefully taken into account Thissen's and your own
review and suggestions, addressing all the comments and suggestions diligently.
Where we have decided not to follow a certain suggestion,
we have striven to explain our reasons in the clearest possible way.
Also, besides addressing the comments,
we have added an additional clarification
about the "covariance matrix whitening",
a terminology previously unknown to us.

Once again, we thank you for considering our reviewed manuscript.
Please feel free to contact us again regarding any further concerns at
[dvmorillo@psi.uned.es](mailto:dvmorillo@psi.uned.es).

<br>

Sincerely,

Daniel Morillo, Ph.D.\
Department of Behavioral Science Methodology \
Universidad Nacional de Educación a Distancia (UNED) \
C. Juan del Rosal 10, Office 2.74 \
28040 Madrid, Spain