### 2025-10-01 notebooks/interp_example.qmd
Try `interp::interp()` example

---
title: "Untitled"
format: html
editor: visual
---

```{r}
library(interp)

data(franke)
## x-y irregular grid points:
oldseed <- set.seed(42)
ni <- 64
xi <- runif(ni,0,1)
yi <- runif(ni,0,1)
xyi <- cbind(xi,yi)
## linear interpolation
fi <- franke.fn(xi,yi,1)
IL <- interp(xi,yi,fi,nx=80,ny=80,method="linear")
## prepare breaks and colors that match for image and contour:
breaks <- pretty(seq(min(IL$z,na.rm=TRUE),max(IL$z,na.rm=TRUE),length=11))
db <- breaks[2]-breaks[1]
nb <- length(breaks)
breaks <- c(breaks[1]-db,breaks,breaks[nb]+db)
colors <- terrain.colors(length(breaks)-1)
image(IL,breaks=breaks,col=colors,main="Franke function 1",
      sub=paste("linear interpolation, ", ni,"points"))
contour(IL,add=TRUE,levels=breaks)
points(xi,yi)
## spline interpolation
fi <- franke.fn(xi,yi,1)
IS <- interp(xi,yi,fi,method="akima",
             kernel="gaussian",solver="QR")
## prepare breaks and colors that match for image and contour:
breaks <- pretty(seq(min(IS$z,na.rm=TRUE),max(IS$z,na.rm=TRUE),length=11))
db <- breaks[2]-breaks[1]
nb <- length(breaks)
breaks <- c(breaks[1]-db,breaks,breaks[nb]+db)
colors <- terrain.colors(length(breaks)-1)
image(IS,breaks=breaks,col=colors,main="Franke function 1",
      sub=paste("spline interpolation, ", ni,"points"))
contour(IS,add=TRUE,levels=breaks)
        points(xi,yi)
```