### 2025-10-01 notebooks/interp_example.qmd
Try "deldir" triangulation method

---
title: "Untitled"
format: html
editor: visual
---

```{r setup}
#| include: false
library(interp)

ROOT_DIR <- here::here()
knitr::opts_knit$set(root.dir = ROOT_DIR)

```

```{r sources}
#| include: false
source("src/Graphical_example_paper.R", encoding = "UTF-8")
```

```{r}
mvn_densities <- mvn_densities |>
  bind_cols(latent_space_grid_rot |> rename_with(~paste0(., '_rot')))

interp(
  mvn_densities$trait_1_rot,
  mvn_densities$trait_2_rot,
  mvn_densities$corr_norm,
  deltri = "deldir"
)
```