### 2025-10-01 notebooks/interp_example.qmd
Plot interpolated oblique contour plots
Computed by rotating the correlated distribution, and then interpolating.

---
title: "Untitled"
format: html
editor: visual
---

```{r setup}
#| include: false
library(interp)
library(magrittr)

ROOT_DIR <- here::here()
knitr::opts_knit$set(root.dir = ROOT_DIR)

```

```{r sources}
#| include: false
source("src/Graphical_example_paper.R", encoding = "UTF-8")
```

```{r}
mvn_densities <- mvn_densities |>
  bind_cols(latent_space_grid_rot |> rename_with(~paste0(., '_rot')))

mvn_dens_obl_interp <- interp(
  mvn_densities$trait_1_rot,
  mvn_densities$trait_2_rot,
  mvn_densities$corr_norm
)

mvn_dens_obl <- mvn_dens_obl_interp %$%
  expand_grid(trait_1 = x, trait_2 = y) |>
  add_column(dens = mvn_dens_obl_interp$z |> as.vector())

contour_obl <- mvn_dens_obl |> geom_contour(
  mapping     = aes(trait_1, trait_2, z = dens),
  color        = CONTOUR_COLOR,
  alpha        = .3,
  linewidth    = VECTOR_WIDTH,
  breaks       = CONTOUR_BREAKS
)

grid_oblique + contour_obl
```

These contour plots do not seem to be correct.
Do they actually represent a standard normal distribution with the intended covariance structure, in the test space?