### 2025-10-17 notebooks/ESM_latent-space-representation.Rmd
Draft "latent space plotting" supplementary file

---
output:
  officedown::rdocx_document:
    tables:
      style: APA_Like_style
      caption:
        style: Image Caption
        pre: "Table S."
        sep: ' '
    mapstyles:
      Body Text: First Paragraph
    base_format: bookdown::word_document2
    reference_docx: ../www/Template.docx
    number_sections: no
    keep_md:         no
    fig_width:       4
    fig_height:      4
bibliography: [../www/Multidimensional-parameters-MCLM.bib, ../www/packages.bib]
csl:          ../www/apa-old-doi-prefix.csl
editor_options: 
  chunk_output_type: console
---

```{r setup}
#| include: false

# Includes:

library(here)
library(knitr)
library(patchwork)
library(officer)

ROOT_DIR <- here()

opts_knit$set(root.dir = ROOT_DIR)
opts_chunk$set(
  echo     = FALSE,
  message  = FALSE,
  cache    = FALSE,
  dev      = "png",
  dpi      = 1200L,
  dev.args = list(type = "cairo-png"),
  fig.cap.fp_text = fp_text_lite(italic = FALSE, bold = TRUE)
)

# Output configuration options:
options(digits = 3)
```

```{r reset-root-wd}
#| cache: false

# Necessary for knitr to find the source file:
opts_knit$set(root.dir = ROOT_DIR)
```

```{r sources}
source("src/Graphical_example_paper.R", encoding = 'UTF-8')
```

# Representation in the latent space

```{r}
# Compute items in latent space (needs to be computed manually, because the
#   covariance matrix is computed from the transform matrix in
#   `compute_mirt_coords()` and the inverse matrix would be used instead):
items_ls <- items_oblique_coords |>
  pivot_longer(origin_1:end_2) |>
  separate(name, into = c('end', 'coord'), sep = "_", remove = FALSE) |>
  group_by(item, end) |>
  mutate(value = transform_matrix %*% value |> drop()) |>
  ungroup() |>
  pivot_wider(id_cols = item:end_transf_2, values_from = value)

## Compose latent space plot:

# Oblique grid:
grid_ls <- transform_grid(
  transform_matrix,
  x_limits = c(-3.2,  3.2),
  y_limits = c(-2.65, 2.65),
  break_step = 1,
  linetype_grid = "17",
  linewidth = LINE_WIDTH,
  axis_ticks = "axis"
)

# Oblique items:
items_geom_ls_obl <- geom_segment(
  mapping   = aes(
    origin_1, origin_2,
    xend  = end_1, yend = end_2,
    color = item
  ),
  data      = items_ls |> arrange(desc(item)), # Plot them in reverse order
  arrow     = arrow(angle = 20, length = unit(10, "points"), type = "closed"),
  linejoin  = "mitre",
  linewidth = VECTOR_WIDTH
)

plot_oblique_ls <- grid_ls +
  contour_orth +
  items_geom_ls_obl +
  colorscale
```

```{r items-latent-space-plot-out}
#| cache:          false
#| fig.topcaption: true
#| fig.height:     4.7
#| fig.width:      6.65
#| warning:        false
plot_oblique_ls
```