### 2025-10-29 output/Letter_to_editor-2.Rmd
Avoid referring to editor and reviewer directly
Refer to them by their role (also, correct reference to the editor, as it actually refers to the action editor instead)

---
output:
  officedown::rdocx_document:
    mapstyles:
      First Paragraph: Body Text
    base_format: bookdown::word_document2
    reference_docx: ../www/Template.docx
    number_sections: no
    keep_md:         no
bibliography: ../www/Multidimensional-parameters-MCLM.bib
csl:          ../www/apa-old-doi-prefix.csl
editor_options: 
  chunk_output_type: console
---

October XX <!-- # TODO: Complete -->, 2025.

<br>

Sandip Sinharay, PhD \
Editor, *Psychometrika* \
Educational Testing Service, USA

<br>

Dear Dr. Sinharay:

Please find attached our second review of the manuscript
*"A generalized definition of multidimensional item response theory
parameters"*.
we have carefully taken into account the reviewer and action editor's
review and suggestions, addressing all the comments and suggestions diligently.
Where we have decided not to follow a certain suggestion,
we have striven to explain our reasons in the clearest possible way.
Besides addressing the comments,
we have thoroughly reviewed the grammar and style,
in the hope of avoiding these kind of errors in our new submission,
and have also added an additional clarification
about the "covariance matrix whitening",
a terminology previously unknown to us.

Once again, we thank you for considering our reviewed manuscript.
Please feel free to contact us again regarding any further concerns at
[dvmorillo@psi.uned.es](mailto:dvmorillo@psi.uned.es).

<br>

Sincerely,

Daniel Morillo, Ph.D.\
Department of Behavioral Science Methodology \
Universidad Nacional de Educación a Distancia (UNED) \
C. Juan del Rosal 10, Office 2.74 \
28040 Madrid, Spain