### 2025-11-04 output/Title_page.Rmd

---
output:
  officedown::rdocx_document:
    mapstyles:
      First Paragraph: Body Text
    base_format: bookdown::word_document2
    reference_docx: ../www/Template.docx
    number_sections: no
    keep_md:         no
editor_options: 
  chunk_output_type: console
---

**Article title:**
A Generalized Definition of Multidimensional Item Response Theory Parameters

**Running title:** Generalized MIRT parameters

# List of authors

- Daniel Morillo-Cuadrado. Department of Behavioral Science Methodology,
  Faculty of Psychology, Universidad Nacional de Educación a Distancia (UNED),
  Spain.

- Mario Luzardo-Verde. Department of Mathematics and Applications,
  Universidad de La República, Uruguay.

## Corresponding author:

Daniel Morillo-Cuadrado \
Faculty of Psychology, Office 2.74. \
C. de Juan del Rosal 10, 28040 Madrid

Phone nº: +34 91398 6584 \
Email address: dvmorillo@psi.uned.es

# Acknowledgements

The authors would like to thank:
[Patricia Recio-Saboya, Ph.D.](https://orcid.org/0000-0002-0530-4404),
for her guidance regarding multidimensional estimation
in IRT software packages;
[Jay Verkuilen, Ph.D.](https://orcid.org/0000-0003-3546-3411),
for his comments and suggestions regarding covariance whitening;
David Thissen and three additional anonymous reviewers for
their valuable comments on early drafts of this manuscript.

# Data Availability Statement

All the materials and code of this paper are available as
a Software Heritage repository, under the following SWHID persistent identifier:

[swh:1:dir:6195936c6e16023c4a0b6a4500b81a34469d64f1;origin=https://github.com/DaniMori/mirt-parameters;visit=swh:1:snp:410cf1a6fb4dd4c10bdf0ce7f12f6890f560549d;anchor=swh:1:rev:1f1a708a7477dd61c0ba5c82dcf1d8523ab3d8ff](https://archive.softwareheritage.org/swh:1:dir:6195936c6e16023c4a0b6a4500b81a34469d64f1;origin=https://github.com/DaniMori/mirt-parameters;visit=swh:1:snp:410cf1a6fb4dd4c10bdf0ce7f12f6890f560549d;anchor=swh:1:rev:1f1a708a7477dd61c0ba5c82dcf1d8523ab3d8ff)

# Financial Support

This research received no specific grant funding form any funding agency,
commercial or not-for-profit sectors.

# Competing Interests Statement

Competing interests: The authors declare none.